---
title: "Retinalogik vs HFA study"
author: "[Eadie Technologies, Inc.](https://www.eadietech.com/)"
date: "2025-02-06"
output: 
  html_document:
    theme: yeti
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: 
      collapsed: true
    includes:
        after_body: footer.html
---

## Setup

### Versioning

Last updated on `r format(Sys.time(), '%Y-%b-%d')` at `r format(Sys.time(), '%I:%M %p')`.

* 2025-Feb-06: Created git & RStudio project.
* 2025-Feb-11: Added individual plots.
* 2025-Feb-18: Added cleaned up data and placeholder for global sensitivity indices.
* 2025-Feb-19: Added Bland-Altman plots for MD & PSD and scatterplots for pointwise sensitivity data. Also separated individual plots into tabs.
* 2025-Feb-20: Tabs were messing up slickR javascript. Removed for now.
* 2025-Feb-24: Added section for test duration.
* 2025-Feb-27: Added section for reliability indices.
* 2025-Mar-01: Added boxplots for FP, FN, and FL.
* 2025-Mar-14: Added newly collected data.
* 2025-Mar-18: Added Bland-Altman plots for the most recent data sets where reliability indices <30% and test-retest variability.
* 2025-Mar-25: Added Wilcoxon rank test results for pointwise differences.
* 2025-Mar-26: Added scatter and Bland-Altman plot for all locations.
* 2025-Apr-10: Added a hexbin plot.
* 2025-May-02: Added new data after extensive data extraction script changes to account for changing layouts.
* 2025-May-26: Added plot showing Deming regression line.
* 2025-Oct-14: Added links to all the other sub-pages under new subsection: Links.

### Acknowledgements
Data collected by Research Assistants Luc St-Pierre, Reann Post, and Research Coordinator Lynn Murphy.

```{r setup, collapse=TRUE}

# load packages
library(tidyverse)
library(here)
library(DT)
library(plotly)
library(viridis)
library(svglite)
library(htmltools)
library(gridExtra)

sessionInfo()

# default chunk options
knitr::opts_chunk$set(
  comment = '>', collapse = TRUE, cache = FALSE, dev= c("png")
  )

# load processed data
load(here("dBdat.Rda"))

```

# Study summary
We compared 24-2 visual field (VF) test results between Retinalogik and the Humphrey Field Analyzer (HFA) using SITA Standard test strategy in patients with early glaucoma and moderate to advanced glaucoma.

# Links to other subpages
As this page grows and grows, it's starting to get cluttered. The following sub-pages were created to compartmentalize different analyses:

* [COS stats and plots](https://vvnejl.github.io/retinalogik-study/cos.html)
* [Scotoma analysis](https://vvnejl.github.io/retinalogik-study/scotoma_analysis.html)

# Methods (briefly)
## Subjects

Eligible participants were identified among patients of Dr. Brennan Eadie at the Eadie Eye Centre. If deemed eligible for the study, subjects were recruited consecutively.

Each participant underwent five study visits. At each visit, they performed a VF test on both eyes using the two devices. The order of device tested was randomized at the first (baseline) study visit.

Exclusion criteria were:

* age <18 years
* diagnosis of secondary glaucoma
* diagnosis of non-glaucomatous optic neuropathy
* significant media opacity
* previous intraocular surgery
* pregnancy
* seizure disorder
* cardiac pacemaker or other implantable device
* severe vertigo or balance disturbance
* inability to demonstrate competence to make an informed decision regarding study participation

The study adhered to the tenets of the Declaration of Helsinki for research involving human subjects and the protocol was approved by the Nova Scotia Health Research Ethics Board (#1030608). All participants gave their written informed consent before enrollment in the study.

## Analysis
```{r demographics, echo=FALSE}
gender <- dBdat %>%
  distinct(id, gender) %>%
  count(gender)

vft <- dBdat %>%
  distinct(id, visit, eye) %>%
  count(eye)

# eyes <- dBdat %>%
#   distinct(id, eye) %>%
#   count(eye)
# 
# healthy <- dBdat %>%
#   subset(md >0) %>%
#   distinct(id, eye) %>%
#   count(eye)
# 
# early <- dBdat %>%
#   subset(md <0 & md > -6) %>%
#   distinct(id, eye) %>%
#   count(eye)
# 
# moderate <- dBdat %>%
#   subset(md < -6 & md > -12) %>%
#   distinct(id, eye) %>%
#   count(eye)
# 
# advanced <- dBdat %>%
#   subset(md < -12) %>%
#   distinct(id, eye) %>%
#   count(eye)

```

All OS data will be transposed to OD format before analyses. `r sum(vft$n)` VF tests from `r length(unique(dBdat$id))` participants (`r sum(gender$n[gender$gender == "F"], na.rm = T)` females, `r sum(gender$n[gender$gender == "M"], na.rm = T)` males) aged `r min(dBdat$age)` to `r max(dBdat$age)` (*M* = `r round(mean(dBdat$age), 2)`, *SD* = `r round(sd(dBdat$age), 2)`) were included in the analysis.

# Results

## Reliability indices {.tabset}
Reliability indices such as fixation losses (FL), false positives (FP), and false negatives (FN) were extracted for analysis.

```{r reliability, warning=FALSE}
library(rempsyc)
library(report)

reliabilityDat <- dBdat %>%
  group_by(device) %>%
  distinct(id, visit, eye, fpPerc, fnPerc, flPerc)

# descriptive stats
descriptive.data <- reliabilityDat %>%
  group_by(device) %>%
  summarize(across(fpPerc:flPerc,
    list(m = mean, sd = sd),
    .names = "{.col}.{.fn}"
  ))

reliabilityDat %>%
   group_by(device) %>% 
   summarize(across(fpPerc:flPerc,
                    list(~ str_c(round(mean(.), 2), " (", round(sd(.), 2), ")")))) %>%
  rename("Device" = device, "FP (SD)" = fpPerc_1, "FN (SD)" = fnPerc_1, "FL (SD)" = flPerc_1) %>%
  nice_table(note = c("Shown are the mean percentages and standard deviation (SD) for each device."))

fpTest <- wilcox.test(fpPerc ~ device, data = reliabilityDat)
# fpTest <- t.test(fpPerc ~ device, data = reliabilityDat)

fnTest <- wilcox.test(fnPerc ~ device, data = reliabilityDat)
flTest <- wilcox.test(flPerc ~ device, data = reliabilityDat)
# 
# stats.table <- report(fpTest)
# nice_table(fpTest, report = "t.test")
# t.test(flPerc ~ device, data = reliabilityDat, na.action = na.omit) 
  
```

### False Positives
Mann–Whitney U test showed that false positives were comparable between Retinalogik and HFA (p = `r if (fpTest$p.value < 0.01) ".01" else round(fpTest$p.value, 2)`)

```{r fp, warning=FALSE}
library(hrbrthemes)
library(viridis)

fpPlot <- reliabilityDat %>%
  ggplot( aes(x=device, y=fpPerc, fill=device, label=paste(id, visit, eye))) +
  geom_boxplot() +
  scale_fill_viridis(discrete = TRUE, alpha=0.6) +
  geom_jitter(color="black", size=0.4, alpha=0.9) +
  xlab("Device") +
  ylab("False Positives (%)") +
  theme_bw() +
  theme(legend.position="none")

ggplotly(fpPlot)

```

### False Negatives
Mann–Whitney U test showed that false negatives were significantly lower for HFA (p = `r if (fnTest$p.value < 0.01) ".01" else round(fnTest$p.value, 2)`)

```{r fn, warning=FALSE}
fnPlot <- reliabilityDat %>%
  ggplot( aes(x=device, y=fnPerc, fill=device, label=paste(id, visit, eye))) +
  geom_boxplot() +
  scale_fill_viridis(discrete = TRUE, alpha=0.6) +
  geom_jitter(color="black", size=0.4, alpha=0.9) +
  xlab("Device") +
  ylab("False Negatives (%)") +
  theme_bw() +
  theme(legend.position="none")

ggplotly(fnPlot)
```

### Fixation Losses
Mann–Whitney U test showed that fixation losses were significantly lower for Retinalogik compared to HFA (p < `r if (flTest$p.value < 0.01) ".01" else round(flTest$p.value, 2)`).

```{r fl, warning=FALSE}
flPlot <- reliabilityDat %>%
  ggplot( aes(x=device, y=flPerc, fill=device, label=paste(id, visit, eye))) +
  geom_boxplot() +
  scale_fill_viridis(discrete = TRUE, alpha=0.6) +
  geom_jitter(color="black", size=0.4, alpha=0.9) +
  xlab("Device") +
  ylab("Fixation Losses (%)") +
  theme_bw() +
  theme(legend.position="none")

ggplotly(flPlot)
```

## Individual plots
Swipe/scroll from right to left or click on the arrows for subsequent visits. Currently not plotting RL05 data where patient withdrew from the study.

### RL01

```{r rl01, warning=FALSE}
library(ggplot2)
library(slickR)

# make sure coordinates are numeric, then flip OS to OD for plotting purposes
dBdat %<>%
  mutate(x = as.numeric(x), y = as.numeric(y)) %>%
  mutate(x = case_when(
    eye == "L" & device == "hfa" ~ x*-1,
    TRUE ~ as.numeric(x)))

dBdat %>%
  mutate(dB = as.numeric(dB)) %>%
  filter(device == "hfa" & id == "RL01") %>%
  group_by(id, visit, eye) %>%
  arrange(id, visit, eye) %>%
  # Use the group_by %>% nest pattern to group data by id
  nest() %>% 
  # Use map2 so the id can be used as the title
  mutate(graphs = map2(data, id,
                       ~ggplot(data = .x, aes(x, y, dB)) +
                         geom_raster(aes(x = x, y = y, fill = dB)) +
                         geom_text(aes(label = dB, x = x, y = y), size = 8) +
                         coord_fixed(ratio = 1) +
                         scale_fill_gradientn(colours = viridis(47), limits = c(-1, 46),
                                              na.value="darkred") +
                         theme_bw() +
                         ggtitle(paste0("HFA ", id, " Visit ", visit, " Eye: ", eye))
                       )
         ) %>% 
  # pull is the pipe-able equivalent of .[['graphs']]
  pull(graphs) %>% 
  # Return the svg of graphs
  map(function(gr) svglite::xmlSVG(show(gr), standalone = TRUE)) -> 
  hfa_individual_graphs

dBdat %>%
  mutate(dB = as.numeric(dB)) %>%
  filter(device == "retinalogik" & id == "RL01") %>%
  group_by(id, visit, eye) %>%
  arrange(id, visit, eye) %>%
  # Use the group_by %>% nest pattern to group data by id
  nest() %>% 
  # Use map2 so the id can be used as the title
  mutate(graphs = map2(data, id,
                       ~ggplot(data = .x, aes(x, y, dB)) +
                         geom_raster(aes(x = x, y = y, fill = dB)) +
                         geom_text(aes(label = dB, x = x, y = y), size = 8) +
                         coord_fixed(ratio = 1) +
                         scale_fill_gradientn(colours = viridis(47), limits = c(-1, 46),
                                              na.value="darkred") +
                         theme_bw() +
                         ggtitle(paste0("Retinalogik ", id, " Visit ", visit, " Eye: ", eye))
                       )
         ) %>% 
  # pull is the pipe-able equivalent of .[['graphs']]
  pull(graphs) %>% 
  # Return the svg of graphs
  map(function(gr) svglite::xmlSVG(show(gr), standalone = TRUE)) -> 
  rt_individual_graphs

# carousels
hfa_plots <- slickR(hfa_individual_graphs, height = 350, width = "95%", padding = 0, slideId = "slick1") +
  settings(slidesToShow = 2, slidesToScroll = 2)

rt_plots <- slickR(rt_individual_graphs, height = 350, width = "95%", padding = 0, slideId = "slick2") +
  settings(slidesToShow = 2, slidesToScroll = 2)

rt_plots %synch% hfa_plots 

```

### RL02

```{r rl02, warning=FALSE}
dBdat %>%
  mutate(dB = as.numeric(dB)) %>%
  filter(device == "hfa" & id == "RL02") %>%
  group_by(id, visit, eye) %>%
  arrange(id, visit, eye) %>%
  # Use the group_by %>% nest pattern to group data by id
  nest() %>% 
  # Use map2 so the id can be used as the title
  mutate(graphs = map2(data, id,
                       ~ggplot(data = .x, aes(x, y, dB)) +
                         geom_raster(aes(x = x, y = y, fill = dB)) +
                         geom_text(aes(label = dB, x = x, y = y), size = 8) +
                         coord_fixed(ratio = 1) +
                         scale_fill_gradientn(colours = viridis(47), limits = c(-1, 46),
                                              na.value="darkred") +
                         theme_bw() +
                         ggtitle(paste0("HFA ", id, " Visit ", visit, " Eye: ", eye))
                       )
         ) %>% 
  # pull is the pipe-able equivalent of .[['graphs']]
  pull(graphs) %>% 
  # Return the svg of graphs
  map(function(gr) svglite::xmlSVG(show(gr), standalone = TRUE)) -> 
  hfa_individual_graphs

dBdat %>%
  mutate(dB = as.numeric(dB)) %>%
  filter(device == "retinalogik" & id == "RL02") %>%
  group_by(id, visit, eye) %>%
  arrange(id, visit, eye) %>%
  # Use the group_by %>% nest pattern to group data by id
  nest() %>% 
  # Use map2 so the id can be used as the title
  mutate(graphs = map2(data, id,
                       ~ggplot(data = .x, aes(x, y, dB)) +
                         geom_raster(aes(x = x, y = y, fill = dB)) +
                         geom_text(aes(label = dB, x = x, y = y), size = 8) +
                         coord_fixed(ratio = 1) +
                         scale_fill_gradientn(colours = viridis(47), limits = c(-1, 46),
                                              na.value="darkred") +
                         theme_bw() +
                         ggtitle(paste0("Retinalogik ", id, " Visit ", visit, " Eye: ", eye))
                       )
         ) %>% 
  # pull is the pipe-able equivalent of .[['graphs']]
  pull(graphs) %>% 
  # Return the svg of graphs
  map(function(gr) svglite::xmlSVG(show(gr), standalone = TRUE)) -> 
  rt_individual_graphs

# carousels
hfa_plots <- slickR(hfa_individual_graphs, height = 350, width = "95%", padding = 0, slideId = "slick3") +
  settings(slidesToShow = 2, slidesToScroll = 2)

rt_plots <- slickR(rt_individual_graphs, height = 350, width = "95%", padding = 0, slideId = "slick4") +
  settings(slidesToShow = 2, slidesToScroll = 2)

rt_plots %synch% hfa_plots
```

### RL03

```{r rl03, warning=FALSE}
dBdat %>%
  mutate(dB = as.numeric(dB)) %>%
  filter(device == "hfa" & id == "RL03") %>%
  group_by(id, visit, eye) %>%
  arrange(id, visit, eye) %>%
  # Use the group_by %>% nest pattern to group data by id
  nest() %>% 
  # Use map2 so the id can be used as the title
  mutate(graphs = map2(data, id,
                       ~ggplot(data = .x, aes(x, y, dB)) +
                         geom_raster(aes(x = x, y = y, fill = dB)) +
                         geom_text(aes(label = dB, x = x, y = y), size = 8) +
                         coord_fixed(ratio = 1) +
                         scale_fill_gradientn(colours = viridis(47), limits = c(-1, 46),
                                              na.value="darkred") +
                         theme_bw() +
                         ggtitle(paste0("HFA ", id, " Visit ", visit, " Eye: ", eye))
                       )
         ) %>% 
  # pull is the pipe-able equivalent of .[['graphs']]
  pull(graphs) %>% 
  # Return the svg of graphs
  map(function(gr) svglite::xmlSVG(show(gr), standalone = TRUE)) -> 
  hfa_individual_graphs

dBdat %>%
  mutate(dB = as.numeric(dB)) %>%
  filter(device == "retinalogik" & id == "RL03") %>%
  group_by(id, visit, eye) %>%
  arrange(id, visit, eye) %>%
  # Use the group_by %>% nest pattern to group data by id
  nest() %>% 
  # Use map2 so the id can be used as the title
  mutate(graphs = map2(data, id,
                       ~ggplot(data = .x, aes(x, y, dB)) +
                         geom_raster(aes(x = x, y = y, fill = dB)) +
                         geom_text(aes(label = dB, x = x, y = y), size = 8) +
                         coord_fixed(ratio = 1) +
                         scale_fill_gradientn(colours = viridis(47), limits = c(-1, 46),
                                              na.value="darkred") +
                         theme_bw() +
                         ggtitle(paste0("Retinalogik ", id, " Visit ", visit, " Eye: ", eye))
                       )
         ) %>% 
  # pull is the pipe-able equivalent of .[['graphs']]
  pull(graphs) %>% 
  # Return the svg of graphs
  map(function(gr) svglite::xmlSVG(show(gr), standalone = TRUE)) -> 
  rt_individual_graphs

# carousels
hfa_plots <- slickR(hfa_individual_graphs, height = 350, width = "95%", padding = 0) +
  settings(slidesToShow = 2, slidesToScroll = 2)

rt_plots <- slickR(rt_individual_graphs, height = 350, width = "95%", padding = 0) +
  settings(slidesToShow = 2, slidesToScroll = 2)

rt_plots %synch% hfa_plots 

```

### RL04

```{r rl04, warning=FALSE}
dBdat %>%
  mutate(dB = as.numeric(dB)) %>%
  filter(device == "hfa" & id == "RL04") %>%
  group_by(id, visit, eye) %>%
  arrange(id, visit, eye) %>%
  # Use the group_by %>% nest pattern to group data by id
  nest() %>% 
  # Use map2 so the id can be used as the title
  mutate(graphs = map2(data, id,
                       ~ggplot(data = .x, aes(x, y, dB)) +
                         geom_raster(aes(x = x, y = y, fill = dB)) +
                         geom_text(aes(label = dB, x = x, y = y), size = 8) +
                         coord_fixed(ratio = 1) +
                         scale_fill_gradientn(colours = viridis(47), limits = c(-1, 46),
                                              na.value="darkred") +
                         theme_bw() +
                         ggtitle(paste0("HFA ", id, " Visit ", visit, " Eye: ", eye))
                       )
         ) %>% 
  # pull is the pipe-able equivalent of .[['graphs']]
  pull(graphs) %>% 
  # Return the svg of graphs
  map(function(gr) svglite::xmlSVG(show(gr), standalone = TRUE)) -> 
  hfa_individual_graphs

dBdat %>%
  mutate(dB = as.numeric(dB)) %>%
  filter(device == "retinalogik" & id == "RL04") %>%
  group_by(id, visit, eye) %>%
  arrange(id, visit, eye) %>%
  # Use the group_by %>% nest pattern to group data by id
  nest() %>% 
  # Use map2 so the id can be used as the title
  mutate(graphs = map2(data, id,
                       ~ggplot(data = .x, aes(x, y, dB)) +
                         geom_raster(aes(x = x, y = y, fill = dB)) +
                         geom_text(aes(label = dB, x = x, y = y), size = 8) +
                         coord_fixed(ratio = 1) +
                         scale_fill_gradientn(colours = viridis(47), limits = c(-1, 46),
                                              na.value="darkred") +
                         theme_bw() +
                         ggtitle(paste0("Retinalogik ", id, " Visit ", visit, " Eye: ", eye))
                       )
         ) %>% 
  # pull is the pipe-able equivalent of .[['graphs']]
  pull(graphs) %>% 
  # Return the svg of graphs
  map(function(gr) svglite::xmlSVG(show(gr), standalone = TRUE)) -> 
  rt_individual_graphs

# carousels
hfa_plots <- slickR(hfa_individual_graphs, height = 350, width = "95%", padding = 0) +
  settings(slidesToShow = 2, slidesToScroll = 2)

rt_plots <- slickR(rt_individual_graphs, height = 350, width = "95%", padding = 0) +
  settings(slidesToShow = 2, slidesToScroll = 2)

rt_plots %synch% hfa_plots 

```

### RL06

```{r rl06, warning=FALSE}
dBdat %>%
  mutate(dB = as.numeric(dB)) %>%
  filter(device == "hfa" & id == "RL06") %>%
  group_by(id, visit, eye) %>%
  arrange(id, visit, eye) %>%
  # Use the group_by %>% nest pattern to group data by id
  nest() %>% 
  # Use map2 so the id can be used as the title
  mutate(graphs = map2(data, id,
                       ~ggplot(data = .x, aes(x, y, dB)) +
                         geom_raster(aes(x = x, y = y, fill = dB)) +
                         geom_text(aes(label = dB, x = x, y = y), size = 8) +
                         coord_fixed(ratio = 1) +
                         scale_fill_gradientn(colours = viridis(47), limits = c(-1, 46),
                                              na.value="darkred") +
                         theme_bw() +
                         ggtitle(paste0("HFA ", id, " Visit ", visit, " Eye: ", eye))
                       )
         ) %>% 
  # pull is the pipe-able equivalent of .[['graphs']]
  pull(graphs) %>% 
  # Return the svg of graphs
  map(function(gr) svglite::xmlSVG(show(gr), standalone = TRUE)) -> 
  hfa_individual_graphs

dBdat %>%
  mutate(dB = as.numeric(dB)) %>%
  filter(device == "retinalogik" & id == "RL06") %>%
  group_by(id, visit, eye) %>%
  arrange(id, visit, eye) %>%
  # Use the group_by %>% nest pattern to group data by id
  nest() %>% 
  # Use map2 so the id can be used as the title
  mutate(graphs = map2(data, id,
                       ~ggplot(data = .x, aes(x, y, dB)) +
                         geom_raster(aes(x = x, y = y, fill = dB)) +
                         geom_text(aes(label = dB, x = x, y = y), size = 8) +
                         coord_fixed(ratio = 1) +
                         scale_fill_gradientn(colours = viridis(47), limits = c(-1, 46),
                                              na.value="darkred") +
                         theme_bw() +
                         ggtitle(paste0("Retinalogik ", id, " Visit ", visit, " Eye: ", eye))
                       )
         ) %>% 
  # pull is the pipe-able equivalent of .[['graphs']]
  pull(graphs) %>% 
  # Return the svg of graphs
  map(function(gr) svglite::xmlSVG(show(gr), standalone = TRUE)) -> 
  rt_individual_graphs

# carousels
hfa_plots <- slickR(hfa_individual_graphs, height = 350, width = "95%", padding = 0) +
  settings(slidesToShow = 2, slidesToScroll = 2)

rt_plots <- slickR(rt_individual_graphs, height = 350, width = "95%", padding = 0) +
  settings(slidesToShow = 2, slidesToScroll = 2)

rt_plots %synch% hfa_plots 
```

### RL07

```{r rl07, warning=FALSE}
# make sure coordinates are numeric, then flip OS to OD for plotting purposes
dBdat %>%
  mutate(dB = as.numeric(dB)) %>%
  filter(device == "hfa" & id == "RL07") %>%
  group_by(id, visit, eye) %>%
  arrange(id, visit, eye) %>%
  # Use the group_by %>% nest pattern to group data by id
  nest() %>% 
  # Use map2 so the id can be used as the title
  mutate(graphs = map2(data, id,
                       ~ggplot(data = .x, aes(x, y, dB)) +
                         geom_raster(aes(x = x, y = y, fill = dB)) +
                         geom_text(aes(label = dB, x = x, y = y), size = 8) +
                         coord_fixed(ratio = 1) +
                         scale_fill_gradientn(colours = viridis(47), limits = c(-1, 46),
                                              na.value="darkred") +
                         theme_bw() +
                         ggtitle(paste0("HFA ", id, " Visit ", visit, " Eye: ", eye))
                       )
         ) %>% 
  # pull is the pipe-able equivalent of .[['graphs']]
  pull(graphs) %>% 
  # Return the svg of graphs
  map(function(gr) svglite::xmlSVG(show(gr), standalone = TRUE)) -> 
  hfa_individual_graphs

dBdat %>%
  mutate(dB = as.numeric(dB)) %>%
  filter(device == "retinalogik" & id == "RL07") %>%
  group_by(id, visit, eye) %>%
  arrange(id, visit, eye) %>%
  # Use the group_by %>% nest pattern to group data by id
  nest() %>% 
  # Use map2 so the id can be used as the title
  mutate(graphs = map2(data, id,
                       ~ggplot(data = .x, aes(x, y, dB)) +
                         geom_raster(aes(x = x, y = y, fill = dB)) +
                         geom_text(aes(label = dB, x = x, y = y), size = 8) +
                         coord_fixed(ratio = 1) +
                         scale_fill_gradientn(colours = viridis(47), limits = c(-1, 46),
                                              na.value="darkred") +
                         theme_bw() +
                         ggtitle(paste0("Retinalogik ", id, " Visit ", visit, " Eye: ", eye))
                       )
         ) %>% 
  # pull is the pipe-able equivalent of .[['graphs']]
  pull(graphs) %>% 
  # Return the svg of graphs
  map(function(gr) svglite::xmlSVG(show(gr), standalone = TRUE)) -> 
  rt_individual_graphs

# carousels
hfa_plots <- slickR(hfa_individual_graphs, height = 350, width = "95%", padding = 0) +
  settings(slidesToShow = 1, slidesToScroll = 1)

rt_plots <- slickR(rt_individual_graphs, height = 350, width = "95%", padding = 0) +
  settings(slidesToShow = 1, slidesToScroll = 1)

rt_plots %synch% hfa_plots 
```

### RL08

```{r rl08, warning=FALSE}
# make sure coordinates are numeric, then flip OS to OD for plotting purposes
dBdat %>%
  mutate(dB = as.numeric(dB)) %>%
  filter(device == "hfa" & id == "RL08") %>%
  group_by(id, visit, eye) %>%
  arrange(id, visit, eye) %>%
  # Use the group_by %>% nest pattern to group data by id
  nest() %>% 
  # Use map2 so the id can be used as the title
  mutate(graphs = map2(data, id,
                       ~ggplot(data = .x, aes(x, y, dB)) +
                         geom_raster(aes(x = x, y = y, fill = dB)) +
                         geom_text(aes(label = dB, x = x, y = y), size = 8) +
                         coord_fixed(ratio = 1) +
                         scale_fill_gradientn(colours = viridis(47), limits = c(-1, 46),
                                              na.value="darkred") +
                         theme_bw() +
                         ggtitle(paste0("HFA ", id, " Visit ", visit, " Eye: ", eye))
                       )
         ) %>% 
  # pull is the pipe-able equivalent of .[['graphs']]
  pull(graphs) %>% 
  # Return the svg of graphs
  map(function(gr) svglite::xmlSVG(show(gr), standalone = TRUE)) -> 
  hfa_individual_graphs

dBdat %>%
  mutate(dB = as.numeric(dB)) %>%
  filter(device == "retinalogik" & id == "RL08") %>%
  group_by(id, visit, eye) %>%
  arrange(id, visit, eye) %>%
  # Use the group_by %>% nest pattern to group data by id
  nest() %>% 
  # Use map2 so the id can be used as the title
  mutate(graphs = map2(data, id,
                       ~ggplot(data = .x, aes(x, y, dB)) +
                         geom_raster(aes(x = x, y = y, fill = dB)) +
                         geom_text(aes(label = dB, x = x, y = y), size = 8) +
                         coord_fixed(ratio = 1) +
                         scale_fill_gradientn(colours = viridis(47), limits = c(-1, 46),
                                              na.value="darkred") +
                         theme_bw() +
                         ggtitle(paste0("Retinalogik ", id, " Visit ", visit, " Eye: ", eye))
                       )
         ) %>% 
  # pull is the pipe-able equivalent of .[['graphs']]
  pull(graphs) %>% 
  # Return the svg of graphs
  map(function(gr) svglite::xmlSVG(show(gr), standalone = TRUE)) -> 
  rt_individual_graphs

# carousels
hfa_plots <- slickR(hfa_individual_graphs, height = 350, width = "95%", padding = 0) +
  settings(slidesToShow = 2, slidesToScroll = 2)

rt_plots <- slickR(rt_individual_graphs, height = 350, width = "95%", padding = 0) +
  settings(slidesToShow = 2, slidesToScroll = 2)

rt_plots %synch% hfa_plots 
```

### RL09

```{r rl09, warning=FALSE}
# make sure coordinates are numeric, then flip OS to OD for plotting purposes
dBdat %>%
  mutate(dB = as.numeric(dB)) %>%
  filter(device == "hfa" & id == "RL09") %>%
  group_by(id, visit, eye) %>%
  arrange(id, visit, eye) %>%
  # Use the group_by %>% nest pattern to group data by id
  nest() %>% 
  # Use map2 so the id can be used as the title
  mutate(graphs = map2(data, id,
                       ~ggplot(data = .x, aes(x, y, dB)) +
                         geom_raster(aes(x = x, y = y, fill = dB)) +
                         geom_text(aes(label = dB, x = x, y = y), size = 8) +
                         coord_fixed(ratio = 1) +
                         scale_fill_gradientn(colours = viridis(47), limits = c(-1, 46),
                                              na.value="darkred") +
                         theme_bw() +
                         ggtitle(paste0("HFA ", id, " Visit ", visit, " Eye: ", eye))
                       )
         ) %>% 
  # pull is the pipe-able equivalent of .[['graphs']]
  pull(graphs) %>% 
  # Return the svg of graphs
  map(function(gr) svglite::xmlSVG(show(gr), standalone = TRUE)) -> 
  hfa_individual_graphs

dBdat %>%
  mutate(dB = as.numeric(dB)) %>%
  filter(device == "retinalogik" & id == "RL09") %>%
  group_by(id, visit, eye) %>%
  arrange(id, visit, eye) %>%
  # Use the group_by %>% nest pattern to group data by id
  nest() %>% 
  # Use map2 so the id can be used as the title
  mutate(graphs = map2(data, id,
                       ~ggplot(data = .x, aes(x, y, dB)) +
                         geom_raster(aes(x = x, y = y, fill = dB)) +
                         geom_text(aes(label = dB, x = x, y = y), size = 8) +
                         coord_fixed(ratio = 1) +
                         scale_fill_gradientn(colours = viridis(47), limits = c(-1, 46),
                                              na.value="darkred") +
                         theme_bw() +
                         ggtitle(paste0("Retinalogik ", id, " Visit ", visit, " Eye: ", eye))
                       )
         ) %>% 
  # pull is the pipe-able equivalent of .[['graphs']]
  pull(graphs) %>% 
  # Return the svg of graphs
  map(function(gr) svglite::xmlSVG(show(gr), standalone = TRUE)) -> 
  rt_individual_graphs

# carousels
hfa_plots <- slickR(hfa_individual_graphs, height = 350, width = "95%", padding = 0) +
  settings(slidesToShow = 2, slidesToScroll = 2)

rt_plots <- slickR(rt_individual_graphs, height = 350, width = "95%", padding = 0) +
  settings(slidesToShow = 2, slidesToScroll = 2)

rt_plots %synch% hfa_plots 
```

### RL10

```{r rl10, warning=FALSE}
# make sure coordinates are numeric, then flip OS to OD for plotting purposes
dBdat %>%
  mutate(dB = as.numeric(dB)) %>%
  filter(device == "hfa" & id == "RL10") %>%
  group_by(id, visit, eye) %>%
  arrange(id, visit, eye) %>%
  # Use the group_by %>% nest pattern to group data by id
  nest() %>% 
  # Use map2 so the id can be used as the title
  mutate(graphs = map2(data, id,
                       ~ggplot(data = .x, aes(x, y, dB)) +
                         geom_raster(aes(x = x, y = y, fill = dB)) +
                         geom_text(aes(label = dB, x = x, y = y), size = 8) +
                         coord_fixed(ratio = 1) +
                         scale_fill_gradientn(colours = viridis(47), limits = c(-1, 46),
                                              na.value="darkred") +
                         theme_bw() +
                         ggtitle(paste0("HFA ", id, " Visit ", visit, " Eye: ", eye))
                       )
         ) %>% 
  # pull is the pipe-able equivalent of .[['graphs']]
  pull(graphs) %>% 
  # Return the svg of graphs
  map(function(gr) svglite::xmlSVG(show(gr), standalone = TRUE)) -> 
  hfa_individual_graphs

dBdat %>%
  mutate(dB = as.numeric(dB)) %>%
  filter(device == "retinalogik" & id == "RL10") %>%
  group_by(id, visit, eye) %>%
  arrange(id, visit, eye) %>%
  # Use the group_by %>% nest pattern to group data by id
  nest() %>% 
  # Use map2 so the id can be used as the title
  mutate(graphs = map2(data, id,
                       ~ggplot(data = .x, aes(x, y, dB)) +
                         geom_raster(aes(x = x, y = y, fill = dB)) +
                         geom_text(aes(label = dB, x = x, y = y), size = 8) +
                         coord_fixed(ratio = 1) +
                         scale_fill_gradientn(colours = viridis(47), limits = c(-1, 46),
                                              na.value="darkred") +
                         theme_bw() +
                         ggtitle(paste0("Retinalogik ", id, " Visit ", visit, " Eye: ", eye))
                       )
         ) %>% 
  # pull is the pipe-able equivalent of .[['graphs']]
  pull(graphs) %>% 
  # Return the svg of graphs
  map(function(gr) svglite::xmlSVG(show(gr), standalone = TRUE)) -> 
  rt_individual_graphs

# carousels
hfa_plots <- slickR(hfa_individual_graphs, height = 350, width = "95%", padding = 0) +
  settings(slidesToShow = 2, slidesToScroll = 2)

rt_plots <- slickR(rt_individual_graphs, height = 350, width = "95%", padding = 0) +
  settings(slidesToShow = 2, slidesToScroll = 2)

rt_plots %synch% hfa_plots 
```

### RL11

```{r rl11, warning=FALSE}
# make sure coordinates are numeric, then flip OS to OD for plotting purposes
dBdat %>%
  mutate(dB = as.numeric(dB)) %>%
  filter(device == "hfa" & id == "RL11") %>%
  group_by(id, visit, eye) %>%
  arrange(id, visit, eye) %>%
  # Use the group_by %>% nest pattern to group data by id
  nest() %>% 
  # Use map2 so the id can be used as the title
  mutate(graphs = map2(data, id,
                       ~ggplot(data = .x, aes(x, y, dB)) +
                         geom_raster(aes(x = x, y = y, fill = dB)) +
                         geom_text(aes(label = dB, x = x, y = y), size = 8) +
                         coord_fixed(ratio = 1) +
                         scale_fill_gradientn(colours = viridis(47), limits = c(-1, 46),
                                              na.value="darkred") +
                         theme_bw() +
                         ggtitle(paste0("HFA ", id, " Visit ", visit, " Eye: ", eye))
                       )
         ) %>% 
  # pull is the pipe-able equivalent of .[['graphs']]
  pull(graphs) %>% 
  # Return the svg of graphs
  map(function(gr) svglite::xmlSVG(show(gr), standalone = TRUE)) -> 
  hfa_individual_graphs

dBdat %>%
  mutate(dB = as.numeric(dB)) %>%
  filter(device == "retinalogik" & id == "RL11") %>%
  group_by(id, visit, eye) %>%
  arrange(id, visit, eye) %>%
  # Use the group_by %>% nest pattern to group data by id
  nest() %>% 
  # Use map2 so the id can be used as the title
  mutate(graphs = map2(data, id,
                       ~ggplot(data = .x, aes(x, y, dB)) +
                         geom_raster(aes(x = x, y = y, fill = dB)) +
                         geom_text(aes(label = dB, x = x, y = y), size = 8) +
                         coord_fixed(ratio = 1) +
                         scale_fill_gradientn(colours = viridis(47), limits = c(-1, 46),
                                              na.value="darkred") +
                         theme_bw() +
                         ggtitle(paste0("Retinalogik ", id, " Visit ", visit, " Eye: ", eye))
                       )
         ) %>% 
  # pull is the pipe-able equivalent of .[['graphs']]
  pull(graphs) %>% 
  # Return the svg of graphs
  map(function(gr) svglite::xmlSVG(show(gr), standalone = TRUE)) -> 
  rt_individual_graphs

# carousels
hfa_plots <- slickR(hfa_individual_graphs, height = 350, width = "95%", padding = 0) +
  settings(slidesToShow = 2, slidesToScroll = 2)

rt_plots <- slickR(rt_individual_graphs, height = 350, width = "95%", padding = 0) +
  settings(slidesToShow = 2, slidesToScroll = 2)

rt_plots %synch% hfa_plots 
```

### RL12

```{r rl12, warning=FALSE}
# make sure coordinates are numeric, then flip OS to OD for plotting purposes
dBdat %>%
  mutate(dB = as.numeric(dB)) %>%
  filter(device == "hfa" & id == "RL12") %>%
  group_by(id, visit, eye) %>%
  arrange(id, visit, eye) %>%
  # Use the group_by %>% nest pattern to group data by id
  nest() %>% 
  # Use map2 so the id can be used as the title
  mutate(graphs = map2(data, id,
                       ~ggplot(data = .x, aes(x, y, dB)) +
                         geom_raster(aes(x = x, y = y, fill = dB)) +
                         geom_text(aes(label = dB, x = x, y = y), size = 8) +
                         coord_fixed(ratio = 1) +
                         scale_fill_gradientn(colours = viridis(47), limits = c(-1, 46),
                                              na.value="darkred") +
                         theme_bw() +
                         ggtitle(paste0("HFA ", id, " Visit ", visit, " Eye: ", eye))
                       )
         ) %>% 
  # pull is the pipe-able equivalent of .[['graphs']]
  pull(graphs) %>% 
  # Return the svg of graphs
  map(function(gr) svglite::xmlSVG(show(gr), standalone = TRUE)) -> 
  hfa_individual_graphs

dBdat %>%
  mutate(dB = as.numeric(dB)) %>%
  filter(device == "retinalogik" & id == "RL12") %>%
  group_by(id, visit, eye) %>%
  arrange(id, visit, eye) %>%
  # Use the group_by %>% nest pattern to group data by id
  nest() %>% 
  # Use map2 so the id can be used as the title
  mutate(graphs = map2(data, id,
                       ~ggplot(data = .x, aes(x, y, dB)) +
                         geom_raster(aes(x = x, y = y, fill = dB)) +
                         geom_text(aes(label = dB, x = x, y = y), size = 8) +
                         coord_fixed(ratio = 1) +
                         scale_fill_gradientn(colours = viridis(47), limits = c(-1, 46),
                                              na.value="darkred") +
                         theme_bw() +
                         ggtitle(paste0("Retinalogik ", id, " Visit ", visit, " Eye: ", eye))
                       )
         ) %>% 
  # pull is the pipe-able equivalent of .[['graphs']]
  pull(graphs) %>% 
  # Return the svg of graphs
  map(function(gr) svglite::xmlSVG(show(gr), standalone = TRUE)) -> 
  rt_individual_graphs

# carousels
hfa_plots <- slickR(hfa_individual_graphs, height = 350, width = "95%", padding = 0) +
  settings(slidesToShow = 2, slidesToScroll = 2)

rt_plots <- slickR(rt_individual_graphs, height = 350, width = "95%", padding = 0) +
  settings(slidesToShow = 2, slidesToScroll = 2)

rt_plots %synch% hfa_plots 
```

### RL13

```{r rl13, warning=FALSE}
# make sure coordinates are numeric, then flip OS to OD for plotting purposes
dBdat %>%
  mutate(dB = as.numeric(dB)) %>%
  filter(device == "hfa" & id == "RL13") %>%
  group_by(id, visit, eye) %>%
  arrange(id, visit, eye) %>%
  # Use the group_by %>% nest pattern to group data by id
  nest() %>% 
  # Use map2 so the id can be used as the title
  mutate(graphs = map2(data, id,
                       ~ggplot(data = .x, aes(x, y, dB)) +
                         geom_raster(aes(x = x, y = y, fill = dB)) +
                         geom_text(aes(label = dB, x = x, y = y), size = 8) +
                         coord_fixed(ratio = 1) +
                         scale_fill_gradientn(colours = viridis(47), limits = c(-1, 46),
                                              na.value="darkred") +
                         theme_bw() +
                         ggtitle(paste0("HFA ", id, " Visit ", visit, " Eye: ", eye))
                       )
         ) %>% 
  # pull is the pipe-able equivalent of .[['graphs']]
  pull(graphs) %>% 
  # Return the svg of graphs
  map(function(gr) svglite::xmlSVG(show(gr), standalone = TRUE)) -> 
  hfa_individual_graphs

dBdat %>%
  mutate(dB = as.numeric(dB)) %>%
  filter(device == "retinalogik" & id == "RL13") %>%
  group_by(id, visit, eye) %>%
  arrange(id, visit, eye) %>%
  # Use the group_by %>% nest pattern to group data by id
  nest() %>% 
  # Use map2 so the id can be used as the title
  mutate(graphs = map2(data, id,
                       ~ggplot(data = .x, aes(x, y, dB)) +
                         geom_raster(aes(x = x, y = y, fill = dB)) +
                         geom_text(aes(label = dB, x = x, y = y), size = 8) +
                         coord_fixed(ratio = 1) +
                         scale_fill_gradientn(colours = viridis(47), limits = c(-1, 46),
                                              na.value="darkred") +
                         theme_bw() +
                         ggtitle(paste0("Retinalogik ", id, " Visit ", visit, " Eye: ", eye))
                       )
         ) %>% 
  # pull is the pipe-able equivalent of .[['graphs']]
  pull(graphs) %>% 
  # Return the svg of graphs
  map(function(gr) svglite::xmlSVG(show(gr), standalone = TRUE)) -> 
  rt_individual_graphs

# carousels
hfa_plots <- slickR(hfa_individual_graphs, height = 350, width = "95%", padding = 0) +
  settings(slidesToShow = 2, slidesToScroll = 2)

rt_plots <- slickR(rt_individual_graphs, height = 350, width = "95%", padding = 0) +
  settings(slidesToShow = 2, slidesToScroll = 2)

rt_plots %synch% hfa_plots 
```

### RL14

```{r rl14, warning=FALSE}
# make sure coordinates are numeric, then flip OS to OD for plotting purposes
dBdat %>%
  mutate(dB = as.numeric(dB)) %>%
  filter(device == "hfa" & id == "RL14") %>%
  group_by(id, visit, eye) %>%
  arrange(id, visit, eye) %>%
  # Use the group_by %>% nest pattern to group data by id
  nest() %>% 
  # Use map2 so the id can be used as the title
  mutate(graphs = map2(data, id,
                       ~ggplot(data = .x, aes(x, y, dB)) +
                         geom_raster(aes(x = x, y = y, fill = dB)) +
                         geom_text(aes(label = dB, x = x, y = y), size = 8) +
                         coord_fixed(ratio = 1) +
                         scale_fill_gradientn(colours = viridis(47), limits = c(-1, 46),
                                              na.value="darkred") +
                         theme_bw() +
                         ggtitle(paste0("HFA ", id, " Visit ", visit, " Eye: ", eye))
                       )
         ) %>% 
  # pull is the pipe-able equivalent of .[['graphs']]
  pull(graphs) %>% 
  # Return the svg of graphs
  map(function(gr) svglite::xmlSVG(show(gr), standalone = TRUE)) -> 
  hfa_individual_graphs

dBdat %>%
  mutate(dB = as.numeric(dB)) %>%
  filter(device == "retinalogik" & id == "RL14") %>%
  group_by(id, visit, eye) %>%
  arrange(id, visit, eye) %>%
  # Use the group_by %>% nest pattern to group data by id
  nest() %>% 
  # Use map2 so the id can be used as the title
  mutate(graphs = map2(data, id,
                       ~ggplot(data = .x, aes(x, y, dB)) +
                         geom_raster(aes(x = x, y = y, fill = dB)) +
                         geom_text(aes(label = dB, x = x, y = y), size = 8) +
                         coord_fixed(ratio = 1) +
                         scale_fill_gradientn(colours = viridis(47), limits = c(-1, 46),
                                              na.value="darkred") +
                         theme_bw() +
                         ggtitle(paste0("Retinalogik ", id, " Visit ", visit, " Eye: ", eye))
                       )
         ) %>% 
  # pull is the pipe-able equivalent of .[['graphs']]
  pull(graphs) %>% 
  # Return the svg of graphs
  map(function(gr) svglite::xmlSVG(show(gr), standalone = TRUE)) -> 
  rt_individual_graphs

# carousels
hfa_plots <- slickR(hfa_individual_graphs, height = 350, width = "95%", padding = 0) +
  settings(slidesToShow = 2, slidesToScroll = 2)

rt_plots <- slickR(rt_individual_graphs, height = 350, width = "95%", padding = 0) +
  settings(slidesToShow = 2, slidesToScroll = 2)

rt_plots %synch% hfa_plots 
```

### RL15

```{r rl15, warning=FALSE}
# make sure coordinates are numeric, then flip OS to OD for plotting purposes
dBdat %>%
  mutate(dB = as.numeric(dB)) %>%
  filter(device == "hfa" & id == "RL15") %>%
  group_by(id, visit, eye) %>%
  arrange(id, visit, eye) %>%
  # Use the group_by %>% nest pattern to group data by id
  nest() %>% 
  # Use map2 so the id can be used as the title
  mutate(graphs = map2(data, id,
                       ~ggplot(data = .x, aes(x, y, dB)) +
                         geom_raster(aes(x = x, y = y, fill = dB)) +
                         geom_text(aes(label = dB, x = x, y = y), size = 8) +
                         coord_fixed(ratio = 1) +
                         scale_fill_gradientn(colours = viridis(47), limits = c(-1, 46),
                                              na.value="darkred") +
                         theme_bw() +
                         ggtitle(paste0("HFA ", id, " Visit ", visit, " Eye: ", eye))
                       )
         ) %>% 
  # pull is the pipe-able equivalent of .[['graphs']]
  pull(graphs) %>% 
  # Return the svg of graphs
  map(function(gr) svglite::xmlSVG(show(gr), standalone = TRUE)) -> 
  hfa_individual_graphs

dBdat %>%
  mutate(dB = as.numeric(dB)) %>%
  filter(device == "retinalogik" & id == "RL15") %>%
  group_by(id, visit, eye) %>%
  arrange(id, visit, eye) %>%
  # Use the group_by %>% nest pattern to group data by id
  nest() %>% 
  # Use map2 so the id can be used as the title
  mutate(graphs = map2(data, id,
                       ~ggplot(data = .x, aes(x, y, dB)) +
                         geom_raster(aes(x = x, y = y, fill = dB)) +
                         geom_text(aes(label = dB, x = x, y = y), size = 8) +
                         coord_fixed(ratio = 1) +
                         scale_fill_gradientn(colours = viridis(47), limits = c(-1, 46),
                                              na.value="darkred") +
                         theme_bw() +
                         ggtitle(paste0("Retinalogik ", id, " Visit ", visit, " Eye: ", eye))
                       )
         ) %>% 
  # pull is the pipe-able equivalent of .[['graphs']]
  pull(graphs) %>% 
  # Return the svg of graphs
  map(function(gr) svglite::xmlSVG(show(gr), standalone = TRUE)) -> 
  rt_individual_graphs

# carousels
hfa_plots <- slickR(hfa_individual_graphs, height = 350, width = "95%", padding = 0) +
  settings(slidesToShow = 2, slidesToScroll = 2)

rt_plots <- slickR(rt_individual_graphs, height = 350, width = "95%", padding = 0) +
  settings(slidesToShow = 2, slidesToScroll = 2)

rt_plots %synch% hfa_plots 
```

### RL16

```{r rl16, warning=FALSE}
# make sure coordinates are numeric, then flip OS to OD for plotting purposes
dBdat %>%
  mutate(dB = as.numeric(dB)) %>%
  filter(device == "hfa" & id == "RL16") %>%
  group_by(id, visit, eye) %>%
  arrange(id, visit, eye) %>%
  # Use the group_by %>% nest pattern to group data by id
  nest() %>% 
  # Use map2 so the id can be used as the title
  mutate(graphs = map2(data, id,
                       ~ggplot(data = .x, aes(x, y, dB)) +
                         geom_raster(aes(x = x, y = y, fill = dB)) +
                         geom_text(aes(label = dB, x = x, y = y), size = 8) +
                         coord_fixed(ratio = 1) +
                         scale_fill_gradientn(colours = viridis(47), limits = c(-1, 46),
                                              na.value="darkred") +
                         theme_bw() +
                         ggtitle(paste0("HFA ", id, " Visit ", visit, " Eye: ", eye))
                       )
         ) %>% 
  # pull is the pipe-able equivalent of .[['graphs']]
  pull(graphs) %>% 
  # Return the svg of graphs
  map(function(gr) svglite::xmlSVG(show(gr), standalone = TRUE)) -> 
  hfa_individual_graphs

dBdat %>%
  mutate(dB = as.numeric(dB)) %>%
  filter(device == "retinalogik" & id == "RL16") %>%
  group_by(id, visit, eye) %>%
  arrange(id, visit, eye) %>%
  # Use the group_by %>% nest pattern to group data by id
  nest() %>% 
  # Use map2 so the id can be used as the title
  mutate(graphs = map2(data, id,
                       ~ggplot(data = .x, aes(x, y, dB)) +
                         geom_raster(aes(x = x, y = y, fill = dB)) +
                         geom_text(aes(label = dB, x = x, y = y), size = 8) +
                         coord_fixed(ratio = 1) +
                         scale_fill_gradientn(colours = viridis(47), limits = c(-1, 46),
                                              na.value="darkred") +
                         theme_bw() +
                         ggtitle(paste0("Retinalogik ", id, " Visit ", visit, " Eye: ", eye))
                       )
         ) %>% 
  # pull is the pipe-able equivalent of .[['graphs']]
  pull(graphs) %>% 
  # Return the svg of graphs
  map(function(gr) svglite::xmlSVG(show(gr), standalone = TRUE)) -> 
  rt_individual_graphs

# carousels
hfa_plots <- slickR(hfa_individual_graphs, height = 350, width = "95%", padding = 0) +
  settings(slidesToShow = 2, slidesToScroll = 2)

rt_plots <- slickR(rt_individual_graphs, height = 350, width = "95%", padding = 0) +
  settings(slidesToShow = 2, slidesToScroll = 2)

rt_plots %synch% hfa_plots 
```

### RL17

```{r rl17, warning=FALSE}
# make sure coordinates are numeric, then flip OS to OD for plotting purposes
dBdat %>%
  mutate(dB = as.numeric(dB)) %>%
  filter(device == "hfa" & id == "RL17") %>%
  group_by(id, visit, eye) %>%
  arrange(id, visit, eye) %>%
  # Use the group_by %>% nest pattern to group data by id
  nest() %>% 
  # Use map2 so the id can be used as the title
  mutate(graphs = map2(data, id,
                       ~ggplot(data = .x, aes(x, y, dB)) +
                         geom_raster(aes(x = x, y = y, fill = dB)) +
                         geom_text(aes(label = dB, x = x, y = y), size = 8) +
                         coord_fixed(ratio = 1) +
                         scale_fill_gradientn(colours = viridis(47), limits = c(-1, 46),
                                              na.value="darkred") +
                         theme_bw() +
                         ggtitle(paste0("HFA ", id, " Visit ", visit, " Eye: ", eye))
                       )
         ) %>% 
  # pull is the pipe-able equivalent of .[['graphs']]
  pull(graphs) %>% 
  # Return the svg of graphs
  map(function(gr) svglite::xmlSVG(show(gr), standalone = TRUE)) -> 
  hfa_individual_graphs

dBdat %>%
  mutate(dB = as.numeric(dB)) %>%
  filter(device == "retinalogik" & id == "RL17") %>%
  group_by(id, visit, eye) %>%
  arrange(id, visit, eye) %>%
  # Use the group_by %>% nest pattern to group data by id
  nest() %>% 
  # Use map2 so the id can be used as the title
  mutate(graphs = map2(data, id,
                       ~ggplot(data = .x, aes(x, y, dB)) +
                         geom_raster(aes(x = x, y = y, fill = dB)) +
                         geom_text(aes(label = dB, x = x, y = y), size = 8) +
                         coord_fixed(ratio = 1) +
                         scale_fill_gradientn(colours = viridis(47), limits = c(-1, 46),
                                              na.value="darkred") +
                         theme_bw() +
                         ggtitle(paste0("Retinalogik ", id, " Visit ", visit, " Eye: ", eye))
                       )
         ) %>% 
  # pull is the pipe-able equivalent of .[['graphs']]
  pull(graphs) %>% 
  # Return the svg of graphs
  map(function(gr) svglite::xmlSVG(show(gr), standalone = TRUE)) -> 
  rt_individual_graphs

# carousels
hfa_plots <- slickR(hfa_individual_graphs, height = 350, width = "95%", padding = 0) +
  settings(slidesToShow = 2, slidesToScroll = 2)

rt_plots <- slickR(rt_individual_graphs, height = 350, width = "95%", padding = 0) +
  settings(slidesToShow = 2, slidesToScroll = 2)

rt_plots %synch% hfa_plots 
```

### RL18

```{r rl18, warning=FALSE}
# make sure coordinates are numeric, then flip OS to OD for plotting purposes
dBdat %>%
  mutate(dB = as.numeric(dB)) %>%
  filter(device == "hfa" & id == "RL18") %>%
  group_by(id, visit, eye) %>%
  arrange(id, visit, eye) %>%
  # Use the group_by %>% nest pattern to group data by id
  nest() %>% 
  # Use map2 so the id can be used as the title
  mutate(graphs = map2(data, id,
                       ~ggplot(data = .x, aes(x, y, dB)) +
                         geom_raster(aes(x = x, y = y, fill = dB)) +
                         geom_text(aes(label = dB, x = x, y = y), size = 8) +
                         coord_fixed(ratio = 1) +
                         scale_fill_gradientn(colours = viridis(47), limits = c(-1, 46),
                                              na.value="darkred") +
                         theme_bw() +
                         ggtitle(paste0("HFA ", id, " Visit ", visit, " Eye: ", eye))
                       )
         ) %>% 
  # pull is the pipe-able equivalent of .[['graphs']]
  pull(graphs) %>% 
  # Return the svg of graphs
  map(function(gr) svglite::xmlSVG(show(gr), standalone = TRUE)) -> 
  hfa_individual_graphs

dBdat %>%
  mutate(dB = as.numeric(dB)) %>%
  filter(device == "retinalogik" & id == "RL18") %>%
  group_by(id, visit, eye) %>%
  arrange(id, visit, eye) %>%
  # Use the group_by %>% nest pattern to group data by id
  nest() %>% 
  # Use map2 so the id can be used as the title
  mutate(graphs = map2(data, id,
                       ~ggplot(data = .x, aes(x, y, dB)) +
                         geom_raster(aes(x = x, y = y, fill = dB)) +
                         geom_text(aes(label = dB, x = x, y = y), size = 8) +
                         coord_fixed(ratio = 1) +
                         scale_fill_gradientn(colours = viridis(47), limits = c(-1, 46),
                                              na.value="darkred") +
                         theme_bw() +
                         ggtitle(paste0("Retinalogik ", id, " Visit ", visit, " Eye: ", eye))
                       )
         ) %>% 
  # pull is the pipe-able equivalent of .[['graphs']]
  pull(graphs) %>% 
  # Return the svg of graphs
  map(function(gr) svglite::xmlSVG(show(gr), standalone = TRUE)) -> 
  rt_individual_graphs

# carousels
hfa_plots <- slickR(hfa_individual_graphs, height = 350, width = "95%", padding = 0) +
  settings(slidesToShow = 2, slidesToScroll = 2)

rt_plots <- slickR(rt_individual_graphs, height = 350, width = "95%", padding = 0) +
  settings(slidesToShow = 2, slidesToScroll = 2)

rt_plots %synch% hfa_plots 
```

### RL19

```{r rl19, warning=FALSE}
# make sure coordinates are numeric, then flip OS to OD for plotting purposes
dBdat %>%
  mutate(dB = as.numeric(dB)) %>%
  filter(device == "hfa" & id == "RL19") %>%
  group_by(id, visit, eye) %>%
  arrange(id, visit, eye) %>%
  # Use the group_by %>% nest pattern to group data by id
  nest() %>% 
  # Use map2 so the id can be used as the title
  mutate(graphs = map2(data, id,
                       ~ggplot(data = .x, aes(x, y, dB)) +
                         geom_raster(aes(x = x, y = y, fill = dB)) +
                         geom_text(aes(label = dB, x = x, y = y), size = 8) +
                         coord_fixed(ratio = 1) +
                         scale_fill_gradientn(colours = viridis(47), limits = c(-1, 46),
                                              na.value="darkred") +
                         theme_bw() +
                         ggtitle(paste0("HFA ", id, " Visit ", visit, " Eye: ", eye))
                       )
         ) %>% 
  # pull is the pipe-able equivalent of .[['graphs']]
  pull(graphs) %>% 
  # Return the svg of graphs
  map(function(gr) svglite::xmlSVG(show(gr), standalone = TRUE)) -> 
  hfa_individual_graphs

dBdat %>%
  mutate(dB = as.numeric(dB)) %>%
  filter(device == "retinalogik" & id == "RL19") %>%
  group_by(id, visit, eye) %>%
  arrange(id, visit, eye) %>%
  # Use the group_by %>% nest pattern to group data by id
  nest() %>% 
  # Use map2 so the id can be used as the title
  mutate(graphs = map2(data, id,
                       ~ggplot(data = .x, aes(x, y, dB)) +
                         geom_raster(aes(x = x, y = y, fill = dB)) +
                         geom_text(aes(label = dB, x = x, y = y), size = 8) +
                         coord_fixed(ratio = 1) +
                         scale_fill_gradientn(colours = viridis(47), limits = c(-1, 46),
                                              na.value="darkred") +
                         theme_bw() +
                         ggtitle(paste0("Retinalogik ", id, " Visit ", visit, " Eye: ", eye))
                       )
         ) %>% 
  # pull is the pipe-able equivalent of .[['graphs']]
  pull(graphs) %>% 
  # Return the svg of graphs
  map(function(gr) svglite::xmlSVG(show(gr), standalone = TRUE)) -> 
  rt_individual_graphs

# carousels
hfa_plots <- slickR(hfa_individual_graphs, height = 350, width = "95%", padding = 0) +
  settings(slidesToShow = 2, slidesToScroll = 2)

rt_plots <- slickR(rt_individual_graphs, height = 350, width = "95%", padding = 0) +
  settings(slidesToShow = 2, slidesToScroll = 2)

rt_plots %synch% hfa_plots 
```

### RL20

```{r rl20, warning=FALSE}
# make sure coordinates are numeric, then flip OS to OD for plotting purposes
dBdat %>%
  mutate(dB = as.numeric(dB)) %>%
  filter(device == "hfa" & id == "RL20") %>%
  group_by(id, visit, eye) %>%
  arrange(id, visit, eye) %>%
  # Use the group_by %>% nest pattern to group data by id
  nest() %>% 
  # Use map2 so the id can be used as the title
  mutate(graphs = map2(data, id,
                       ~ggplot(data = .x, aes(x, y, dB)) +
                         geom_raster(aes(x = x, y = y, fill = dB)) +
                         geom_text(aes(label = dB, x = x, y = y), size = 8) +
                         coord_fixed(ratio = 1) +
                         scale_fill_gradientn(colours = viridis(47), limits = c(-1, 46),
                                              na.value="darkred") +
                         theme_bw() +
                         ggtitle(paste0("HFA ", id, " Visit ", visit, " Eye: ", eye))
                       )
         ) %>% 
  # pull is the pipe-able equivalent of .[['graphs']]
  pull(graphs) %>% 
  # Return the svg of graphs
  map(function(gr) svglite::xmlSVG(show(gr), standalone = TRUE)) -> 
  hfa_individual_graphs

dBdat %>%
  mutate(dB = as.numeric(dB)) %>%
  filter(device == "retinalogik" & id == "RL20") %>%
  group_by(id, visit, eye) %>%
  arrange(id, visit, eye) %>%
  # Use the group_by %>% nest pattern to group data by id
  nest() %>% 
  # Use map2 so the id can be used as the title
  mutate(graphs = map2(data, id,
                       ~ggplot(data = .x, aes(x, y, dB)) +
                         geom_raster(aes(x = x, y = y, fill = dB)) +
                         geom_text(aes(label = dB, x = x, y = y), size = 8) +
                         coord_fixed(ratio = 1) +
                         scale_fill_gradientn(colours = viridis(47), limits = c(-1, 46),
                                              na.value="darkred") +
                         theme_bw() +
                         ggtitle(paste0("Retinalogik ", id, " Visit ", visit, " Eye: ", eye))
                       )
         ) %>% 
  # pull is the pipe-able equivalent of .[['graphs']]
  pull(graphs) %>% 
  # Return the svg of graphs
  map(function(gr) svglite::xmlSVG(show(gr), standalone = TRUE)) -> 
  rt_individual_graphs

# carousels
hfa_plots <- slickR(hfa_individual_graphs, height = 350, width = "95%", padding = 0) +
  settings(slidesToShow = 2, slidesToScroll = 2)

rt_plots <- slickR(rt_individual_graphs, height = 350, width = "95%", padding = 0) +
  settings(slidesToShow = 2, slidesToScroll = 2)

rt_plots %synch% hfa_plots 
```

### RL21

```{r rl21, warning=FALSE}
# make sure coordinates are numeric, then flip OS to OD for plotting purposes
dBdat %>%
  mutate(dB = as.numeric(dB)) %>%
  filter(device == "hfa" & id == "RL21") %>%
  group_by(id, visit, eye) %>%
  arrange(id, visit, eye) %>%
  # Use the group_by %>% nest pattern to group data by id
  nest() %>% 
  # Use map2 so the id can be used as the title
  mutate(graphs = map2(data, id,
                       ~ggplot(data = .x, aes(x, y, dB)) +
                         geom_raster(aes(x = x, y = y, fill = dB)) +
                         geom_text(aes(label = dB, x = x, y = y), size = 8) +
                         coord_fixed(ratio = 1) +
                         scale_fill_gradientn(colours = viridis(47), limits = c(-1, 46),
                                              na.value="darkred") +
                         theme_bw() +
                         ggtitle(paste0("HFA ", id, " Visit ", visit, " Eye: ", eye))
                       )
         ) %>% 
  # pull is the pipe-able equivalent of .[['graphs']]
  pull(graphs) %>% 
  # Return the svg of graphs
  map(function(gr) svglite::xmlSVG(show(gr), standalone = TRUE)) -> 
  hfa_individual_graphs

dBdat %>%
  mutate(dB = as.numeric(dB)) %>%
  filter(device == "retinalogik" & id == "RL21") %>%
  group_by(id, visit, eye) %>%
  arrange(id, visit, eye) %>%
  # Use the group_by %>% nest pattern to group data by id
  nest() %>% 
  # Use map2 so the id can be used as the title
  mutate(graphs = map2(data, id,
                       ~ggplot(data = .x, aes(x, y, dB)) +
                         geom_raster(aes(x = x, y = y, fill = dB)) +
                         geom_text(aes(label = dB, x = x, y = y), size = 8) +
                         coord_fixed(ratio = 1) +
                         scale_fill_gradientn(colours = viridis(47), limits = c(-1, 46),
                                              na.value="darkred") +
                         theme_bw() +
                         ggtitle(paste0("Retinalogik ", id, " Visit ", visit, " Eye: ", eye))
                       )
         ) %>% 
  # pull is the pipe-able equivalent of .[['graphs']]
  pull(graphs) %>% 
  # Return the svg of graphs
  map(function(gr) svglite::xmlSVG(show(gr), standalone = TRUE)) -> 
  rt_individual_graphs

# carousels
hfa_plots <- slickR(hfa_individual_graphs, height = 350, width = "95%", padding = 0) +
  settings(slidesToShow = 2, slidesToScroll = 2)

rt_plots <- slickR(rt_individual_graphs, height = 350, width = "95%", padding = 0) +
  settings(slidesToShow = 2, slidesToScroll = 2)

rt_plots %synch% hfa_plots 
```

### RL22

```{r rl22, warning=FALSE}
# make sure coordinates are numeric, then flip OS to OD for plotting purposes
dBdat %>%
  mutate(dB = as.numeric(dB)) %>%
  filter(device == "hfa" & id == "RL22") %>%
  group_by(id, visit, eye) %>%
  arrange(id, visit, eye) %>%
  # Use the group_by %>% nest pattern to group data by id
  nest() %>% 
  # Use map2 so the id can be used as the title
  mutate(graphs = map2(data, id,
                       ~ggplot(data = .x, aes(x, y, dB)) +
                         geom_raster(aes(x = x, y = y, fill = dB)) +
                         geom_text(aes(label = dB, x = x, y = y), size = 8) +
                         coord_fixed(ratio = 1) +
                         scale_fill_gradientn(colours = viridis(47), limits = c(-1, 46),
                                              na.value="darkred") +
                         theme_bw() +
                         ggtitle(paste0("HFA ", id, " Visit ", visit, " Eye: ", eye))
                       )
         ) %>% 
  # pull is the pipe-able equivalent of .[['graphs']]
  pull(graphs) %>% 
  # Return the svg of graphs
  map(function(gr) svglite::xmlSVG(show(gr), standalone = TRUE)) -> 
  hfa_individual_graphs

dBdat %>%
  mutate(dB = as.numeric(dB)) %>%
  filter(device == "retinalogik" & id == "RL22") %>%
  group_by(id, visit, eye) %>%
  arrange(id, visit, eye) %>%
  # Use the group_by %>% nest pattern to group data by id
  nest() %>% 
  # Use map2 so the id can be used as the title
  mutate(graphs = map2(data, id,
                       ~ggplot(data = .x, aes(x, y, dB)) +
                         geom_raster(aes(x = x, y = y, fill = dB)) +
                         geom_text(aes(label = dB, x = x, y = y), size = 8) +
                         coord_fixed(ratio = 1) +
                         scale_fill_gradientn(colours = viridis(47), limits = c(-1, 46),
                                              na.value="darkred") +
                         theme_bw() +
                         ggtitle(paste0("Retinalogik ", id, " Visit ", visit, " Eye: ", eye))
                       )
         ) %>% 
  # pull is the pipe-able equivalent of .[['graphs']]
  pull(graphs) %>% 
  # Return the svg of graphs
  map(function(gr) svglite::xmlSVG(show(gr), standalone = TRUE)) -> 
  rt_individual_graphs

# carousels
hfa_plots <- slickR(hfa_individual_graphs, height = 350, width = "95%", padding = 0) +
  settings(slidesToShow = 2, slidesToScroll = 2)

rt_plots <- slickR(rt_individual_graphs, height = 350, width = "95%", padding = 0) +
  settings(slidesToShow = 2, slidesToScroll = 2)

rt_plots %synch% hfa_plots 
```

### RL23

```{r rl23, warning=FALSE}
dBdat %>%
  mutate(dB = as.numeric(dB)) %>%
  filter(device == "hfa" & id == "RL23") %>%
  group_by(id, visit, eye) %>%
  arrange(id, visit, eye) %>%
  # Use the group_by %>% nest pattern to group data by id
  nest() %>% 
  # Use map2 so the id can be used as the title
  mutate(graphs = map2(data, id,
                       ~ggplot(data = .x, aes(x, y, dB)) +
                         geom_raster(aes(x = x, y = y, fill = dB)) +
                         geom_text(aes(label = dB, x = x, y = y), size = 8) +
                         coord_fixed(ratio = 1) +
                         scale_fill_gradientn(colours = viridis(47), limits = c(-1, 46),
                                              na.value="darkred") +
                         theme_bw() +
                         ggtitle(paste0("HFA ", id, " Visit ", visit, " Eye: ", eye))
                       )
         ) %>% 
  # pull is the pipe-able equivalent of .[['graphs']]
  pull(graphs) %>% 
  # Return the svg of graphs
  map(function(gr) svglite::xmlSVG(show(gr), standalone = TRUE)) -> 
  hfa_individual_graphs

dBdat %>%
  mutate(dB = as.numeric(dB)) %>%
  filter(device == "retinalogik" & id == "RL23") %>%
  group_by(id, visit, eye) %>%
  arrange(id, visit, eye) %>%
  # Use the group_by %>% nest pattern to group data by id
  nest() %>% 
  # Use map2 so the id can be used as the title
  mutate(graphs = map2(data, id,
                       ~ggplot(data = .x, aes(x, y, dB)) +
                         geom_raster(aes(x = x, y = y, fill = dB)) +
                         geom_text(aes(label = dB, x = x, y = y), size = 8) +
                         coord_fixed(ratio = 1) +
                         scale_fill_gradientn(colours = viridis(47), limits = c(-1, 46),
                                              na.value="darkred") +
                         theme_bw() +
                         ggtitle(paste0("Retinalogik ", id, " Visit ", visit, " Eye: ", eye))
                       )
         ) %>% 
  # pull is the pipe-able equivalent of .[['graphs']]
  pull(graphs) %>% 
  # Return the svg of graphs
  map(function(gr) svglite::xmlSVG(show(gr), standalone = TRUE)) -> 
  rt_individual_graphs

# carousels
hfa_plots <- slickR(hfa_individual_graphs, height = 350, width = "95%", padding = 0) +
  settings(slidesToShow = 2, slidesToScroll = 2)

rt_plots <- slickR(rt_individual_graphs, height = 350, width = "95%", padding = 0) +
  settings(slidesToShow = 2, slidesToScroll = 2)

rt_plots %synch% hfa_plots 

```

### RL24

```{r rl24, warning=FALSE}
dBdat %>%
  mutate(dB = as.numeric(dB)) %>%
  filter(device == "hfa" & id == "RL24") %>%
  group_by(id, visit, eye) %>%
  arrange(id, visit, eye) %>%
  # Use the group_by %>% nest pattern to group data by id
  nest() %>% 
  # Use map2 so the id can be used as the title
  mutate(graphs = map2(data, id,
                       ~ggplot(data = .x, aes(x, y, dB)) +
                         geom_raster(aes(x = x, y = y, fill = dB)) +
                         geom_text(aes(label = dB, x = x, y = y), size = 8) +
                         coord_fixed(ratio = 1) +
                         scale_fill_gradientn(colours = viridis(47), limits = c(-1, 46),
                                              na.value="darkred") +
                         theme_bw() +
                         ggtitle(paste0("HFA ", id, " Visit ", visit, " Eye: ", eye))
                       )
         ) %>% 
  # pull is the pipe-able equivalent of .[['graphs']]
  pull(graphs) %>% 
  # Return the svg of graphs
  map(function(gr) svglite::xmlSVG(show(gr), standalone = TRUE)) -> 
  hfa_individual_graphs

dBdat %>%
  mutate(dB = as.numeric(dB)) %>%
  filter(device == "retinalogik" & id == "RL24") %>%
  group_by(id, visit, eye) %>%
  arrange(id, visit, eye) %>%
  # Use the group_by %>% nest pattern to group data by id
  nest() %>% 
  # Use map2 so the id can be used as the title
  mutate(graphs = map2(data, id,
                       ~ggplot(data = .x, aes(x, y, dB)) +
                         geom_raster(aes(x = x, y = y, fill = dB)) +
                         geom_text(aes(label = dB, x = x, y = y), size = 8) +
                         coord_fixed(ratio = 1) +
                         scale_fill_gradientn(colours = viridis(47), limits = c(-1, 46),
                                              na.value="darkred") +
                         theme_bw() +
                         ggtitle(paste0("Retinalogik ", id, " Visit ", visit, " Eye: ", eye))
                       )
         ) %>% 
  # pull is the pipe-able equivalent of .[['graphs']]
  pull(graphs) %>% 
  # Return the svg of graphs
  map(function(gr) svglite::xmlSVG(show(gr), standalone = TRUE)) -> 
  rt_individual_graphs

# carousels
hfa_plots <- slickR(hfa_individual_graphs, height = 350, width = "95%", padding = 0) +
  settings(slidesToShow = 2, slidesToScroll = 2)

rt_plots <- slickR(rt_individual_graphs, height = 350, width = "95%", padding = 0) +
  settings(slidesToShow = 2, slidesToScroll = 2)

rt_plots %synch% hfa_plots 

```

### RL25

```{r rl25, warning=FALSE}
dBdat %>%
  mutate(dB = as.numeric(dB)) %>%
  filter(device == "hfa" & id == "RL25") %>%
  group_by(id, visit, eye) %>%
  arrange(id, visit, eye) %>%
  # Use the group_by %>% nest pattern to group data by id
  nest() %>% 
  # Use map2 so the id can be used as the title
  mutate(graphs = map2(data, id,
                       ~ggplot(data = .x, aes(x, y, dB)) +
                         geom_raster(aes(x = x, y = y, fill = dB)) +
                         geom_text(aes(label = dB, x = x, y = y), size = 8) +
                         coord_fixed(ratio = 1) +
                         scale_fill_gradientn(colours = viridis(47), limits = c(-1, 46),
                                              na.value="darkred") +
                         theme_bw() +
                         ggtitle(paste0("HFA ", id, " Visit ", visit, " Eye: ", eye))
                       )
         ) %>% 
  # pull is the pipe-able equivalent of .[['graphs']]
  pull(graphs) %>% 
  # Return the svg of graphs
  map(function(gr) svglite::xmlSVG(show(gr), standalone = TRUE)) -> 
  hfa_individual_graphs

dBdat %>%
  mutate(dB = as.numeric(dB)) %>%
  filter(device == "retinalogik" & id == "RL25") %>%
  group_by(id, visit, eye) %>%
  arrange(id, visit, eye) %>%
  # Use the group_by %>% nest pattern to group data by id
  nest() %>% 
  # Use map2 so the id can be used as the title
  mutate(graphs = map2(data, id,
                       ~ggplot(data = .x, aes(x, y, dB)) +
                         geom_raster(aes(x = x, y = y, fill = dB)) +
                         geom_text(aes(label = dB, x = x, y = y), size = 8) +
                         coord_fixed(ratio = 1) +
                         scale_fill_gradientn(colours = viridis(47), limits = c(-1, 46),
                                              na.value="darkred") +
                         theme_bw() +
                         ggtitle(paste0("Retinalogik ", id, " Visit ", visit, " Eye: ", eye))
                       )
         ) %>% 
  # pull is the pipe-able equivalent of .[['graphs']]
  pull(graphs) %>% 
  # Return the svg of graphs
  map(function(gr) svglite::xmlSVG(show(gr), standalone = TRUE)) -> 
  rt_individual_graphs

# carousels
hfa_plots <- slickR(hfa_individual_graphs, height = 350, width = "95%", padding = 0) +
  settings(slidesToShow = 2, slidesToScroll = 2)

rt_plots <- slickR(rt_individual_graphs, height = 350, width = "95%", padding = 0) +
  settings(slidesToShow = 2, slidesToScroll = 2)

rt_plots %synch% hfa_plots 

```

### RL26

```{r rl26, warning=FALSE}
dBdat %>%
  mutate(dB = as.numeric(dB)) %>%
  filter(device == "hfa" & id == "RL26") %>%
  group_by(id, visit, eye) %>%
  arrange(id, visit, eye) %>%
  # Use the group_by %>% nest pattern to group data by id
  nest() %>% 
  # Use map2 so the id can be used as the title
  mutate(graphs = map2(data, id,
                       ~ggplot(data = .x, aes(x, y, dB)) +
                         geom_raster(aes(x = x, y = y, fill = dB)) +
                         geom_text(aes(label = dB, x = x, y = y), size = 8) +
                         coord_fixed(ratio = 1) +
                         scale_fill_gradientn(colours = viridis(47), limits = c(-1, 46),
                                              na.value="darkred") +
                         theme_bw() +
                         ggtitle(paste0("HFA ", id, " Visit ", visit, " Eye: ", eye))
                       )
         ) %>% 
  # pull is the pipe-able equivalent of .[['graphs']]
  pull(graphs) %>% 
  # Return the svg of graphs
  map(function(gr) svglite::xmlSVG(show(gr), standalone = TRUE)) -> 
  hfa_individual_graphs

dBdat %>%
  mutate(dB = as.numeric(dB)) %>%
  filter(device == "retinalogik" & id == "RL26") %>%
  group_by(id, visit, eye) %>%
  arrange(id, visit, eye) %>%
  # Use the group_by %>% nest pattern to group data by id
  nest() %>% 
  # Use map2 so the id can be used as the title
  mutate(graphs = map2(data, id,
                       ~ggplot(data = .x, aes(x, y, dB)) +
                         geom_raster(aes(x = x, y = y, fill = dB)) +
                         geom_text(aes(label = dB, x = x, y = y), size = 8) +
                         coord_fixed(ratio = 1) +
                         scale_fill_gradientn(colours = viridis(47), limits = c(-1, 46),
                                              na.value="darkred") +
                         theme_bw() +
                         ggtitle(paste0("Retinalogik ", id, " Visit ", visit, " Eye: ", eye))
                       )
         ) %>% 
  # pull is the pipe-able equivalent of .[['graphs']]
  pull(graphs) %>% 
  # Return the svg of graphs
  map(function(gr) svglite::xmlSVG(show(gr), standalone = TRUE)) -> 
  rt_individual_graphs

# carousels
hfa_plots <- slickR(hfa_individual_graphs, height = 350, width = "95%", padding = 0) +
  settings(slidesToShow = 2, slidesToScroll = 2)

rt_plots <- slickR(rt_individual_graphs, height = 350, width = "95%", padding = 0) +
  settings(slidesToShow = 2, slidesToScroll = 2)

rt_plots %synch% hfa_plots 

```

### RL27

```{r rl27, warning=FALSE}
dBdat %>%
  mutate(dB = as.numeric(dB)) %>%
  filter(device == "hfa" & id == "RL27") %>%
  group_by(id, visit, eye) %>%
  arrange(id, visit, eye) %>%
  # Use the group_by %>% nest pattern to group data by id
  nest() %>% 
  # Use map2 so the id can be used as the title
  mutate(graphs = map2(data, id,
                       ~ggplot(data = .x, aes(x, y, dB)) +
                         geom_raster(aes(x = x, y = y, fill = dB)) +
                         geom_text(aes(label = dB, x = x, y = y), size = 8) +
                         coord_fixed(ratio = 1) +
                         scale_fill_gradientn(colours = viridis(47), limits = c(-1, 46),
                                              na.value="darkred") +
                         theme_bw() +
                         ggtitle(paste0("HFA ", id, " Visit ", visit, " Eye: ", eye))
                       )
         ) %>% 
  # pull is the pipe-able equivalent of .[['graphs']]
  pull(graphs) %>% 
  # Return the svg of graphs
  map(function(gr) svglite::xmlSVG(show(gr), standalone = TRUE)) -> 
  hfa_individual_graphs

dBdat %>%
  mutate(dB = as.numeric(dB)) %>%
  filter(device == "retinalogik" & id == "RL27") %>%
  group_by(id, visit, eye) %>%
  arrange(id, visit, eye) %>%
  # Use the group_by %>% nest pattern to group data by id
  nest() %>% 
  # Use map2 so the id can be used as the title
  mutate(graphs = map2(data, id,
                       ~ggplot(data = .x, aes(x, y, dB)) +
                         geom_raster(aes(x = x, y = y, fill = dB)) +
                         geom_text(aes(label = dB, x = x, y = y), size = 8) +
                         coord_fixed(ratio = 1) +
                         scale_fill_gradientn(colours = viridis(47), limits = c(-1, 46),
                                              na.value="darkred") +
                         theme_bw() +
                         ggtitle(paste0("Retinalogik ", id, " Visit ", visit, " Eye: ", eye))
                       )
         ) %>% 
  # pull is the pipe-able equivalent of .[['graphs']]
  pull(graphs) %>% 
  # Return the svg of graphs
  map(function(gr) svglite::xmlSVG(show(gr), standalone = TRUE)) -> 
  rt_individual_graphs

# carousels
hfa_plots <- slickR(hfa_individual_graphs, height = 350, width = "95%", padding = 0) +
  settings(slidesToShow = 2, slidesToScroll = 2)

rt_plots <- slickR(rt_individual_graphs, height = 350, width = "95%", padding = 0) +
  settings(slidesToShow = 2, slidesToScroll = 2)

rt_plots %synch% hfa_plots 

```

### RL28

```{r rl28, warning=FALSE}
dBdat %>%
  mutate(dB = as.numeric(dB)) %>%
  filter(device == "hfa" & id == "RL28") %>%
  group_by(id, visit, eye) %>%
  arrange(id, visit, eye) %>%
  # Use the group_by %>% nest pattern to group data by id
  nest() %>% 
  # Use map2 so the id can be used as the title
  mutate(graphs = map2(data, id,
                       ~ggplot(data = .x, aes(x, y, dB)) +
                         geom_raster(aes(x = x, y = y, fill = dB)) +
                         geom_text(aes(label = dB, x = x, y = y), size = 8) +
                         coord_fixed(ratio = 1) +
                         scale_fill_gradientn(colours = viridis(47), limits = c(-1, 46),
                                              na.value="darkred") +
                         theme_bw() +
                         ggtitle(paste0("HFA ", id, " Visit ", visit, " Eye: ", eye))
                       )
         ) %>% 
  # pull is the pipe-able equivalent of .[['graphs']]
  pull(graphs) %>% 
  # Return the svg of graphs
  map(function(gr) svglite::xmlSVG(show(gr), standalone = TRUE)) -> 
  hfa_individual_graphs

dBdat %>%
  mutate(dB = as.numeric(dB)) %>%
  filter(device == "retinalogik" & id == "RL28") %>%
  group_by(id, visit, eye) %>%
  arrange(id, visit, eye) %>%
  # Use the group_by %>% nest pattern to group data by id
  nest() %>% 
  # Use map2 so the id can be used as the title
  mutate(graphs = map2(data, id,
                       ~ggplot(data = .x, aes(x, y, dB)) +
                         geom_raster(aes(x = x, y = y, fill = dB)) +
                         geom_text(aes(label = dB, x = x, y = y), size = 8) +
                         coord_fixed(ratio = 1) +
                         scale_fill_gradientn(colours = viridis(47), limits = c(-1, 46),
                                              na.value="darkred") +
                         theme_bw() +
                         ggtitle(paste0("Retinalogik ", id, " Visit ", visit, " Eye: ", eye))
                       )
         ) %>% 
  # pull is the pipe-able equivalent of .[['graphs']]
  pull(graphs) %>% 
  # Return the svg of graphs
  map(function(gr) svglite::xmlSVG(show(gr), standalone = TRUE)) -> 
  rt_individual_graphs

# carousels
hfa_plots <- slickR(hfa_individual_graphs, height = 350, width = "95%", padding = 0) +
  settings(slidesToShow = 2, slidesToScroll = 2)

rt_plots <- slickR(rt_individual_graphs, height = 350, width = "95%", padding = 0) +
  settings(slidesToShow = 2, slidesToScroll = 2)

rt_plots %synch% hfa_plots 

```

### RL29

```{r rl29, warning=FALSE}
dBdat %>%
  mutate(dB = as.numeric(dB)) %>%
  filter(device == "hfa" & id == "RL29") %>%
  group_by(id, visit, eye) %>%
  arrange(id, visit, eye) %>%
  # Use the group_by %>% nest pattern to group data by id
  nest() %>% 
  # Use map2 so the id can be used as the title
  mutate(graphs = map2(data, id,
                       ~ggplot(data = .x, aes(x, y, dB)) +
                         geom_raster(aes(x = x, y = y, fill = dB)) +
                         geom_text(aes(label = dB, x = x, y = y), size = 8) +
                         coord_fixed(ratio = 1) +
                         scale_fill_gradientn(colours = viridis(47), limits = c(-1, 46),
                                              na.value="darkred") +
                         theme_bw() +
                         ggtitle(paste0("HFA ", id, " Visit ", visit, " Eye: ", eye))
                       )
         ) %>% 
  # pull is the pipe-able equivalent of .[['graphs']]
  pull(graphs) %>% 
  # Return the svg of graphs
  map(function(gr) svglite::xmlSVG(show(gr), standalone = TRUE)) -> 
  hfa_individual_graphs

dBdat %>%
  mutate(dB = as.numeric(dB)) %>%
  filter(device == "retinalogik" & id == "RL29") %>%
  group_by(id, visit, eye) %>%
  arrange(id, visit, eye) %>%
  # Use the group_by %>% nest pattern to group data by id
  nest() %>% 
  # Use map2 so the id can be used as the title
  mutate(graphs = map2(data, id,
                       ~ggplot(data = .x, aes(x, y, dB)) +
                         geom_raster(aes(x = x, y = y, fill = dB)) +
                         geom_text(aes(label = dB, x = x, y = y), size = 8) +
                         coord_fixed(ratio = 1) +
                         scale_fill_gradientn(colours = viridis(47), limits = c(-1, 46),
                                              na.value="darkred") +
                         theme_bw() +
                         ggtitle(paste0("Retinalogik ", id, " Visit ", visit, " Eye: ", eye))
                       )
         ) %>% 
  # pull is the pipe-able equivalent of .[['graphs']]
  pull(graphs) %>% 
  # Return the svg of graphs
  map(function(gr) svglite::xmlSVG(show(gr), standalone = TRUE)) -> 
  rt_individual_graphs

# carousels
hfa_plots <- slickR(hfa_individual_graphs, height = 350, width = "95%", padding = 0) +
  settings(slidesToShow = 2, slidesToScroll = 2)

rt_plots <- slickR(rt_individual_graphs, height = 350, width = "95%", padding = 0) +
  settings(slidesToShow = 2, slidesToScroll = 2)

rt_plots %synch% hfa_plots 

```

### RL30

```{r rl30, warning=FALSE}
dBdat %>%
  mutate(dB = as.numeric(dB)) %>%
  filter(device == "hfa" & id == "RL30") %>%
  group_by(id, visit, eye) %>%
  arrange(id, visit, eye) %>%
  # Use the group_by %>% nest pattern to group data by id
  nest() %>% 
  # Use map2 so the id can be used as the title
  mutate(graphs = map2(data, id,
                       ~ggplot(data = .x, aes(x, y, dB)) +
                         geom_raster(aes(x = x, y = y, fill = dB)) +
                         geom_text(aes(label = dB, x = x, y = y), size = 8) +
                         coord_fixed(ratio = 1) +
                         scale_fill_gradientn(colours = viridis(47), limits = c(-1, 46),
                                              na.value="darkred") +
                         theme_bw() +
                         ggtitle(paste0("HFA ", id, " Visit ", visit, " Eye: ", eye))
                       )
         ) %>% 
  # pull is the pipe-able equivalent of .[['graphs']]
  pull(graphs) %>% 
  # Return the svg of graphs
  map(function(gr) svglite::xmlSVG(show(gr), standalone = TRUE)) -> 
  hfa_individual_graphs

dBdat %>%
  mutate(dB = as.numeric(dB)) %>%
  filter(device == "retinalogik" & id == "RL30") %>%
  group_by(id, visit, eye) %>%
  arrange(id, visit, eye) %>%
  # Use the group_by %>% nest pattern to group data by id
  nest() %>% 
  # Use map2 so the id can be used as the title
  mutate(graphs = map2(data, id,
                       ~ggplot(data = .x, aes(x, y, dB)) +
                         geom_raster(aes(x = x, y = y, fill = dB)) +
                         geom_text(aes(label = dB, x = x, y = y), size = 8) +
                         coord_fixed(ratio = 1) +
                         scale_fill_gradientn(colours = viridis(47), limits = c(-1, 46),
                                              na.value="darkred") +
                         theme_bw() +
                         ggtitle(paste0("Retinalogik ", id, " Visit ", visit, " Eye: ", eye))
                       )
         ) %>% 
  # pull is the pipe-able equivalent of .[['graphs']]
  pull(graphs) %>% 
  # Return the svg of graphs
  map(function(gr) svglite::xmlSVG(show(gr), standalone = TRUE)) -> 
  rt_individual_graphs

# carousels
hfa_plots <- slickR(hfa_individual_graphs, height = 350, width = "95%", padding = 0) +
  settings(slidesToShow = 2, slidesToScroll = 2)

rt_plots <- slickR(rt_individual_graphs, height = 350, width = "95%", padding = 0) +
  settings(slidesToShow = 2, slidesToScroll = 2)

rt_plots %synch% hfa_plots 

```

### RL31

```{r rl31, warning=FALSE}
dBdat %>%
  mutate(dB = as.numeric(dB)) %>%
  filter(device == "hfa" & id == "RL31") %>%
  group_by(id, visit, eye) %>%
  arrange(id, visit, eye) %>%
  # Use the group_by %>% nest pattern to group data by id
  nest() %>% 
  # Use map2 so the id can be used as the title
  mutate(graphs = map2(data, id,
                       ~ggplot(data = .x, aes(x, y, dB)) +
                         geom_raster(aes(x = x, y = y, fill = dB)) +
                         geom_text(aes(label = dB, x = x, y = y), size = 8) +
                         coord_fixed(ratio = 1) +
                         scale_fill_gradientn(colours = viridis(47), limits = c(-1, 46),
                                              na.value="darkred") +
                         theme_bw() +
                         ggtitle(paste0("HFA ", id, " Visit ", visit, " Eye: ", eye))
                       )
         ) %>% 
  # pull is the pipe-able equivalent of .[['graphs']]
  pull(graphs) %>% 
  # Return the svg of graphs
  map(function(gr) svglite::xmlSVG(show(gr), standalone = TRUE)) -> 
  hfa_individual_graphs

dBdat %>%
  mutate(dB = as.numeric(dB)) %>%
  filter(device == "retinalogik" & id == "RL31") %>%
  group_by(id, visit, eye) %>%
  arrange(id, visit, eye) %>%
  # Use the group_by %>% nest pattern to group data by id
  nest() %>% 
  # Use map2 so the id can be used as the title
  mutate(graphs = map2(data, id,
                       ~ggplot(data = .x, aes(x, y, dB)) +
                         geom_raster(aes(x = x, y = y, fill = dB)) +
                         geom_text(aes(label = dB, x = x, y = y), size = 8) +
                         coord_fixed(ratio = 1) +
                         scale_fill_gradientn(colours = viridis(47), limits = c(-1, 46),
                                              na.value="darkred") +
                         theme_bw() +
                         ggtitle(paste0("Retinalogik ", id, " Visit ", visit, " Eye: ", eye))
                       )
         ) %>% 
  # pull is the pipe-able equivalent of .[['graphs']]
  pull(graphs) %>% 
  # Return the svg of graphs
  map(function(gr) svglite::xmlSVG(show(gr), standalone = TRUE)) -> 
  rt_individual_graphs

# carousels
hfa_plots <- slickR(hfa_individual_graphs, height = 350, width = "95%", padding = 0) +
  settings(slidesToShow = 2, slidesToScroll = 2)

rt_plots <- slickR(rt_individual_graphs, height = 350, width = "95%", padding = 0) +
  settings(slidesToShow = 2, slidesToScroll = 2)

rt_plots %synch% hfa_plots 

```

### RL32

```{r rl32, warning=FALSE}
dBdat %>%
  mutate(dB = as.numeric(dB)) %>%
  filter(device == "hfa" & id == "RL32") %>%
  group_by(id, visit, eye) %>%
  arrange(id, visit, eye) %>%
  # Use the group_by %>% nest pattern to group data by id
  nest() %>% 
  # Use map2 so the id can be used as the title
  mutate(graphs = map2(data, id,
                       ~ggplot(data = .x, aes(x, y, dB)) +
                         geom_raster(aes(x = x, y = y, fill = dB)) +
                         geom_text(aes(label = dB, x = x, y = y), size = 8) +
                         coord_fixed(ratio = 1) +
                         scale_fill_gradientn(colours = viridis(47), limits = c(-1, 46),
                                              na.value="darkred") +
                         theme_bw() +
                         ggtitle(paste0("HFA ", id, " Visit ", visit, " Eye: ", eye))
                       )
         ) %>% 
  # pull is the pipe-able equivalent of .[['graphs']]
  pull(graphs) %>% 
  # Return the svg of graphs
  map(function(gr) svglite::xmlSVG(show(gr), standalone = TRUE)) -> 
  hfa_individual_graphs

dBdat %>%
  mutate(dB = as.numeric(dB)) %>%
  filter(device == "retinalogik" & id == "RL32") %>%
  group_by(id, visit, eye) %>%
  arrange(id, visit, eye) %>%
  # Use the group_by %>% nest pattern to group data by id
  nest() %>% 
  # Use map2 so the id can be used as the title
  mutate(graphs = map2(data, id,
                       ~ggplot(data = .x, aes(x, y, dB)) +
                         geom_raster(aes(x = x, y = y, fill = dB)) +
                         geom_text(aes(label = dB, x = x, y = y), size = 8) +
                         coord_fixed(ratio = 1) +
                         scale_fill_gradientn(colours = viridis(47), limits = c(-1, 46),
                                              na.value="darkred") +
                         theme_bw() +
                         ggtitle(paste0("Retinalogik ", id, " Visit ", visit, " Eye: ", eye))
                       )
         ) %>% 
  # pull is the pipe-able equivalent of .[['graphs']]
  pull(graphs) %>% 
  # Return the svg of graphs
  map(function(gr) svglite::xmlSVG(show(gr), standalone = TRUE)) -> 
  rt_individual_graphs

# carousels
hfa_plots <- slickR(hfa_individual_graphs, height = 350, width = "95%", padding = 0) +
  settings(slidesToShow = 2, slidesToScroll = 2)

rt_plots <- slickR(rt_individual_graphs, height = 350, width = "95%", padding = 0) +
  settings(slidesToShow = 2, slidesToScroll = 2)

rt_plots %synch% hfa_plots 

```

### RL33

```{r rl33, warning=FALSE}
dBdat %>%
  mutate(dB = as.numeric(dB)) %>%
  filter(device == "hfa" & id == "RL33") %>%
  group_by(id, visit, eye) %>%
  arrange(id, visit, eye) %>%
  # Use the group_by %>% nest pattern to group data by id
  nest() %>% 
  # Use map2 so the id can be used as the title
  mutate(graphs = map2(data, id,
                       ~ggplot(data = .x, aes(x, y, dB)) +
                         geom_raster(aes(x = x, y = y, fill = dB)) +
                         geom_text(aes(label = dB, x = x, y = y), size = 8) +
                         coord_fixed(ratio = 1) +
                         scale_fill_gradientn(colours = viridis(47), limits = c(-1, 46),
                                              na.value="darkred") +
                         theme_bw() +
                         ggtitle(paste0("HFA ", id, " Visit ", visit, " Eye: ", eye))
                       )
         ) %>% 
  # pull is the pipe-able equivalent of .[['graphs']]
  pull(graphs) %>% 
  # Return the svg of graphs
  map(function(gr) svglite::xmlSVG(show(gr), standalone = TRUE)) -> 
  hfa_individual_graphs

dBdat %>%
  mutate(dB = as.numeric(dB)) %>%
  filter(device == "retinalogik" & id == "RL33") %>%
  group_by(id, visit, eye) %>%
  arrange(id, visit, eye) %>%
  # Use the group_by %>% nest pattern to group data by id
  nest() %>% 
  # Use map2 so the id can be used as the title
  mutate(graphs = map2(data, id,
                       ~ggplot(data = .x, aes(x, y, dB)) +
                         geom_raster(aes(x = x, y = y, fill = dB)) +
                         geom_text(aes(label = dB, x = x, y = y), size = 8) +
                         coord_fixed(ratio = 1) +
                         scale_fill_gradientn(colours = viridis(47), limits = c(-1, 46),
                                              na.value="darkred") +
                         theme_bw() +
                         ggtitle(paste0("Retinalogik ", id, " Visit ", visit, " Eye: ", eye))
                       )
         ) %>% 
  # pull is the pipe-able equivalent of .[['graphs']]
  pull(graphs) %>% 
  # Return the svg of graphs
  map(function(gr) svglite::xmlSVG(show(gr), standalone = TRUE)) -> 
  rt_individual_graphs

# carousels
hfa_plots <- slickR(hfa_individual_graphs, height = 350, width = "95%", padding = 0) +
  settings(slidesToShow = 2, slidesToScroll = 2)

rt_plots <- slickR(rt_individual_graphs, height = 350, width = "95%", padding = 0) +
  settings(slidesToShow = 2, slidesToScroll = 2)

rt_plots %synch% hfa_plots 

```

### RL34

```{r rl34, warning=FALSE}
dBdat %>%
  mutate(dB = as.numeric(dB)) %>%
  filter(device == "hfa" & id == "RL34") %>%
  group_by(id, visit, eye) %>%
  arrange(id, visit, eye) %>%
  # Use the group_by %>% nest pattern to group data by id
  nest() %>% 
  # Use map2 so the id can be used as the title
  mutate(graphs = map2(data, id,
                       ~ggplot(data = .x, aes(x, y, dB)) +
                         geom_raster(aes(x = x, y = y, fill = dB)) +
                         geom_text(aes(label = dB, x = x, y = y), size = 8) +
                         coord_fixed(ratio = 1) +
                         scale_fill_gradientn(colours = viridis(47), limits = c(-1, 46),
                                              na.value="darkred") +
                         theme_bw() +
                         ggtitle(paste0("HFA ", id, " Visit ", visit, " Eye: ", eye))
                       )
         ) %>% 
  # pull is the pipe-able equivalent of .[['graphs']]
  pull(graphs) %>% 
  # Return the svg of graphs
  map(function(gr) svglite::xmlSVG(show(gr), standalone = TRUE)) -> 
  hfa_individual_graphs

dBdat %>%
  mutate(dB = as.numeric(dB)) %>%
  filter(device == "retinalogik" & id == "RL34") %>%
  group_by(id, visit, eye) %>%
  arrange(id, visit, eye) %>%
  # Use the group_by %>% nest pattern to group data by id
  nest() %>% 
  # Use map2 so the id can be used as the title
  mutate(graphs = map2(data, id,
                       ~ggplot(data = .x, aes(x, y, dB)) +
                         geom_raster(aes(x = x, y = y, fill = dB)) +
                         geom_text(aes(label = dB, x = x, y = y), size = 8) +
                         coord_fixed(ratio = 1) +
                         scale_fill_gradientn(colours = viridis(47), limits = c(-1, 46),
                                              na.value="darkred") +
                         theme_bw() +
                         ggtitle(paste0("Retinalogik ", id, " Visit ", visit, " Eye: ", eye))
                       )
         ) %>% 
  # pull is the pipe-able equivalent of .[['graphs']]
  pull(graphs) %>% 
  # Return the svg of graphs
  map(function(gr) svglite::xmlSVG(show(gr), standalone = TRUE)) -> 
  rt_individual_graphs

# carousels
hfa_plots <- slickR(hfa_individual_graphs, height = 350, width = "95%", padding = 0) +
  settings(slidesToShow = 2, slidesToScroll = 2)

rt_plots <- slickR(rt_individual_graphs, height = 350, width = "95%", padding = 0) +
  settings(slidesToShow = 2, slidesToScroll = 2)

rt_plots %synch% hfa_plots 

```

### RL35

```{r rl35, warning=FALSE}
dBdat %>%
  mutate(dB = as.numeric(dB)) %>%
  filter(device == "hfa" & id == "RL35") %>%
  group_by(id, visit, eye) %>%
  arrange(id, visit, eye) %>%
  # Use the group_by %>% nest pattern to group data by id
  nest() %>% 
  # Use map2 so the id can be used as the title
  mutate(graphs = map2(data, id,
                       ~ggplot(data = .x, aes(x, y, dB)) +
                         geom_raster(aes(x = x, y = y, fill = dB)) +
                         geom_text(aes(label = dB, x = x, y = y), size = 8) +
                         coord_fixed(ratio = 1) +
                         scale_fill_gradientn(colours = viridis(47), limits = c(-1, 46),
                                              na.value="darkred") +
                         theme_bw() +
                         ggtitle(paste0("HFA ", id, " Visit ", visit, " Eye: ", eye))
                       )
         ) %>% 
  # pull is the pipe-able equivalent of .[['graphs']]
  pull(graphs) %>% 
  # Return the svg of graphs
  map(function(gr) svglite::xmlSVG(show(gr), standalone = TRUE)) -> 
  hfa_individual_graphs

dBdat %>%
  mutate(dB = as.numeric(dB)) %>%
  filter(device == "retinalogik" & id == "RL35") %>%
  group_by(id, visit, eye) %>%
  arrange(id, visit, eye) %>%
  # Use the group_by %>% nest pattern to group data by id
  nest() %>% 
  # Use map2 so the id can be used as the title
  mutate(graphs = map2(data, id,
                       ~ggplot(data = .x, aes(x, y, dB)) +
                         geom_raster(aes(x = x, y = y, fill = dB)) +
                         geom_text(aes(label = dB, x = x, y = y), size = 8) +
                         coord_fixed(ratio = 1) +
                         scale_fill_gradientn(colours = viridis(47), limits = c(-1, 46),
                                              na.value="darkred") +
                         theme_bw() +
                         ggtitle(paste0("Retinalogik ", id, " Visit ", visit, " Eye: ", eye))
                       )
         ) %>% 
  # pull is the pipe-able equivalent of .[['graphs']]
  pull(graphs) %>% 
  # Return the svg of graphs
  map(function(gr) svglite::xmlSVG(show(gr), standalone = TRUE)) -> 
  rt_individual_graphs

# carousels
hfa_plots <- slickR(hfa_individual_graphs, height = 350, width = "95%", padding = 0) +
  settings(slidesToShow = 2, slidesToScroll = 2)

rt_plots <- slickR(rt_individual_graphs, height = 350, width = "95%", padding = 0) +
  settings(slidesToShow = 2, slidesToScroll = 2)

rt_plots %synch% hfa_plots 

```

## Global sensitivity indices
Bland–Altman analysis for mean deviation comparing the latest reliable VF tests for the HFA and Retinalogik, where FL, FP, and FNs were below 30%. Only data from the eye with the worse MD were included. The red solid line indicates the mean bias, the black dashed lines indicate the 95% limits of agreement, the blue solid line indicates the regression line and the black solid line indicates y = 0.

```{r md, warning=FALSE}
library(BlandAltmanLeh)

dBdat_flipped <- dBdat %>%
  mutate(x = as.numeric(x), y = as.numeric(y)) %>%
  mutate(x = case_when(
    eye == "L" ~ x * -1,
    TRUE ~ as.numeric(x)
  ))

check_visit <- function(df, visit_number) {
  df %>%
    filter(visit == visit_number) %>%
    group_by(eye) %>%
    mutate(is_below_30 = flPerc < 30 & fpPerc < 30 & fnPerc < 30) %>%
    filter(all(is_below_30)) %>%
    ungroup()
}

unique_ids <- unique(dBdat_flipped$id)
all_processed_data <- tibble::tibble()

for (current_id in unique_ids) {
  current_data <- dBdat_flipped %>% filter(id == current_id)
  unique_visits <- current_data %>%
    pull(visit) %>%
    unique() %>%
    sort(decreasing = TRUE)

  processed_data_single_id <- tibble::tibble()
  qualifying_visit <- NA

  if (length(unique_visits) == 1) {
    visit_number <- unique_visits[1]
    temp_processed_data <- current_data %>%
      check_visit(visit_number)
    if (nrow(temp_processed_data) > 0) {
      processed_data_single_id <- temp_processed_data
      qualifying_visit <- visit_number
    }
  } else {
    for (visit_number in unique_visits) {
      temp_processed_data <- current_data %>%
        check_visit(visit_number)
      if (nrow(temp_processed_data) > 0) {
        processed_data_single_id <- temp_processed_data
        qualifying_visit <- visit_number
        break
      }
    }
  }

  if (nrow(processed_data_single_id) > 0) {
    all_processed_data <- bind_rows(all_processed_data, processed_data_single_id)
  }
}

print("Contents of all_processed_data:")
print(all_processed_data)

# Filter for patients with qualifying data for both devices
qualifying_patients <- all_processed_data %>%
  distinct(id, device) %>%
  group_by(id) %>%
  filter(n_distinct(device) == 2) %>%
  pull(id) %>%
  unique()

if (length(qualifying_patients) > 0) {
  processed_data_for_bland_altman <- all_processed_data %>%
    filter(id %in% qualifying_patients)

  if ("hfa" %in% unique(processed_data_for_bland_altman$device) &&
      "retinalogik" %in% unique(processed_data_for_bland_altman$device)) {

    min_md_eye_all <- processed_data_for_bland_altman %>%
      group_by(id, visit, device, eye) %>%
      summarise(min_md = min(md, na.rm=TRUE)) %>%
      group_by(id) %>%
      filter(min_md == min(min_md, na.rm=TRUE)) %>%
      distinct(id, eye, visit)

    df_wide_all <- dBdat_flipped %>%
      distinct(id, device, visit, eye, md) %>%
      right_join(min_md_eye_all) %>%
      pivot_wider(names_from = device, values_from = md)

    print("Contents of df_wide_all (for qualifying patients):")
    print(df_wide_all)

    if (nrow(df_wide_all) > 0 && all(c("hfa", "retinalogik") %in% names(df_wide_all))) {
      p <- bland.altman.plot(df_wide_all$hfa, df_wide_all$retinalogik,
                              graph.sys = "ggplot2")
      meanbias <- mean(df_wide_all$hfa, na.rm=T) - mean(df_wide_all$retinalogik, na.rm=T)

      print(p + geom_smooth(method = "lm", se = FALSE) +
              geom_point(position = "jitter") +
              geom_hline(yintercept = 0, color = "black") +
              geom_hline(yintercept = meanbias, color = "red", linetype = "solid", size = 2) +
              annotate("text", x = -6, y = 0.5, label=paste("Mean bias =", round(meanbias, 2), "dB")) +
              xlab("MD (dB)") +
              ylab("Difference in MD (HFA-Retinalogik)") +
              ggtitle("Bland-Altman Plot (Lowest MD Eye)",
                      subtitle = paste("Mean HFA MD:", round(mean(df_wide_all$hfa, na.rm=T), 2),
                                       "\nMean Retinalogik MD:", round(mean(df_wide_all$retinalogik, na.rm=T), 2))))
    } else {
      cat("Note: Could not create Bland-Altman plot for the selected qualifying patients. Ensure both 'hfa' and 'retinalogik' data are available for them.\n\n")
    }
  } else {
    cat("Note: Among the patients with qualifying visits, either 'hfa' or 'retinalogik' data is missing. Cannot create Bland-Altman plot.\n\n")
  }
} else {
  cat("Note: No patients found with qualifying visits for both 'hfa' and 'retinalogik'. Cannot create Bland-Altman plot.\n\n")
}
```

## Mean sensitivity and variability
Mean threshold sensitivities (SDs in parentheses) are shown for the latest reliable VF tests, where FL, FP, and FNs were below 30%. Only data from the eye with the worse MD were included. The grid plots below display results from both HFA and Retinalogik devices, followed by the mean differences in threshold sensitivities between the two devices and the SDs of those differences (in parentheses).

```{r variability, echo=FALSE, warning=FALSE, message=FALSE, fig.width=10, fig.height=10}
library(patchwork)

bs <- data.frame(x = c(15, 15),
                 y = c(-3, 3),
                 meanRT = NA, sdRT = NA,
                 meanHFA = NA, meanRT = NA,
                 meanDiff = NA, sdDiff = NA)

bs2 <- data.frame(x = c(15, 15),
                 y = c(-3, 3),
                 results = NA,
                 hfa = NA,
                 retinalogik = NA,
                 greater = NA)

meanvariability <- dBdat_flipped %>%
  filter(!(x == 15 & y == -3)) %>%
  filter(!(x == 15 & y == +3)) %>%
  distinct(id, device, visit, eye, x, y, dB) %>%
  right_join(min_md_eye_all) %>%
  pivot_wider(names_from = device, values_from = dB) %>%
  mutate(diff = hfa-retinalogik) %>%
  group_by(x, y) %>%
  summarize(meanRT = mean(retinalogik), sdRT = sd(retinalogik),
            meanHFA = mean(hfa), sdHFA = sd(hfa),
            meanDiff = mean(diff), sdDiff = sd(diff))

wilcoxdf <- dBdat_flipped %>%
  filter(!(x == 15 & y == -3)) %>%
  filter(!(x == 15 & y == +3)) %>%
  distinct(id, device, visit, eye, x, y, dB) %>%
  right_join(min_md_eye_all) %>%
  pivot_wider(names_from = device, values_from = dB) %>%
  # mutate(diff = hfa-retinalogik) %>%
  group_by(x, y) %>%
  summarize(results = wilcox.test(Pair(hfa, retinalogik) ~1)$p.value,
            hfa = mean(hfa), retinalogik = mean(retinalogik),
            hfagreater = case_when(hfa > retinalogik ~ "greater",
                                hfa < retinalogik ~ "lesser"))

hfaplot <- meanvariability %>%
  rbind(bs) %>%
  ggplot(aes(x, y, meanHFA)) +
  geom_raster(aes(x = x, y = y, fill = meanHFA)) +
  geom_text(aes(label = ifelse(is.na(meanHFA), "", paste0(round(meanHFA, 2), "\n(", round(sdHFA, 2), ")")),
                  x = x, y = y), size = 3) +
  coord_fixed(ratio = 1) +
  scale_fill_gradientn(colours = viridis(47), limits = c(-1, 46),
                       na.value="darkred") +
  theme_bw() +
  ggtitle("HFA") +
  theme(legend.position="none", plot.margin = unit(c(5.5, 5.5, 1, 5.5), "pt")) # Reduced bottom margin

rtplot <- meanvariability %>%
  rbind(bs) %>%
  ggplot(aes(x, y, meanRT)) +
  geom_raster(aes(x = x, y = y, fill = meanRT)) +
  geom_text(aes(label = ifelse(is.na(meanRT), "", paste0(round(meanRT, 2), "\n(", round(sdRT, 2), ")")),
                  x = x, y = y), size = 3) +
  coord_fixed(ratio = 1) +
  scale_fill_gradientn(colours = viridis(47), limits = c(-1, 46),
                       na.value="darkred") +
  theme_bw() +
  ggtitle("Retinalogik") +
  theme(legend.position="none", plot.margin = unit(c(5.5, 5.5, 1, 5.5), "pt")) # Reduced bottom margin

diffplot <- meanvariability %>%
  rbind(bs) %>%
  ggplot(aes(x, y, meanDiff)) +
  geom_raster(aes(x = x, y = y, fill = meanDiff)) +
  geom_text(aes(label = ifelse(is.na(meanDiff), "", paste0(round(meanDiff, 2), "\n(", round(sdDiff, 2), ")")),
                  x = x, y = y), size = 3) +
  coord_fixed(ratio = 1) +
  scale_fill_gradientn(colours = rainbow(11), limits = c(-5, 5),
                       na.value="darkred") +
  theme_bw() +
  ggtitle("HFA-Retinalogik") +
  theme(legend.position="none", plot.margin = unit(c(1, 5.5, 5.5, 5.5), "pt")) # Reduced top margin
  
pvalueplot <- wilcoxdf %>%
  rbind(bs2) %>%
  ggplot(aes(x, y, results)) +
  geom_raster(aes(x = x, y = y, fill = hfagreater)) +
  geom_text(aes(label = ifelse(results < 0.05, "<.01", round(results, 2)),
                x = x, y = y), size = 4) +
  coord_fixed(ratio = 1) +
  scale_fill_manual(values = c("thistle", "darksalmon"), na.value="darkred",
                    name = "HFA > Retinalogik?") +
  theme_bw() +
  ggtitle("HFA-Retinalogik p-values") +
  theme(legend.position="none", plot.margin = unit(c(1, 5.5, 5.5, 5.5), "pt")) # Reduced top margin

final_plot <- (hfaplot + rtplot) / (diffplot + pvalueplot) +
  plot_layout(heights = c(1, 1), widths = c(1, 1)) # Ensure equal heights

print(final_plot)

# Bland-Altman
# p <- meanvariability %>%
#   rbind(bs) %>%
#   bland.altman.plot(.$meanHFA, .$meanRT, graph.sys = "ggplot2")
# meanbias <- mean(mdpsd$md_hfa, na.rm=T) - mean(mdpsd$md_retinalogik)
# 
# p1 <- print(p + geom_smooth(method = "lm", se = FALSE) +
#               geom_hline(yintercept = 0, color = "black") +
#               geom_hline(yintercept = meanbias, color = "red", linetype = "solid", size = 2) +
#               annotate("text", x = -17, y = -2, label=paste("Mean bias =", round(meanbias, 2), "dB")) +
#               xlab("Mean deviation (MD)") +
#               ylab("Difference in mean deviation (dB)") +
#               ggtitle("Bland-Altman plots (HFA-Retinalogik)"))

# p <- bland.altman.plot(mdpsd$psd_hfa, mdpsd$psd_retinalogik, graph.sys = "ggplot2")
# meanbias <- mean(mdpsd$psd_hfa, na.rm=T) - mean(mdpsd$psd_retinalogik)
# 
# p2 <- print(p + geom_smooth(method = "lm", se = FALSE) +
#               geom_hline(yintercept = 0, color = "black") +
#               geom_hline(yintercept = meanbias, color = "red", linetype = "solid", size=2) +
#               annotate("text", x = 11, y = 1, label=paste("Mean bias =", round(meanbias, 2), "dB")) +
#               xlab("Pattern standard deviation (PSD)") + ylab("Difference in pattern standard deviation (dB)"))

# grid.arrange(p1, p2, ncol=2, top = "Bland-Altman plots")

# dBdat %>% 
#   distinct(id, age, visit, device, eye, md, psd) %>% 
#   datatable(extensions = 'Buttons', options = list(
#     dom = 'Bfrtip',
#     buttons = 
#       list('copy', 'print', list(
#         extend = 'collection',
#         buttons = c('csv', 'excel', 'pdf'),
#         text = 'Download'
#       ))
# ))
```

In the bottom right panel, Wilcoxon signed-rank tests were used for between-group comparisons. Salmon coloured areas represents higher Retinalogik thresholds compared to HFA whereas the lavender area represents greater HFA thresholds compared to Retinalogik.

## Scatter, Bland-Altman, & Hexbin plots
Scatterplot and Bland–Altman plots of VF sensitivity at each test point between the HFA and Retinalogik. Plotted only the latest reliable VF tests, where FL, FP, and FNs were below 30%. Only data from the eye with the worse MD were included. Pearson product-moment correlation coefficient (r) and p values are shown in the scatterplot.

```{r scatterblandaltman, warning=FALSE}

df <- dBdat_flipped %>%
  filter(!(x == 15 & y == -3)) %>%
  filter(!(x == 15 & y == +3)) %>%
  distinct(id, device, visit, eye, x, y, dB) %>%
  right_join(min_md_eye_all) %>%
  pivot_wider(names_from = device, values_from = dB) %>%
  mutate(retinalogik = as.numeric(retinalogik), hfa = as.numeric(hfa)) 

pearson <- cor.test(df$hfa, df$retinalogik, method = "pearson")

scatterplot0 <- df %>%
  ggplot(., aes(x=hfa, y=retinalogik)) + 
  geom_point(aes(color=id), position = "jitter") +
  geom_smooth(method = "loess") +
  annotate("text", x = 5, y = 34,
           label = paste0("r = ", round(pearson$estimate, 2), "\np ",
                   ifelse(pearson$p.value < 0.05, "< .01", round(pearson$p.value, 2)))) +
  scale_x_continuous(breaks = seq(0, 44, by=2), limits=c(-1,40)) +
  scale_y_continuous(breaks = seq(0, 44, by=2), limits=c(-1,40)) +
  geom_abline(intercept = 0, slope = 1, color = "gray50") +
  labs(title="Pointwise sensitivity for all locations") +
  xlab("HFA (dB)") +
  ylab("Retinalogik (dB)") +
  theme_bw()

ggplotly(scatterplot0)

blandplot <- bland.altman.plot(df$hfa, df$retinalogik, graph.sys = "ggplot2", geom_count=T)
bias <- mean(df$hfa) - mean(df$retinalogik)

print(blandplot +
        geom_smooth(method = "lm", se = FALSE) +
        # geom_point(position = "jitter") +
        geom_hline(yintercept = 0, color = "black") +
        geom_hline(yintercept = bias, color = "red", linetype = "solid", size = 2) +
        xlab("dB") +
        ylab("Difference in dB (HFA-Retinalogik)") +
        labs(subtitle=paste("Bias =", round(bias, 2), "dB")) +
        ggtitle("Bland-Altman plot for all locations (Lowest MD only)"))

```
```{r hexbin, fig.width=6.5, fig.height=6, warning=FALSE}
library(hexbin)
library(RColorBrewer)

# early <- dBdat %>%
#   filter(md < 0 & md > -6) %>%
#   mutate(x = as.numeric(x), y = as.numeric(y)) %>%
#   mutate(x = case_when(
#     eye == "L" ~ x * -1,
#     TRUE ~ as.numeric(x)
#   )) %>%
#   filter(!(x == 15 & y == -3)) %>%
#   filter(!(x == 15 & y == +3)) %>%
#   distinct(id, device, visit, eye, x, y, dB) %>%
#   #right_join(min_md_eye_all) %>%
#   pivot_wider(names_from = device, values_from = dB) %>%
#   mutate(retinalogik = as.numeric(retinalogik), hfa = as.numeric(hfa)) 

# bin <- hexbin(df$hfa, df$retinalogik, xbins=40)
# my_colors=colorRampPalette(rev(brewer.pal(11,'Spectral')))
# plot(bin, main="" , colramp=my_colors , legend=F ) 

ggplot(df, aes(x = df$hfa, y = df$retinalogik)) +
  geom_hex(bins = 20) +
  scale_fill_viridis_c() +
  labs(title="Hexbin plot of threshold sensitivities for all locations",
       subtitle="Lowest MD only (Size of bin = 20)") +
  xlab("HFA (dB)") +
  ylab("Retinalogik (dB)") +
  theme_bw()

```

## Deming regression
Deming regression (red line) applied to all sensitivity data combined. The area of each data point (blue circles) is proportional to the number of occurrences. Blindspots were removed from analysis.

```{r deming, message=FALSE, warning=FALSE}
library(MethComp)

demingwide <- dBdat_flipped %>%
  filter(!(x == 15 & y == -3)) %>%
  filter(!(x == 15 & y == +3)) %>%
  select(id, eye, visit, x, y, device, dB) %>%
  pivot_wider(names_from = device, values_from = dB) %>%
  drop_na()

error_var_df <- demingwide %>%
  group_by(id, eye, x, y) %>%
  summarise(
    n = n(),
    var_H = ifelse(n > 1, var(hfa, na.rm = TRUE), NA),
    var_R = ifelse(n > 1, var(retinalogik, na.rm = TRUE), NA),
    .groups = "drop"
  ) %>%
  filter(n > 1)

# Mean within-location variances
sigma2_H <- mean(error_var_df$var_H, na.rm = TRUE)
sigma2_R <- mean(error_var_df$var_R, na.rm = TRUE)

# Error ratio for Deming regression
lambda <- sigma2_R / sigma2_H

# Run Deming regression
deming_result <- Deming(demingwide$hfa, demingwide$retinalogik, vr=lambda)

summary(deming_result)
intercept <- deming_result[1]
slope <- deming_result[2]

p <- demingwide %>%
  ggplot(., aes(x=hfa, y=retinalogik)) + 
  #geom_point(aes(color=id), position = "jitter") +
  scale_x_continuous(breaks = seq(0, 50, by=5), limits=c(-2,50)) +
  scale_y_continuous(breaks = seq(0, 50, by=5), limits=c(-2,50)) +
  geom_abline(intercept = 0, slope = 1, color = "gray50", linetype="dashed") +
  geom_abline(intercept = intercept, slope = slope, color = "red", linetype = "solid", size = 1) +  # Deming line
  geom_count(color = "blue") +
  annotate("text", x = 20, y = 40, label = paste("Slope =", round(slope, 2)), color = "red") +
  labs(title = "Include all dBs",
       x = "HFA",
       y = "Retinalogik") +
  theme_bw()

ggplotly(p, width = 600, height= 600)

```

## Individual Bland-Altman plots {.tabset}
Shown in each tab below are the Bland-Altman plots for each patient's latest reliable VF where FL, FP, and FN were <30%. The worse eye according to their MD is plotted.

```{r onlyreliable, message = FALSE, warning=FALSE, results='asis'}
library(ggrepel)

dBdat_split <- dBdat_flipped %>%
  filter(!(x == 15 & y == -3)) %>%
  filter(!(x == 15 & y == +3)) %>%
  group_split(id)

# Process each data frame and generate Bland-Altman plots
for (i in seq_along(dBdat_split)) {
  current_id <- unique(dBdat_split[[i]]$id)

  # Function to check a specific visit
  check_visit <- function(df, visit_number) {
    df %>%
      filter(visit == visit_number) %>%
      group_by(eye) %>%
      mutate(is_below_30 = flPerc < 30 & fpPerc < 30 & fnPerc < 30) %>%
      filter(all(is_below_30)) %>%
      ungroup()
  }

  # Get all unique visit numbers in descending order
  unique_visits <- dBdat_split[[i]] %>%
    pull(visit) %>%
    unique() %>%
    sort(decreasing = TRUE)

  processed_data_single_id <- tibble::tibble() # Initialize as empty
  qualifying_visit <- NA

  # Iterate through visits from latest to earliest
  if (length(unique_visits) == 1) {
    # Only one visit, check it directly
    visit_number <- unique_visits[1]
    temp_processed_data <- dBdat_split[[i]] %>%
      check_visit(visit_number)
    if (nrow(temp_processed_data) > 0) {
      processed_data_single_id <- temp_processed_data
      qualifying_visit <- visit_number
    }
  } else {
    # Iterate through visits from latest to earliest
    for (visit_number in unique_visits) {
      temp_processed_data <- dBdat_split[[i]] %>%
        check_visit(visit_number)
      if (nrow(temp_processed_data) > 0) {
        processed_data_single_id <- temp_processed_data
        qualifying_visit <- visit_number
        break # Stop checking once a qualifying visit is found
      }
    }
  }

  cat(paste0('\n\n### ', current_id, '\n\n')) # Tab title with more newlines

  if ("hfa" %in% unique(dBdat$device) && "retinalogik" %in% unique(dBdat$device)) {
    if (nrow(processed_data_single_id) > 0) {
      # Check if the qualifying visit has data for both devices
      devices_in_qualifying_visit <- unique(processed_data_single_id$device)
      has_hfa <- "hfa" %in% devices_in_qualifying_visit
      has_retinalogik <- "retinalogik" %in% devices_in_qualifying_visit

      if (has_hfa && has_retinalogik) {
        min_md_eye <- processed_data_single_id %>%
          group_by(eye) %>%
          summarise(min_md = min(md, na.rm = TRUE)) %>%
          filter(min_md == min(min_md, na.rm = TRUE)) %>%
          pull(eye) # Extract the eye value(s)

        md_values <- processed_data_single_id %>%
          filter(eye %in% min_md_eye) %>%
          distinct(device, md)

        df_wide <- processed_data_single_id %>%
          filter(eye %in% min_md_eye) %>%
          distinct(id, visit, device, eye, x, y, dB) %>%
          pivot_wider(names_from = device, values_from = dB) %>%
          mutate(location = paste0("(", x, ", ", y, ")"))

        if (nrow(df_wide) > 0 && all(c("hfa", "retinalogik") %in% names(df_wide))) {
          df_wide$difference <- df_wide$hfa - df_wide$retinalogik
          df_wide$average <- (df_wide$hfa + df_wide$retinalogik) / 2

          p <- ggplot(df_wide, aes(x = average, y = difference, label = location)) +
            geom_point(position = "jitter") +
            geom_smooth(method = "lm", se = FALSE) +
            geom_hline(yintercept = 0, color = "black") +
            geom_label_repel(
              aes(label = ifelse(abs(difference) > 5, as.character(location), '')),
              box.padding = 0.35,
              point.padding = 0.5,
              segment.color = 'grey50'
            ) +
            xlab("dB") +
            ylab("Difference in dB (HFA - Retinalogik)") +
            ggtitle(
              paste("ID:", current_id,
                    "| Visit:", qualifying_visit,
                    "| Eye:", paste(min_md_eye, collapse = ", ")),
              subtitle = paste(
                "HFA MD:", round(md_values$md[md_values$device == "hfa"], 2),
                "\nRetinalogik MD:", round(md_values$md[md_values$device == "retinalogik"], 2)
              )
            )

          print(p)

        } else {
          cat("Note: Could not create Bland-Altman plot for ID:", current_id, ". Missing 'hfa' or 'retinalogik' dB values in the qualifying visit.\n\n")
        }
      } else {
        cat("Note: Could not create Bland-Altman plot for ID:", current_id, ". Qualifying visit does not contain data for both 'hfa' and 'retinalogik'.\n\n")
      }
    } else {
      cat("Note: No qualifying visit found for ID:", current_id, " based on the <30% criteria.\n\n")
    }
  } else {
    cat("Note: Either 'hfa' or 'retinalogik' is missing in the overall dataset. Cannot proceed with Bland-Altman plots.\n\n")
  }
  cat('\n')
}
```

## Test-retest variability {.tabset}
Calculate the SD across all MDs, and divide that by the absolute MD:

### Early
```{r testretest_early, warning=FALSE}
dBdat %>%
  filter(md < 0 & md > -6) %>%
  distinct(id, device, visit, eye, md) %>%
  group_by(device) %>%
  summarize(mean = mean(md), sd = sd(md)) %>%
  mutate(variability = sd/mean) %>%
  nice_table()
```

### Moderate
```{r testretest_moderate, warning=FALSE}
dBdat %>%
  filter(md < -6 & md > -12) %>%
  distinct(id, device, visit, eye, md) %>%
  group_by(device) %>%
  summarize(mean = mean(md), sd = sd(md)) %>%
  mutate(variability = sd/mean) %>%
  nice_table()
```

### Advanced
```{r testretest_advanced, warning=FALSE}
dBdat %>%
  filter(md < -12) %>%
  distinct(id, device, visit, eye, md) %>%
  group_by(device) %>%
  summarize(mean = mean(md), sd = sd(md)) %>%
  mutate(variability = sd/mean) %>%
  nice_table()
```

## Test duration comparison

```{r duration, warning=FALSE}
duration <- dBdat %>%
  group_by(device) %>%
  mutate(mean = mean(as.duration(duration)), sd = sd(as.duration(duration))) %>%
  distinct(device, mean, sd)

rtduration <- dBdat %>%
  filter(device=="retinalogik") %>%
  distinct(id, visit, eye, duration)

hfaduration <- dBdat %>%
  filter(device=="hfa") %>%
  distinct(id, visit, eye, duration)

t.test(as.numeric(rtduration$duration), as.numeric(hfaduration$duration), paired=T)

```

Mean test duration time was `r round(duration$mean[duration$device=="retinalogik"]/60, 2)` (SD = `r round(duration$sd[duration$device=="retinalogik"]/60, 2)`) mins for Retinalogik and `r round(duration$mean[duration$device=="hfa"]/60, 2)` (SD = `r round(duration$sd[duration$device=="hfa"]/60, 2)`) mins for HFA (p < .01).

## Scatterplots

```{r scatterplot2, warning=FALSE}
widedat <- dBdat %>%
  select(id, device, visit, eye, x, y, dB) %>%
  pivot_wider(., names_from = device, values_from = dB) %>%
  # mutate(retinalogik = ifelse(retinalogik == "<0", "-1", retinalogik)) %>%
  mutate(retinalogik = as.numeric(retinalogik), hfa = as.numeric(hfa)) %>%
  na.omit()

scatterplot <- widedat %>%
  ggplot(., aes(x=hfa, y=retinalogik)) + 
  geom_point(aes(color=id), position = "jitter") +
  geom_smooth(method = "loess") +
  scale_x_continuous(breaks = seq(0, 44, by=2), limits=c(-1,40)) +
  scale_y_continuous(breaks = seq(0, 44, by=2), limits=c(-1,40)) +
  geom_abline(intercept = 0, slope = 1, color = "gray50") +
  labs(title="Pointwise sensitivity for all visits and both eyes if available") +
  xlab("HFA (dB)") +
  ylab("Retinalogik (dB)") +
  theme_bw()

ggplotly(scatterplot)

```

# Discussion

# Data
For those interested in the cleaned up pointwise data.
```{r data}

dBdat %>% 
  select(id, age, device, eye, x, y, dB) %>% 
  datatable(options = list(
  order = list(1, 'asc')
))
```
