---
title: "Retinalogik vs HFA study"
author: "[Eadie Technologies, Inc.](https://www.eadietech.com/)"
date: "2025-02-06"
output: 
  html_document:
    theme: yeti
    code_folding: hide
    toc: true
    toc_depth: 2
    toc_float: 
      collapsed: true
    includes:
        after_body: footer.html
---

## Setup

### Versioning

Last updated on `r format(Sys.time(), '%Y-%b-%d')` at `r format(Sys.time(), '%I:%M %p')`.

* 2025-Feb-06: Created git & RStudio project
* 2025-Feb-11: Added individual plots
* 2025-Feb-18: Added cleaned up data and placeholder for global sensitivity indices
* 2025-Feb-19: Added Bland-Altman plots for MD & PSD and scatterplots for pointwise sensitivity data. Also separated individual plots into tabs.


### Acknowledgements
Data collected by Research Assistant Reann Post and Research Coordinator Lynn Murphy.

```{r setup, collapse=TRUE}

# load packages
library(tidyverse)
library(here)
library(DT)
library(plotly)
library(viridis)
library(svglite)
library(htmltools)
library(gridExtra)

sessionInfo()

# default chunk options
knitr::opts_chunk$set(
  comment = '>', cache = TRUE, collapse = TRUE, cache = FALSE, dev= c("png")
  )

# load processed data
load(here("dBdat.Rda"))

```

# Study summary
We compared 24-2 Full Threshold visual field (VF) test results between Retinalogik and Humphrey Field Analyzer (HFA) in patients with early glaucoma and moderate to advanced glaucoma.

# Methods (briefly)
## Subjects

Eligible participants were identified among patients of Dr. Brennan Eadie at the Eadie Eye Centre. If deemed eligible for the study, subjects were recruited consecutively.

Each participant underwent five study visits. At each visit, they performed a VF test on both eyes using two devices. The order of device tested was randomized at the first (baseline) study visit.

The study adhered to the tenets of the Declaration of Helsinki for research involving human subjects and the protocol was approved by the Nova Scotia Health Research Ethics Board (#1030608). All participants gave their written informed consent before enrollment in the study.

## Analysis
```{r demographics, echo=FALSE}
gender <- dBdat %>%
  distinct(id, gender) %>%
  count(gender)

eyes <- dBdat %>%
  distinct(id, visit, eye) %>%
  count(eye)

```
All OS data will be transposed to OD format before analyses. `r sum(eyes$n)` eyes from `r length(unique(dBdat$id))` participants (`r sum(gender$n[gender$gender == "F"], na.rm = T)` females, `r sum(gender$n[gender$gender == "M"], na.rm = T)+1` males) aged `r min(dBdat$age)` to `r max(dBdat$age)` (*M* = `r round(mean(dBdat$age), 2)`, *SD* = `r round(sd(dBdat$age), 2)`) were included in the analysis.

# Results
## Individual plots {.tabset}
Click on each tab to view each patient's data. Swipe/scroll from right to left for subsequent visits. Currently not plotting RL05 data where patient withdrew from the study. Also not plotting where only HFA/Retinalogik data is available (e.g. RL04). 

### RL01

```{r rl01, warning=FALSE}
library(ggplot2)
library(slickR)

# make sure coordinates are numeric, then flip OS to OD for plotting purposes
dBdat %<>%
  mutate(x = as.numeric(x), y = as.numeric(y)) %>%
  mutate(x = case_when(
    eye == "L" & device == "hfa" ~ x*-1,
    TRUE ~ as.numeric(x)))

dBdat %>%
  mutate(dB = as.numeric(dB)) %>%
  filter(device == "hfa" & id == "RL01") %>%
  group_by(id, visit, eye) %>%
  arrange(id, visit, eye) %>%
  # Use the group_by %>% nest pattern to group data by id
  nest() %>% 
  # Use map2 so the id can be used as the title
  mutate(graphs = map2(data, id,
                       ~ggplot(data = .x, aes(x, y, dB)) +
                         geom_raster(aes(x = x, y = y, fill = dB)) +
                         geom_text(aes(label = dB, x = x, y = y), size = 8) +
                         coord_fixed(ratio = 1) +
                         scale_fill_gradientn(colours = viridis(47), limits = c(-1, 46),
                                              na.value="darkred") +
                         theme_bw() +
                         ggtitle(paste0("HFA ", id, " Visit ", visit, " Eye: ", eye))
                       )
         ) %>% 
  # pull is the pipe-able equivalent of .[['graphs']]
  pull(graphs) %>% 
  # Return the svg of graphs
  map(function(gr) svglite::xmlSVG(show(gr), standalone = TRUE)) -> 
  hfa_individual_graphs

dBdat %>%
  mutate(dB = as.numeric(dB)) %>%
  filter(device == "retinalogik" & id == "RL01") %>%
  group_by(id, visit, eye) %>%
  arrange(id, visit, eye) %>%
  # Use the group_by %>% nest pattern to group data by id
  nest() %>% 
  # Use map2 so the id can be used as the title
  mutate(graphs = map2(data, id,
                       ~ggplot(data = .x, aes(x, y, dB)) +
                         geom_raster(aes(x = x, y = y, fill = dB)) +
                         geom_text(aes(label = dB, x = x, y = y), size = 8) +
                         coord_fixed(ratio = 1) +
                         scale_fill_gradientn(colours = viridis(47), limits = c(-1, 46),
                                              na.value="darkred") +
                         theme_bw() +
                         ggtitle(paste0("Retinalogik ", id, " Visit ", visit, " Eye: ", eye))
                       )
         ) %>% 
  # pull is the pipe-able equivalent of .[['graphs']]
  pull(graphs) %>% 
  # Return the svg of graphs
  map(function(gr) svglite::xmlSVG(show(gr), standalone = TRUE)) -> 
  rt_individual_graphs

# carousels
hfa_plots <- slickR(hfa_individual_graphs, height = 350, width = "95%", padding = 0, slideId = "slick1") +
  settings(slidesToShow = 2, slidesToScroll = 2)

rt_plots <- slickR(rt_individual_graphs, height = 350, width = "95%", padding = 0, slideId = "slick2") +
  settings(slidesToShow = 2, slidesToScroll = 2)

rt_plots %synch% hfa_plots 
```

### RL02

```{r rl02, warning=FALSE}
dBdat %>%
  mutate(dB = as.numeric(dB)) %>%
  filter(device == "hfa" & id == "RL02") %>%
  group_by(id, visit, eye) %>%
  arrange(id, visit, eye) %>%
  # Use the group_by %>% nest pattern to group data by id
  nest() %>% 
  # Use map2 so the id can be used as the title
  mutate(graphs = map2(data, id,
                       ~ggplot(data = .x, aes(x, y, dB)) +
                         geom_raster(aes(x = x, y = y, fill = dB)) +
                         geom_text(aes(label = dB, x = x, y = y), size = 8) +
                         coord_fixed(ratio = 1) +
                         scale_fill_gradientn(colours = viridis(47), limits = c(-1, 46),
                                              na.value="darkred") +
                         theme_bw() +
                         ggtitle(paste0("HFA ", id, " Visit ", visit, " Eye: ", eye))
                       )
         ) %>% 
  # pull is the pipe-able equivalent of .[['graphs']]
  pull(graphs) %>% 
  # Return the svg of graphs
  map(function(gr) svglite::xmlSVG(show(gr), standalone = TRUE)) -> 
  hfa_individual_graphs

dBdat %>%
  mutate(dB = as.numeric(dB)) %>%
  filter(device == "retinalogik" & id == "RL02") %>%
  group_by(id, visit, eye) %>%
  arrange(id, visit, eye) %>%
  # Use the group_by %>% nest pattern to group data by id
  nest() %>% 
  # Use map2 so the id can be used as the title
  mutate(graphs = map2(data, id,
                       ~ggplot(data = .x, aes(x, y, dB)) +
                         geom_raster(aes(x = x, y = y, fill = dB)) +
                         geom_text(aes(label = dB, x = x, y = y), size = 8) +
                         coord_fixed(ratio = 1) +
                         scale_fill_gradientn(colours = viridis(47), limits = c(-1, 46),
                                              na.value="darkred") +
                         theme_bw() +
                         ggtitle(paste0("Retinalogik ", id, " Visit ", visit, " Eye: ", eye))
                       )
         ) %>% 
  # pull is the pipe-able equivalent of .[['graphs']]
  pull(graphs) %>% 
  # Return the svg of graphs
  map(function(gr) svglite::xmlSVG(show(gr), standalone = TRUE)) -> 
  rt_individual_graphs

# carousels
hfa_plots <- slickR(hfa_individual_graphs, height = 350, width = "95%", padding = 0, slideId = "slick3") +
  settings(slidesToShow = 2, slidesToScroll = 2)

rt_plots <- slickR(rt_individual_graphs, height = 350, width = "95%", padding = 0, slideId = "slick4") +
  settings(slidesToShow = 2, slidesToScroll = 2)

rt_plots %synch% hfa_plots 
```

### RL03

```{r rl03, warning=FALSE}
dBdat %>%
  mutate(dB = as.numeric(dB)) %>%
  filter(device == "hfa" & id == "RL03") %>%
  group_by(id, visit, eye) %>%
  arrange(id, visit, eye) %>%
  # Use the group_by %>% nest pattern to group data by id
  nest() %>% 
  # Use map2 so the id can be used as the title
  mutate(graphs = map2(data, id,
                       ~ggplot(data = .x, aes(x, y, dB)) +
                         geom_raster(aes(x = x, y = y, fill = dB)) +
                         geom_text(aes(label = dB, x = x, y = y), size = 8) +
                         coord_fixed(ratio = 1) +
                         scale_fill_gradientn(colours = viridis(47), limits = c(-1, 46),
                                              na.value="darkred") +
                         theme_bw() +
                         ggtitle(paste0("HFA ", id, " Visit ", visit, " Eye: ", eye))
                       )
         ) %>% 
  # pull is the pipe-able equivalent of .[['graphs']]
  pull(graphs) %>% 
  # Return the svg of graphs
  map(function(gr) svglite::xmlSVG(show(gr), standalone = TRUE)) -> 
  hfa_individual_graphs

dBdat %>%
  mutate(dB = as.numeric(dB)) %>%
  filter(device == "retinalogik" & id == "RL03") %>%
  group_by(id, visit, eye) %>%
  arrange(id, visit, eye) %>%
  # Use the group_by %>% nest pattern to group data by id
  nest() %>% 
  # Use map2 so the id can be used as the title
  mutate(graphs = map2(data, id,
                       ~ggplot(data = .x, aes(x, y, dB)) +
                         geom_raster(aes(x = x, y = y, fill = dB)) +
                         geom_text(aes(label = dB, x = x, y = y), size = 8) +
                         coord_fixed(ratio = 1) +
                         scale_fill_gradientn(colours = viridis(47), limits = c(-1, 46),
                                              na.value="darkred") +
                         theme_bw() +
                         ggtitle(paste0("Retinalogik ", id, " Visit ", visit, " Eye: ", eye))
                       )
         ) %>% 
  # pull is the pipe-able equivalent of .[['graphs']]
  pull(graphs) %>% 
  # Return the svg of graphs
  map(function(gr) svglite::xmlSVG(show(gr), standalone = TRUE)) -> 
  rt_individual_graphs

# carousels
hfa_plots <- slickR(hfa_individual_graphs, height = 350, width = "95%", padding = 0) +
  settings(slidesToShow = 2, slidesToScroll = 2)

rt_plots <- slickR(rt_individual_graphs, height = 350, width = "95%", padding = 0) +
  settings(slidesToShow = 2, slidesToScroll = 2)

rt_plots %synch% hfa_plots 
```

### RL06

```{r rl06, warning=FALSE}
dBdat %>%
  mutate(dB = as.numeric(dB)) %>%
  filter(device == "hfa" & id == "RL06") %>%
  group_by(id, visit, eye) %>%
  arrange(id, visit, eye) %>%
  # Use the group_by %>% nest pattern to group data by id
  nest() %>% 
  # Use map2 so the id can be used as the title
  mutate(graphs = map2(data, id,
                       ~ggplot(data = .x, aes(x, y, dB)) +
                         geom_raster(aes(x = x, y = y, fill = dB)) +
                         geom_text(aes(label = dB, x = x, y = y), size = 8) +
                         coord_fixed(ratio = 1) +
                         scale_fill_gradientn(colours = viridis(47), limits = c(-1, 46),
                                              na.value="darkred") +
                         theme_bw() +
                         ggtitle(paste0("HFA ", id, " Visit ", visit, " Eye: ", eye))
                       )
         ) %>% 
  # pull is the pipe-able equivalent of .[['graphs']]
  pull(graphs) %>% 
  # Return the svg of graphs
  map(function(gr) svglite::xmlSVG(show(gr), standalone = TRUE)) -> 
  hfa_individual_graphs

dBdat %>%
  mutate(dB = as.numeric(dB)) %>%
  filter(device == "retinalogik" & id == "RL06") %>%
  group_by(id, visit, eye) %>%
  arrange(id, visit, eye) %>%
  # Use the group_by %>% nest pattern to group data by id
  nest() %>% 
  # Use map2 so the id can be used as the title
  mutate(graphs = map2(data, id,
                       ~ggplot(data = .x, aes(x, y, dB)) +
                         geom_raster(aes(x = x, y = y, fill = dB)) +
                         geom_text(aes(label = dB, x = x, y = y), size = 8) +
                         coord_fixed(ratio = 1) +
                         scale_fill_gradientn(colours = viridis(47), limits = c(-1, 46),
                                              na.value="darkred") +
                         theme_bw() +
                         ggtitle(paste0("Retinalogik ", id, " Visit ", visit, " Eye: ", eye))
                       )
         ) %>% 
  # pull is the pipe-able equivalent of .[['graphs']]
  pull(graphs) %>% 
  # Return the svg of graphs
  map(function(gr) svglite::xmlSVG(show(gr), standalone = TRUE)) -> 
  rt_individual_graphs

# carousels
hfa_plots <- slickR(hfa_individual_graphs, height = 350, width = "95%", padding = 0) +
  settings(slidesToShow = 2, slidesToScroll = 2)

rt_plots <- slickR(rt_individual_graphs, height = 350, width = "95%", padding = 0) +
  settings(slidesToShow = 2, slidesToScroll = 2)

rt_plots %synch% hfa_plots 
```

### RL07

```{r rl07, warning=FALSE}
# make sure coordinates are numeric, then flip OS to OD for plotting purposes
dBdat %>%
  mutate(dB = as.numeric(dB)) %>%
  filter(device == "hfa" & id == "RL07") %>%
  group_by(id, visit, eye) %>%
  arrange(id, visit, eye) %>%
  # Use the group_by %>% nest pattern to group data by id
  nest() %>% 
  # Use map2 so the id can be used as the title
  mutate(graphs = map2(data, id,
                       ~ggplot(data = .x, aes(x, y, dB)) +
                         geom_raster(aes(x = x, y = y, fill = dB)) +
                         geom_text(aes(label = dB, x = x, y = y), size = 8) +
                         coord_fixed(ratio = 1) +
                         scale_fill_gradientn(colours = viridis(47), limits = c(-1, 46),
                                              na.value="darkred") +
                         theme_bw() +
                         ggtitle(paste0("HFA ", id, " Visit ", visit, " Eye: ", eye))
                       )
         ) %>% 
  # pull is the pipe-able equivalent of .[['graphs']]
  pull(graphs) %>% 
  # Return the svg of graphs
  map(function(gr) svglite::xmlSVG(show(gr), standalone = TRUE)) -> 
  hfa_individual_graphs

dBdat %>%
  mutate(dB = as.numeric(dB)) %>%
  filter(device == "retinalogik" & id == "RL07") %>%
  group_by(id, visit, eye) %>%
  arrange(id, visit, eye) %>%
  # Use the group_by %>% nest pattern to group data by id
  nest() %>% 
  # Use map2 so the id can be used as the title
  mutate(graphs = map2(data, id,
                       ~ggplot(data = .x, aes(x, y, dB)) +
                         geom_raster(aes(x = x, y = y, fill = dB)) +
                         geom_text(aes(label = dB, x = x, y = y), size = 8) +
                         coord_fixed(ratio = 1) +
                         scale_fill_gradientn(colours = viridis(47), limits = c(-1, 46),
                                              na.value="darkred") +
                         theme_bw() +
                         ggtitle(paste0("Retinalogik ", id, " Visit ", visit, " Eye: ", eye))
                       )
         ) %>% 
  # pull is the pipe-able equivalent of .[['graphs']]
  pull(graphs) %>% 
  # Return the svg of graphs
  map(function(gr) svglite::xmlSVG(show(gr), standalone = TRUE)) -> 
  rt_individual_graphs

# carousels
hfa_plots <- slickR(hfa_individual_graphs, height = 350, width = "95%", padding = 0) +
  settings(slidesToShow = 1, slidesToScroll = 1)

rt_plots <- slickR(rt_individual_graphs, height = 350, width = "95%", padding = 0) +
  settings(slidesToShow = 1, slidesToScroll = 1)

rt_plots %synch% hfa_plots 
```

### RL08

```{r rl08, warning=FALSE}
# make sure coordinates are numeric, then flip OS to OD for plotting purposes
dBdat %>%
  mutate(dB = as.numeric(dB)) %>%
  filter(device == "hfa" & id == "RL08") %>%
  group_by(id, visit, eye) %>%
  arrange(id, visit, eye) %>%
  # Use the group_by %>% nest pattern to group data by id
  nest() %>% 
  # Use map2 so the id can be used as the title
  mutate(graphs = map2(data, id,
                       ~ggplot(data = .x, aes(x, y, dB)) +
                         geom_raster(aes(x = x, y = y, fill = dB)) +
                         geom_text(aes(label = dB, x = x, y = y), size = 8) +
                         coord_fixed(ratio = 1) +
                         scale_fill_gradientn(colours = viridis(47), limits = c(-1, 46),
                                              na.value="darkred") +
                         theme_bw() +
                         ggtitle(paste0("HFA ", id, " Visit ", visit, " Eye: ", eye))
                       )
         ) %>% 
  # pull is the pipe-able equivalent of .[['graphs']]
  pull(graphs) %>% 
  # Return the svg of graphs
  map(function(gr) svglite::xmlSVG(show(gr), standalone = TRUE)) -> 
  hfa_individual_graphs

dBdat %>%
  mutate(dB = as.numeric(dB)) %>%
  filter(device == "retinalogik" & id == "RL08") %>%
  group_by(id, visit, eye) %>%
  arrange(id, visit, eye) %>%
  # Use the group_by %>% nest pattern to group data by id
  nest() %>% 
  # Use map2 so the id can be used as the title
  mutate(graphs = map2(data, id,
                       ~ggplot(data = .x, aes(x, y, dB)) +
                         geom_raster(aes(x = x, y = y, fill = dB)) +
                         geom_text(aes(label = dB, x = x, y = y), size = 8) +
                         coord_fixed(ratio = 1) +
                         scale_fill_gradientn(colours = viridis(47), limits = c(-1, 46),
                                              na.value="darkred") +
                         theme_bw() +
                         ggtitle(paste0("Retinalogik ", id, " Visit ", visit, " Eye: ", eye))
                       )
         ) %>% 
  # pull is the pipe-able equivalent of .[['graphs']]
  pull(graphs) %>% 
  # Return the svg of graphs
  map(function(gr) svglite::xmlSVG(show(gr), standalone = TRUE)) -> 
  rt_individual_graphs

# carousels
hfa_plots <- slickR(hfa_individual_graphs, height = 350, width = "95%", padding = 0) +
  settings(slidesToShow = 2, slidesToScroll = 2)

rt_plots <- slickR(rt_individual_graphs, height = 350, width = "95%", padding = 0) +
  settings(slidesToShow = 2, slidesToScroll = 2)

rt_plots %synch% hfa_plots 
```

### RL09

```{r rl09, warning=FALSE}
# make sure coordinates are numeric, then flip OS to OD for plotting purposes
dBdat %>%
  mutate(dB = as.numeric(dB)) %>%
  filter(device == "hfa" & id == "RL09") %>%
  group_by(id, visit, eye) %>%
  arrange(id, visit, eye) %>%
  # Use the group_by %>% nest pattern to group data by id
  nest() %>% 
  # Use map2 so the id can be used as the title
  mutate(graphs = map2(data, id,
                       ~ggplot(data = .x, aes(x, y, dB)) +
                         geom_raster(aes(x = x, y = y, fill = dB)) +
                         geom_text(aes(label = dB, x = x, y = y), size = 8) +
                         coord_fixed(ratio = 1) +
                         scale_fill_gradientn(colours = viridis(47), limits = c(-1, 46),
                                              na.value="darkred") +
                         theme_bw() +
                         ggtitle(paste0("HFA ", id, " Visit ", visit, " Eye: ", eye))
                       )
         ) %>% 
  # pull is the pipe-able equivalent of .[['graphs']]
  pull(graphs) %>% 
  # Return the svg of graphs
  map(function(gr) svglite::xmlSVG(show(gr), standalone = TRUE)) -> 
  hfa_individual_graphs

dBdat %>%
  mutate(dB = as.numeric(dB)) %>%
  filter(device == "retinalogik" & id == "RL09") %>%
  group_by(id, visit, eye) %>%
  arrange(id, visit, eye) %>%
  # Use the group_by %>% nest pattern to group data by id
  nest() %>% 
  # Use map2 so the id can be used as the title
  mutate(graphs = map2(data, id,
                       ~ggplot(data = .x, aes(x, y, dB)) +
                         geom_raster(aes(x = x, y = y, fill = dB)) +
                         geom_text(aes(label = dB, x = x, y = y), size = 8) +
                         coord_fixed(ratio = 1) +
                         scale_fill_gradientn(colours = viridis(47), limits = c(-1, 46),
                                              na.value="darkred") +
                         theme_bw() +
                         ggtitle(paste0("Retinalogik ", id, " Visit ", visit, " Eye: ", eye))
                       )
         ) %>% 
  # pull is the pipe-able equivalent of .[['graphs']]
  pull(graphs) %>% 
  # Return the svg of graphs
  map(function(gr) svglite::xmlSVG(show(gr), standalone = TRUE)) -> 
  rt_individual_graphs

# carousels
hfa_plots <- slickR(hfa_individual_graphs, height = 350, width = "95%", padding = 0) +
  settings(slidesToShow = 2, slidesToScroll = 2)

rt_plots <- slickR(rt_individual_graphs, height = 350, width = "95%", padding = 0) +
  settings(slidesToShow = 2, slidesToScroll = 2)

rt_plots %synch% hfa_plots 
```

### RL10

```{r rl10, warning=FALSE}
# make sure coordinates are numeric, then flip OS to OD for plotting purposes
dBdat %>%
  mutate(dB = as.numeric(dB)) %>%
  filter(device == "hfa" & id == "RL10") %>%
  group_by(id, visit, eye) %>%
  arrange(id, visit, eye) %>%
  # Use the group_by %>% nest pattern to group data by id
  nest() %>% 
  # Use map2 so the id can be used as the title
  mutate(graphs = map2(data, id,
                       ~ggplot(data = .x, aes(x, y, dB)) +
                         geom_raster(aes(x = x, y = y, fill = dB)) +
                         geom_text(aes(label = dB, x = x, y = y), size = 8) +
                         coord_fixed(ratio = 1) +
                         scale_fill_gradientn(colours = viridis(47), limits = c(-1, 46),
                                              na.value="darkred") +
                         theme_bw() +
                         ggtitle(paste0("HFA ", id, " Visit ", visit, " Eye: ", eye))
                       )
         ) %>% 
  # pull is the pipe-able equivalent of .[['graphs']]
  pull(graphs) %>% 
  # Return the svg of graphs
  map(function(gr) svglite::xmlSVG(show(gr), standalone = TRUE)) -> 
  hfa_individual_graphs

dBdat %>%
  mutate(dB = as.numeric(dB)) %>%
  filter(device == "retinalogik" & id == "RL10") %>%
  group_by(id, visit, eye) %>%
  arrange(id, visit, eye) %>%
  # Use the group_by %>% nest pattern to group data by id
  nest() %>% 
  # Use map2 so the id can be used as the title
  mutate(graphs = map2(data, id,
                       ~ggplot(data = .x, aes(x, y, dB)) +
                         geom_raster(aes(x = x, y = y, fill = dB)) +
                         geom_text(aes(label = dB, x = x, y = y), size = 8) +
                         coord_fixed(ratio = 1) +
                         scale_fill_gradientn(colours = viridis(47), limits = c(-1, 46),
                                              na.value="darkred") +
                         theme_bw() +
                         ggtitle(paste0("Retinalogik ", id, " Visit ", visit, " Eye: ", eye))
                       )
         ) %>% 
  # pull is the pipe-able equivalent of .[['graphs']]
  pull(graphs) %>% 
  # Return the svg of graphs
  map(function(gr) svglite::xmlSVG(show(gr), standalone = TRUE)) -> 
  rt_individual_graphs

# carousels
hfa_plots <- slickR(hfa_individual_graphs, height = 350, width = "95%", padding = 0) +
  settings(slidesToShow = 2, slidesToScroll = 2)

rt_plots <- slickR(rt_individual_graphs, height = 350, width = "95%", padding = 0) +
  settings(slidesToShow = 2, slidesToScroll = 2)

rt_plots %synch% hfa_plots 
```

### RL11

```{r rl11, warning=FALSE}
# make sure coordinates are numeric, then flip OS to OD for plotting purposes
dBdat %>%
  mutate(dB = as.numeric(dB)) %>%
  filter(device == "hfa" & id == "RL11") %>%
  group_by(id, visit, eye) %>%
  arrange(id, visit, eye) %>%
  # Use the group_by %>% nest pattern to group data by id
  nest() %>% 
  # Use map2 so the id can be used as the title
  mutate(graphs = map2(data, id,
                       ~ggplot(data = .x, aes(x, y, dB)) +
                         geom_raster(aes(x = x, y = y, fill = dB)) +
                         geom_text(aes(label = dB, x = x, y = y), size = 8) +
                         coord_fixed(ratio = 1) +
                         scale_fill_gradientn(colours = viridis(47), limits = c(-1, 46),
                                              na.value="darkred") +
                         theme_bw() +
                         ggtitle(paste0("HFA ", id, " Visit ", visit, " Eye: ", eye))
                       )
         ) %>% 
  # pull is the pipe-able equivalent of .[['graphs']]
  pull(graphs) %>% 
  # Return the svg of graphs
  map(function(gr) svglite::xmlSVG(show(gr), standalone = TRUE)) -> 
  hfa_individual_graphs

dBdat %>%
  mutate(dB = as.numeric(dB)) %>%
  filter(device == "retinalogik" & id == "RL11") %>%
  group_by(id, visit, eye) %>%
  arrange(id, visit, eye) %>%
  # Use the group_by %>% nest pattern to group data by id
  nest() %>% 
  # Use map2 so the id can be used as the title
  mutate(graphs = map2(data, id,
                       ~ggplot(data = .x, aes(x, y, dB)) +
                         geom_raster(aes(x = x, y = y, fill = dB)) +
                         geom_text(aes(label = dB, x = x, y = y), size = 8) +
                         coord_fixed(ratio = 1) +
                         scale_fill_gradientn(colours = viridis(47), limits = c(-1, 46),
                                              na.value="darkred") +
                         theme_bw() +
                         ggtitle(paste0("Retinalogik ", id, " Visit ", visit, " Eye: ", eye))
                       )
         ) %>% 
  # pull is the pipe-able equivalent of .[['graphs']]
  pull(graphs) %>% 
  # Return the svg of graphs
  map(function(gr) svglite::xmlSVG(show(gr), standalone = TRUE)) -> 
  rt_individual_graphs

# carousels
hfa_plots <- slickR(hfa_individual_graphs, height = 350, width = "95%", padding = 0) +
  settings(slidesToShow = 2, slidesToScroll = 2)

rt_plots <- slickR(rt_individual_graphs, height = 350, width = "95%", padding = 0) +
  settings(slidesToShow = 2, slidesToScroll = 2)

rt_plots %synch% hfa_plots 
```

### RL12

```{r rl12, warning=FALSE}
# make sure coordinates are numeric, then flip OS to OD for plotting purposes
dBdat %>%
  mutate(dB = as.numeric(dB)) %>%
  filter(device == "hfa" & id == "RL12") %>%
  group_by(id, visit, eye) %>%
  arrange(id, visit, eye) %>%
  # Use the group_by %>% nest pattern to group data by id
  nest() %>% 
  # Use map2 so the id can be used as the title
  mutate(graphs = map2(data, id,
                       ~ggplot(data = .x, aes(x, y, dB)) +
                         geom_raster(aes(x = x, y = y, fill = dB)) +
                         geom_text(aes(label = dB, x = x, y = y), size = 8) +
                         coord_fixed(ratio = 1) +
                         scale_fill_gradientn(colours = viridis(47), limits = c(-1, 46),
                                              na.value="darkred") +
                         theme_bw() +
                         ggtitle(paste0("HFA ", id, " Visit ", visit, " Eye: ", eye))
                       )
         ) %>% 
  # pull is the pipe-able equivalent of .[['graphs']]
  pull(graphs) %>% 
  # Return the svg of graphs
  map(function(gr) svglite::xmlSVG(show(gr), standalone = TRUE)) -> 
  hfa_individual_graphs

dBdat %>%
  mutate(dB = as.numeric(dB)) %>%
  filter(device == "retinalogik" & id == "RL12") %>%
  group_by(id, visit, eye) %>%
  arrange(id, visit, eye) %>%
  # Use the group_by %>% nest pattern to group data by id
  nest() %>% 
  # Use map2 so the id can be used as the title
  mutate(graphs = map2(data, id,
                       ~ggplot(data = .x, aes(x, y, dB)) +
                         geom_raster(aes(x = x, y = y, fill = dB)) +
                         geom_text(aes(label = dB, x = x, y = y), size = 8) +
                         coord_fixed(ratio = 1) +
                         scale_fill_gradientn(colours = viridis(47), limits = c(-1, 46),
                                              na.value="darkred") +
                         theme_bw() +
                         ggtitle(paste0("Retinalogik ", id, " Visit ", visit, " Eye: ", eye))
                       )
         ) %>% 
  # pull is the pipe-able equivalent of .[['graphs']]
  pull(graphs) %>% 
  # Return the svg of graphs
  map(function(gr) svglite::xmlSVG(show(gr), standalone = TRUE)) -> 
  rt_individual_graphs

# carousels
hfa_plots <- slickR(hfa_individual_graphs, height = 350, width = "95%", padding = 0) +
  settings(slidesToShow = 2, slidesToScroll = 2)

rt_plots <- slickR(rt_individual_graphs, height = 350, width = "95%", padding = 0) +
  settings(slidesToShow = 2, slidesToScroll = 2)

rt_plots %synch% hfa_plots 
```

## Global sensitivity indices
Bland–Altman analysis for mean deviation (top) and pattern standard deviation (bottom) comparing HFA and Retinalogik. The red solid line indicates the mean bias, the black dashed lines indicate the 95% limits of agreement, the blue solid line indicates the regression line and the black solid line indicates y = 0. For both panels, a more positive value indicates that the HFA returned a higher result, while a negative value indicates that the Retinalogik returned a higher result.

```{r mdpsd, warning=FALSE}
library(BlandAltmanLeh)

mdpsd <- dBdat %>% 
  distinct(id, age, visit, device, eye, md, psd) %>%
  pivot_wider(names_from = device,
              values_from = c(md, psd))

p <- bland.altman.plot(mdpsd$md_hfa, mdpsd$md_retinalogik, graph.sys = "ggplot2")
meanbias <- mean(mdpsd$md_hfa, na.rm=T) - mean(mdpsd$md_retinalogik)

p1 <- print(p + geom_smooth(method = "lm", se = FALSE) +
              geom_hline(yintercept = 0, color = "black") +
              geom_hline(yintercept = meanbias, color = "red", linetype = "solid", size = 2) +
              annotate("text", x = -17, y = -2, label=paste("Mean bias =", round(meanbias, 2), "dB")) +
              xlab("Mean deviation (MD)") +
              ylab("Difference in mean deviation (dB)") +
              ggtitle("Bland-Altman plots (HFA-Retinalogik)"))

p <- bland.altman.plot(mdpsd$psd_hfa, mdpsd$psd_retinalogik, graph.sys = "ggplot2")
meanbias <- mean(mdpsd$psd_hfa, na.rm=T) - mean(mdpsd$psd_retinalogik)

p2 <- print(p + geom_smooth(method = "lm", se = FALSE) +
              geom_hline(yintercept = 0, color = "black") +
              geom_hline(yintercept = meanbias, color = "red", linetype = "solid", size=2) +
              annotate("text", x = 11, y = 1, label=paste("Mean bias =", round(meanbias, 2), "dB")) +
              xlab("Pattern standard deviation (PSD)") + ylab("Difference in pattern standard deviation (dB)"))

# grid.arrange(p1, p2, ncol=2, top = "Bland-Altman plots")

dBdat %>% 
  distinct(id, age, visit, device, eye, md, psd) %>% 
  datatable(extensions = 'Buttons', options = list(
    dom = 'Bfrtip',
    buttons = 
      list('copy', 'print', list(
        extend = 'collection',
        buttons = c('csv', 'excel', 'pdf'),
        text = 'Download'
      ))
))
```

## Scatterplots

```{r scatterplots, warning=FALSE}
widedat <- dBdat %>%
  select(id, device, visit, eye, x, y, dB) %>%
  pivot_wider(., names_from = device, values_from = dB) %>%
  mutate(retinalogik = ifelse(retinalogik == "<0", "-1", retinalogik)) %>%
  mutate(retinalogik = as.numeric(retinalogik), hfa = as.numeric(hfa)) %>%
  na.omit()

scatterplot <- widedat %>%
  ggplot(., aes(x=hfa, y=retinalogik)) + 
  geom_point(aes(color=id), position = "jitter") +
  geom_smooth(method = "loess") +
  scale_x_continuous(breaks = seq(0, 44, by=2), limits=c(-1,40)) +
  scale_y_continuous(breaks = seq(0, 44, by=2), limits=c(-1,40)) +
  geom_abline(intercept = 0, slope = 1, color = "gray50") +
  labs(title="Pointwise sensitivity for all visits and both eyes if available") +
  xlab("HFA (dB)") +
  ylab("Retinalogik (dB)") +
  theme_bw()

ggplotly(scatterplot)

```

# Discussion

# Data
For those interested in the cleaned up pointwise data.
```{r data}

dBdat %>% 
  select(id, age, device, eye, x, y, dB) %>% 
  datatable(options = list(
  order = list(1, 'asc')
))
```
