---
title: "Retinalogik vs HFA study JAMA Ophthalmology paper"
author: "[Eadie Technologies, Inc.](https://www.eadietech.com/)"
date: "2025-06-11"
output: 
  html_document:
    theme: yeti
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: 
      collapsed: true
    includes:
        after_body: footer.html
---

## Setup

### Versioning

Last updated on `r format(Sys.time(), '%Y-%b-%d')` at `r format(Sys.time(), '%I:%M %p')`.

* 2025-Jun-11: Created separate .Rmd file as index.Rmd was getting very cluttered.
* 2025-Jun-26: Extract data based on classification during screening.
* 2025-Aug-25: Extract, clean, and run stats on questionnaire data.

### Acknowledgements
Data collected by Research Assistants Luc St-Pierre, Reann Post, and Research Coordinator Lynn Murphy.

```{r setup, collapse=TRUE}

# load packages
library(tidyverse)
library(here)
library(DT)
library(plotly)
library(viridis)
library(svglite)
library(htmltools)
library(gridExtra)
library(sf)
library(spdep)
library(RColorBrewer)
library(patchwork)
library(report)
library(effectsize)
library(rempsyc)
library(psych)

sessionInfo()

# default chunk options
knitr::opts_chunk$set(
  comment = '>', collapse = TRUE, cache = FALSE, dev= c("png")
  )

rm(list = ls()) 

# load processed data
load(here("dBdat.Rda"))

load(here("retinalogik_hfadatawpsd.Rda"))

classification <- readxl::read_excel("RetinaLogik Data Classification.xlsx") %>%
  mutate(
    # Extract the numeric part after "RL"
    numeric_part = as.numeric(str_extract(ID, "\\d+")),
    # Format the numeric part to have at least two digits with leading zeros
    formatted_numeric = str_pad(numeric_part, width = 2, side = "left", pad = "0"),
    # Combine "RL" with the formatted number
    ID = paste0("RL", formatted_numeric)
  ) %>%
  # Remove the temporary columns
  select(-numeric_part, -formatted_numeric) %>%
  rename(id = ID, classification = Classifcation, eye = `Eye (OD/OS)`) %>%
  mutate(eye = case_when(eye == "OS" ~ "L",
                         eye == "OD" ~ "R",
                         eye == "OU" ~ "R"))

dBdat_filtered <- dBdat %>%
  mutate(x = as.numeric(x), y = as.numeric(y)) %>%
  mutate(x = case_when(
    eye == "L" & device == "retinalogik" ~ x*-1,
    TRUE ~ as.numeric(x)
  )) %>%
  # remove blind spots
  filter(!(x == 15 & y == -3)) %>%
  filter(!(x == 15 & y == +3)) %>%
  #remove first 2 visits
  #filter(visit != 1 & visit != 2) %>%
  #remove reliability indices <30%
  #filter(flPerc < 30 & fnPerc < 30 & fpPerc < 30) %>%
  group_by(id, device, eye) %>%
  mutate(meanmd = mean(md))

dBdat_filtered <- merge(dBdat_filtered, classification) %>%
  filter(classification != "Control")

```

### Download plots
[click me to find all generated plots](https://github.com/vvnejl/retinalogik-study/tree/a484221d13b84da1c1692b1bc471db54e97ddd30/plots)

# Study summary
We compared 24-2 visual field (VF) test results between Retinalogik and the Humphrey Field Analyzer (HFA) using SITA Standard test strategy in patients with early glaucoma and moderate to advanced glaucoma [here](https://vvnejl.github.io/retinalogik-study/). A scotoma analysis was performed following glaucomatous VF definitions found in [Park et al. (2011)](https://www.aaojournal.org/article/S0161-6420(11)00172-2/abstract). Scotoma analysis results may be found [here](https://vvnejl.github.io/retinalogik-study/scotoma_analysis.html).

# Methods (very briefly)
## Subjects
Eligible participants were identified among patients of Dr. Brennan Eadie at the Eadie Eye Centre. If deemed eligible for the study, subjects were recruited consecutively.

Each participant underwent five study visits. At each visit, they performed a VF test on both eyes using the two devices. The order of device tested was randomized at the first (baseline) study visit.

Exclusion criteria were:

* age <18 years
* diagnosis of secondary glaucoma
* diagnosis of non-glaucomatous optic neuropathy
* significant media opacity
* previous intraocular surgery
* pregnancy
* seizure disorder
* cardiac pacemaker or other implantable device
* severe vertigo or balance disturbance
* inability to demonstrate competence to make an informed decision regarding study participation

The study adhered to the tenets of the Declaration of Helsinki for research involving human subjects and the protocol was approved by the Nova Scotia Health Research Ethics Board (#1030608). All participants gave their written informed consent before enrollment in the study.

### Demographics
```{r demographics}
vft <- dBdat_filtered %>%
  distinct(id, visit, eye) %>%
  count(eye)

gender <- dBdat_filtered %>%
  distinct(id, gender) %>%
  count(gender)

groups <- dBdat_filtered %>%
  distinct(id, classification) %>%
  mutate(classification = as.factor(classification)) %>%
  count(classification)

```

All OS data are transposed to OD format before analyses. `r sum(vft$n)` VF tests from `r length(unique(dBdat_filtered$id))` participants (`r sum(gender$n[gender$gender == "F"], na.rm = T)` females, `r sum(gender$n[gender$gender == "M"], na.rm = T)` males) aged `r min(dBdat_filtered$age)` to `r max(dBdat_filtered$age)` (*M* = `r round(mean(dBdat_filtered$age), 2)`, *SD* = `r round(sd(dBdat_filtered$age), 2)`) were included in the analysis (Early n = `r groups$n[groups$classification == "Mild"]`; Moderate n = `r groups$n[groups$classification == "Moderate"]`; Advanced n = `r groups$n[groups$classification == "Advanced"]`).

# Results
## Analysis
Of the 5 visits, the first 2 visits were removed from analysis. VFTs where reliability indices (FL, FP, and FN) > 30% were removed from analysis as well.

### Mean sensitivity
In patients with moderate and advanced glaucoma, the mean sensitivity was significantly higher when using the VR-RL perimeter. Wilcoxon signed-rank tests confirmed this difference when looking at the median of the patient-level mean differences between visits.

```{r meansens}
dBdat_filtered <- dBdat_filtered %>%
  #remove first 2 visits
  filter(visit != 1 & visit != 2) %>%
  #remove reliability indices <30%
  filter(flPerc < 30 & fnPerc < 30 & fpPerc < 30)

dBdat_filtered %>%
  group_by(id, device) %>%
  mutate(mean = mean(dB)) %>%
  distinct(id, classification, device, mean) %>%
  pivot_wider(names_from = device, values_from = mean) %>%
  na.omit() %>%
  mutate(classification = factor(classification, levels = c("Mild", "Moderate", "Advanced"))) %>%
  group_by(classification) %>%
  summarise(
    n = n(),
    mean_retinalogik = mean(retinalogik, na.rm = TRUE),
    sd_retinalogik = sd(retinalogik, na.rm = TRUE),
    mean_hfa = mean(hfa, na.rm = TRUE),
    sd_hfa = sd(hfa, na.rm = TRUE)
  ) %>%
  mutate(difference = mean_retinalogik-mean_hfa) %>%
  nice_table()

#########
dBdat_means <- dBdat_filtered %>%
  group_by(id, device) %>%
  mutate(mean = mean(dB)) %>%
  distinct(id, classification, device, mean) %>%
  pivot_wider(names_from = device, values_from = mean) %>%
  na.omit() %>%
  mutate(difference = retinalogik-hfa)

# 1. Group the cleaned, wide data by classification
wilcoxon_results <- dBdat_means %>%
  group_by(classification) %>%
  
  # 2. Nest the data for group-wise application
  nest() %>%
  
  # 3. Apply the Wilcoxon test to the paired columns in each nested group
  mutate(
    wilcox_test = map(data, ~ wilcox.test(.x$retinalogik, .x$hfa, paired = TRUE)),
    
    # 4. Extract the key results cleanly
    W_statistic = map_dbl(wilcox_test, "statistic"),
    p_value = map_dbl(wilcox_test, "p.value")
  ) %>%
  
  # 5. Select the final output columns
  select(classification, W_statistic, p_value)

wilcoxon_results %>%
  mutate(classification = factor(classification, levels = c("Mild", "Moderate", "Advanced"))) %>%
  arrange(classification) %>%
  nice_table(title = c("Wilcoxon signed rank test on paired samples"))

dBdat_filtered %>%
  group_by(id, device) %>%
  mutate(median = median(dB)) %>%
  distinct(id, classification, device, median) %>%
  pivot_wider(names_from = device, values_from = median) %>%
  na.omit() %>%
  mutate(classification = factor(classification, levels = c("Mild", "Moderate", "Advanced"))) %>%
  group_by(classification) %>%
  summarise(
    n = n(),
    median_retinalogik = median(retinalogik, na.rm = TRUE),
    q1_retinalogik = quantile(retinalogik, .25, na.rm = TRUE),
    q3_retinalogik = quantile(retinalogik, .75, na.rm = TRUE),
    iqr_retinalogik = IQR(retinalogik, na.rm = TRUE),
    median_hfa = median(hfa, na.rm = TRUE),
    q1_hfa = quantile(hfa, .25, na.rm = TRUE),
    q3_hfa = quantile(hfa, .75, na.rm = TRUE),
    iqr_hfa = IQR(hfa, na.rm = TRUE)
  ) %>%
  nice_table()

# mild_data <- subset(dBdat_means, classification == "Mild")
# t.test(mild_data$retinalogik, mild_data$hfa, paired = TRUE) %>%
#   report()
# 
# moderate_data <- subset(dBdat_means, classification == "Moderate")
# t.test(moderate_data$retinalogik, moderate_data$hfa, paired = TRUE) %>%
#   report()
# 
# advanced_data <- subset(dBdat_means, classification == "Advanced")
# t.test(advanced_data$retinalogik, advanced_data$hfa, paired = TRUE) %>%
#   report() %>% summary()
# 
# dBdat_medians <- dBdat_filtered %>%
#   group_by(id, device) %>%
#   mutate(median = median(dB)) %>%
#   distinct(id, classification, device, median) %>%
#   pivot_wider(names_from = device, values_from = median) %>%
#   na.omit()
# 
# mild_tmp <- dBdat_filtered %>%
#   subset(classification == "Mild") %>%
#   pivot_wider(names_from = device, values_from = dB) %>%
#   na.omit()
# wilcox.test(mild_tmp$retinalogik, mild_tmp$hfa, paired=TRUE)
# 
# mild_data <- subset(dBdat_medians, classification == "Mild")
# wilcox.test(mild_data$retinalogik, mild_data$hfa, paired = TRUE)
# 
# moderate_data <- subset(dBdat_medians, classification == "Moderate")
# wilcox.test(moderate_data$retinalogik, moderate_data$hfa, paired = TRUE)
# 
# advanced_data <- subset(dBdat_medians, classification == "Advanced")
# wilcox.test(advanced_data$retinalogik, advanced_data$hfa, paired = TRUE)

```

```{r hexbin}
dBdat_filtered %>%
  group_by(id, device, x, y) %>%
  mutate(meandB = mean(dB)) %>%
  distinct(id, device, x, y, meandB) %>%
  pivot_wider(names_from = device, values_from = meandB) %>%
  mutate(retinalogik = as.numeric(retinalogik), hfa = as.numeric(hfa)) %>%
  ggplot(aes(x = hfa, y = retinalogik)) +
  geom_hex(bins = 30) +
  scale_fill_viridis_c() +
  geom_smooth(method = "loess", se = TRUE, colour = "red") +
  geom_abline(intercept = 0, slope = 1) +
  labs(title="Hexbin plot of threshold sensitivities for all locations",
       subtitle="Averaged across reliable datasets between visits 3 to 5\n(Size of bin = 30)") +
  xlab("HFA (dB)") +
  ylab("Retinalogik (dB)") +
  xlim(-2, 40) +
  ylim(-2, 40) +
  theme_bw()

ggsave("plots/hexbin30.png", width = 5, height = 4)
```
```{r mshexbin}
# dBdat_filtered %>%
#   group_by(id, device, visit) %>%
#   summarize(ms = mean(dB)) %>%
#   group_by(id, device) %>%
#   summarize(meanms = mean(ms)) %>%
#   pivot_wider(names_from = device, values_from = meanms) %>% 
#   mutate(retinalogik = as.numeric(retinalogik), hfa = as.numeric(hfa)) %>%
#   ggplot(aes(x = hfa, y = retinalogik)) +
#   geom_hex(bins = 5) +
#   scale_fill_viridis_c() +
#   geom_smooth(method = "loess", se = TRUE, colour = "red") +
#   geom_abline(intercept = 0, slope = 1) +
#   labs(title="Hexbin plot of mean sensitivities",
#        subtitle="Averaged across reliable datasets between visits 3 to 5\n(Size of bin = 5)") +
#   xlab("HFA (dB)") +
#   ylab("Retinalogik (dB)") +
#   xlim(0, 35) +
#   ylim(0, 35) +
#   theme_bw()

#ggsave("plots/mshexbin.png", width = 5, height = 4)
```

```{r msblandaltman}
library(BlandAltmanLeh)

ms_means <- dBdat_filtered %>%
  group_by(id, device, visit) %>%
  summarize(ms = mean(dB)) %>%
  group_by(id, device) %>%
  summarize(meanms = mean(ms)) %>%
  pivot_wider(names_from = device, values_from = meanms) %>%
  na.omit() %>%
  mutate(difference = retinalogik-hfa)

df <- ms_means

blandplot <- bland.altman.plot(df$retinalogik, df$hfa, graph.sys = "ggplot2", geom_count=T)
bias <- mean(df$retinalogik, na.rm=T) - mean(df$hfa, na.rm=T)

print(blandplot +
        geom_smooth(method = "lm", se = FALSE) +
        # geom_point(position = "jitter") +
        geom_hline(yintercept = 0, color = "black") +
        geom_hline(yintercept = bias, color = "red", linetype = "solid", linewidth = 1) +
        ylim(c(-10, 10)) +
        xlab("Mean Sensitivity (MS)") +
        ylab("Difference in MS (Retinalogik-HFA)") +
        labs(subtitle=paste("Bias =", round(bias, 2), "dB")) +
        ggtitle("Bland-Altman plot comparing MS across all stages of glaucoma"))

ggsave("plots/blandaltman-ms-allgon.png", width = 7, height = 5)

```

### Mean deviation
In patients with moderate and advanced glaucoma, the mean deviation was significantly higher when using the VR-RL perimeter. Wilcoxon signed-rank tests confirmed this difference when looking at the median of the patient-level mean differences between visits.

```{r mdtable}
dBdat_filtered %>%
  group_by(id, device) %>%
  mutate(mean = mean(md)) %>%
  distinct(id, classification, device, mean) %>%
  pivot_wider(names_from = device, values_from = mean) %>%
  na.omit() %>%
  mutate(classification = factor(classification, levels = c("Mild", "Moderate", "Advanced"))) %>%
  group_by(classification) %>%
  summarise(
    n = n(),
    mean_retinalogik = mean(retinalogik, na.rm = TRUE),
    sd_retinalogik = sd(retinalogik, na.rm = TRUE),
    mean_hfa = mean(hfa, na.rm = TRUE),
    sd_hfa = sd(hfa, na.rm = TRUE)
  ) %>%
  mutate(difference = mean_retinalogik-mean_hfa) %>%
  nice_table()

#########
md_means <- dBdat_filtered %>%
  group_by(id, device) %>%
  mutate(mean = mean(md)) %>%
  distinct(id, classification, device, mean) %>%
  pivot_wider(names_from = device, values_from = mean) %>%
  na.omit() %>%
  mutate(difference = retinalogik-hfa)

# 1. Group the cleaned, wide data by classification
wilcoxon_results <- md_means %>%
  group_by(classification) %>%
  
  # 2. Nest the data for group-wise application
  nest() %>%
  
  # 3. Apply the Wilcoxon test to the paired columns in each nested group
  mutate(
    wilcox_test = map(data, ~ wilcox.test(.x$retinalogik, .x$hfa, paired = TRUE)),
    
    # 4. Extract the key results cleanly
    W_statistic = map_dbl(wilcox_test, "statistic"),
    p_value = map_dbl(wilcox_test, "p.value")
  ) %>%
  
  # 5. Select the final output columns
  select(classification, W_statistic, p_value)

wilcoxon_results %>%
  mutate(classification = factor(classification, levels = c("Mild", "Moderate", "Advanced"))) %>%
  arrange(classification) %>%
  nice_table(title = c("Wilcoxon signed rank test on paired samples"))

dBdat_filtered %>%
  group_by(id, device) %>%
  mutate(median = median(md)) %>%
  distinct(id, classification, device, median) %>%
  pivot_wider(names_from = device, values_from = median) %>%
  na.omit() %>%
  mutate(classification = factor(classification, levels = c("Mild", "Moderate", "Advanced"))) %>%
  group_by(classification) %>%
  summarise(
    n = n(),
    median_retinalogik = median(retinalogik, na.rm = TRUE),
    q1_retinalogik = quantile(retinalogik, .25, na.rm = TRUE),
    q3_retinalogik = quantile(retinalogik, .75, na.rm = TRUE),
    iqr_retinalogik = IQR(retinalogik, na.rm = TRUE),
    median_hfa = median(hfa, na.rm = TRUE),
    q1_hfa = quantile(hfa, .25, na.rm = TRUE),
    q3_hfa = quantile(hfa, .75, na.rm = TRUE),
    iqr_hfa = IQR(hfa, na.rm = TRUE)
  ) %>%
  nice_table()

```

```{r mdhexbin}
# dBdat_filtered %>%
#   group_by(id, device, visit) %>%
#   summarize(md = mean(md)) %>%
#   group_by(id, device) %>%
#   summarize(meanmd = mean(md)) %>%
#   pivot_wider(names_from = device, values_from = meanmd) %>%
#   mutate(retinalogik = as.numeric(retinalogik), hfa = as.numeric(hfa)) %>%
#   ggplot(aes(x = hfa, y = retinalogik)) +
#   geom_hex(bins = 30) +
#   scale_fill_viridis_c() +
#   geom_smooth(method = "loess", se = TRUE, colour = "red") +
#   geom_abline(intercept = 0, slope = 1) +
#   labs(title="Hexbin plot of mean deviation",
#        subtitle="Averaged across reliable datasets between visits 3 to 5\n(Size of bin = 30)") +
#   xlab("HFA (dB)") +
#   ylab("Retinalogik (dB)") +
#   xlim(-30, 5) +
#   ylim(-30, 5) +
#   theme_bw()

#ggsave("plots/hexbin30.png", width = 5, height = 4)
```

```{r mdblandaltman}
library(BlandAltmanLeh)

df <- md_means

blandplot <- bland.altman.plot(df$retinalogik, df$hfa, graph.sys = "ggplot2", geom_count=T)
bias <- mean(df$retinalogik, na.rm=T) - mean(df$hfa, na.rm=T)

print(blandplot +
        geom_smooth(method = "lm", se = FALSE) +
        # geom_point(position = "jitter") +
        geom_hline(yintercept = 0, color = "black") +
        geom_hline(yintercept = bias, color = "red", linetype = "solid", linewidth = 1) +
        ylim(c(-15, 15)) +
        xlab("MD") +
        ylab("Difference in MD (Retinalogik-HFA)") +
        labs(subtitle=paste("Bias =", round(bias, 2), "dB")) +
        ggtitle("Bland-Altman plot comparing MD across all stages of glaucoma"))

ggsave("plots/blandaltman-md-allgon.png", width = 7, height = 5)

milddf <- df %>%
  filter(classification == "Mild")

blandplotmild <- bland.altman.plot(milddf$retinalogik, milddf$hfa, graph.sys = "ggplot2", geom_count=T)
biasmild <- mean(milddf$retinalogik, na.rm=T) - mean(milddf$hfa, na.rm=T)

 print(blandplotmild +
        geom_smooth(method = "lm", se = FALSE) +
        # geom_point(position = "jitter") +
        geom_hline(yintercept = 0, color = "black") +
        geom_hline(yintercept = biasmild, color = "red", linetype = "solid", linewidth = 2) +
        ylim(c(-15, 15)) +
        xlab("MD") +
        ylab("Difference in MD (Retinalogik-HFA)") +
        labs(subtitle=paste("Bias =", round(biasmild, 2), "dB")) +
        ggtitle("Bland-Altman plot comparing MD in patients with mild glaucoma"))

ggsave("plots/blandaltman-md-mild.png", width = 7, height = 5)

moderatedf <- df %>%
  filter(classification == "Moderate")

blandplotmoderate <- bland.altman.plot(moderatedf$retinalogik, moderatedf$hfa, graph.sys = "ggplot2", geom_count=T)
biasmoderate <- mean(moderatedf$retinalogik, na.rm=T) - mean(moderatedf$hfa, na.rm=T)

print(blandplotmoderate +
        geom_smooth(method = "lm", se = FALSE) +
        # geom_point(position = "jitter") +
        geom_hline(yintercept = 0, color = "black") +
        geom_hline(yintercept = biasmoderate, color = "red", linetype = "solid", linewidth = 2) +
        ylim(c(-15, 15)) +
        xlab("MD") +
        ylab("Difference in MD (Retinalogik-HFA)") +
        labs(subtitle=paste("Bias =", round(biasmoderate, 2), "dB")) +
        ggtitle("Bland-Altman plot comparing MD in patients with moderate glaucoma"))

ggsave("plots/blandaltman-md-moderate.png", width = 7, height = 5)

advanceddf <- df %>%
  filter(classification == "Advanced")

blandplotadvanced <- bland.altman.plot(advanceddf$retinalogik, advanceddf$hfa, graph.sys = "ggplot2", geom_count=T)
biasadvanced <- mean(advanceddf$retinalogik, na.rm=T) - mean(advanceddf$hfa, na.rm=T)

print(blandplotadvanced +
        geom_smooth(method = "lm", se = FALSE) +
        # geom_point(position = "jitter") +
        geom_hline(yintercept = 0, color = "black") +
        geom_hline(yintercept = biasadvanced, color = "red", linetype = "solid", linewidth = 2) +
        ylim(c(-15, 15)) +
        xlab("MD") +
        ylab("Difference in MD (Retinalogik-HFA)") +
        labs(subtitle=paste("Bias =", round(biasadvanced, 2), "dB")) +
        ggtitle("Bland-Altman plot comparing MD in patients with advanced glaucoma"))

ggsave("plots/blandaltman-md-advanced.png", width = 7, height = 5)

```

### Pointwise analysis
```{r pointwise}

## Early GON
nRT <- dBdat_filtered %>%
  filter(classification == "Mild" & device == "retinalogik") %>%
  distinct(id, device, visit, eye) %>%
  n_distinct()

nHFA <- dBdat_filtered %>%
  filter(classification == "Mild" & device == "hfa") %>%
  distinct(id, device, visit, eye) %>%
  n_distinct()

meanSensitivity <- dBdat_filtered %>%
  filter(classification == "Mild") %>%
  distinct(id, device, visit, eye, x, y, dB) %>%
  pivot_wider(names_from = device, values_from = dB) %>%
  group_by(x, y) %>%
  summarize(meanRT = mean(retinalogik, na.rm=T), sdRT = sd(retinalogik, na.rm=T),
            meanHFA = mean(hfa, na.rm=T), sdHFA = sd(hfa, na.rm=T),
            pvalue = wilcox.test(Pair(retinalogik, hfa) ~1)$p.value) %>%
  ungroup() %>%
  add_row(tibble_row(x = 15, y = -3, meanRT = NA, sdRT = NA)) %>%
  add_row(tibble_row(x = 15, y = +3, meanRT = NA, sdRT = NA)) 

rtplot_earlyGON <- meanSensitivity %>%
  ggplot(aes(x, y, meanRT)) +
  geom_raster(aes(x = x, y = y, fill = meanRT)) +
  geom_text(aes(label = ifelse(is.na(meanRT), "", paste0(round(meanRT, 2), "\n(", round(sdRT, 2), ")")),
                x = x, y = y), size = 3, colour="white") +
  coord_fixed(ratio = 1) +
  scale_fill_gradientn(colours = viridis(51), limits = c(-1, 49),
                       na.value="darkred") +
  theme_bw() +
  ggtitle("Retinalogik", subtitle = paste0("(n = ", nRT, " early GON eyes)")) +
  theme(legend.position="none")

hfaplot_earlyGON <- meanSensitivity %>%
  ggplot(aes(x, y, meanHFA)) +
  geom_raster(aes(x = x, y = y, fill = meanHFA)) +
  geom_text(aes(label = ifelse(is.na(meanHFA), "", paste0(round(meanHFA, 2), "\n(", round(sdHFA, 2), ")")),
                x = x, y = y), size = 3, colour="white") +
  coord_fixed(ratio = 1) +
  scale_fill_gradientn(colours = viridis(51), limits = c(-1, 49),
                       na.value="darkred") +
  theme_bw() +
  ggtitle("HFA", subtitle = paste0("(n = ", nHFA, " early GON eyes)")) +
  theme(legend.position="none")

diffplot_earlyGON <- meanSensitivity %>%
  mutate(diff = meanRT-meanHFA) %>%
  ggplot(aes(x, y, diff)) +
  geom_raster(aes(x = x, y = y, fill = diff)) +
  geom_text(aes(label = ifelse(is.na(diff), "", paste0(round(diff, 2))),
                x = x, y = y), size = 3, colour="black") +
  coord_fixed(ratio = 1) +
  scale_fill_gradientn(colours = colorRampPalette(c("red", "white", "red"))(27), limits = c(-13, 13), 
                       na.value="darkred") +
  theme_bw() +
  ggtitle("Difference (Retinalogik-HFA)") +
  theme(legend.position="none")

pplot_earlyGON <- meanSensitivity %>%
  ggplot(aes(x, y)) +
  geom_tile(fill = "white", colour = "grey90") +
  geom_text(aes(label = ifelse(pvalue < 0.01, "<.01", round(pvalue, 2)),
                x = x, y = y),  size = 3, colour="black") +
  coord_fixed(ratio = 1) +
  #scale_fill_gradient(low = "gray", high = "white", na.value="darkred") +
  theme_bw() +
  ggtitle("Wilcoxon p-values (Retinalogik-HFA)") +
  theme(legend.position="none",
        panel.grid = element_blank())

## Moderate GON
nRT <- dBdat_filtered %>%
  filter(classification == "Moderate" & device == "retinalogik") %>%
  distinct(id, device, visit, eye) %>%
  n_distinct()

nHFA <- dBdat_filtered %>%
  filter(classification == "Moderate" & device == "hfa") %>%
  distinct(id, device, visit, eye) %>%
  n_distinct()

meanSensitivity <- dBdat_filtered %>%
  filter(classification == "Moderate") %>%
  distinct(id, device, visit, eye, x, y, dB) %>%
  pivot_wider(names_from = device, values_from = dB) %>%
  group_by(x, y) %>%
  summarize(meanRT = mean(retinalogik, na.rm=T), sdRT = sd(retinalogik, na.rm=T),
            meanHFA = mean(hfa, na.rm=T), sdHFA = sd(hfa, na.rm=T),
            pvalue = wilcox.test(Pair(retinalogik, hfa) ~1)$p.value) %>%
  ungroup() %>%
  add_row(tibble_row(x = 15, y = -3, meanRT = NA, sdRT = NA)) %>%
  add_row(tibble_row(x = 15, y = +3, meanRT = NA, sdRT = NA)) 

rtplot_moderateGON <- meanSensitivity %>%
  ggplot(aes(x, y, meanRT)) +
  geom_raster(aes(x = x, y = y, fill = meanRT)) +
  geom_text(aes(label = ifelse(is.na(meanRT), "", paste0(round(meanRT, 2), "\n(", round(sdRT, 2), ")")),
                x = x, y = y), size = 3, colour="white") +
  coord_fixed(ratio = 1) +
  scale_fill_gradientn(colours = viridis(51), limits = c(-1, 49),
                       na.value="darkred") +
  theme_bw() +
  ggtitle("Retinalogik", subtitle = paste0("(n = ", nRT, " moderate GON eyes)")) +
  theme(legend.position="none")

hfaplot_moderateGON <- meanSensitivity %>%
  ggplot(aes(x, y, meanHFA)) +
  geom_raster(aes(x = x, y = y, fill = meanHFA)) +
  geom_text(aes(label = ifelse(is.na(meanHFA), "", paste0(round(meanHFA, 2), "\n(", round(sdHFA, 2), ")")),
                x = x, y = y), size = 3, colour="white") +
  coord_fixed(ratio = 1) +
  scale_fill_gradientn(colours = viridis(51), limits = c(-1, 49),
                       na.value="darkred") +
  theme_bw() +
  ggtitle("HFA", subtitle = paste0("(n = ", nHFA, " moderate GON eyes)")) +
  theme(legend.position="none")

diffplot_moderateGON <- meanSensitivity %>%
  mutate(diff = meanRT-meanHFA) %>%
  ggplot(aes(x, y, diff)) +
  geom_raster(aes(x = x, y = y, fill = diff)) +
  geom_text(aes(label = ifelse(is.na(diff), "", paste0(round(diff, 2))),
                x = x, y = y), size = 3, colour="black") +
  coord_fixed(ratio = 1) +
  scale_fill_gradientn(colours = colorRampPalette(c("red", "white", "red"))(27), limits = c(-13, 13), 
                       na.value="darkred") +
  theme_bw() +
  ggtitle("Difference (Retinalogik-HFA)") +
  theme(legend.position="none")

pplot_moderateGON <- meanSensitivity %>%
  ggplot(aes(x, y)) +
  geom_tile(fill = "white", colour = "grey90") +
  geom_text(aes(label = ifelse(pvalue < 0.01, "<.01", round(pvalue, 2)),
                x = x, y = y),  size = 3, colour="black") +
  coord_fixed(ratio = 1) +
  #scale_fill_gradient(low = "gray", high = "white", na.value="darkred") +
  theme_bw() +
  ggtitle("Wilcoxon p-values (Retinalogik-HFA)") +
  theme(legend.position="none",
        panel.grid = element_blank())

## Advanced GON
nRT <- dBdat_filtered %>%
  filter(classification == "Advanced" & device == "retinalogik") %>%
  distinct(id, device, visit, eye) %>%
  n_distinct()

nHFA <- dBdat_filtered %>%
  filter(classification == "Advanced" & device == "hfa") %>%
  distinct(id, device, visit, eye) %>%
  n_distinct()

meanSensitivity <- dBdat_filtered %>%
  filter(classification == "Advanced") %>%
  distinct(id, device, visit, eye, x, y, dB) %>%
  pivot_wider(names_from = device, values_from = dB) %>%
  group_by(x, y) %>%
  summarize(meanRT = mean(retinalogik, na.rm=T), sdRT = sd(retinalogik, na.rm=T),
            meanHFA = mean(hfa, na.rm=T), sdHFA = sd(hfa, na.rm=T),
            pvalue = wilcox.test(Pair(retinalogik, hfa) ~1)$p.value) %>%
  ungroup() %>%
  add_row(tibble_row(x = 15, y = -3, meanRT = NA, sdRT = NA)) %>%
  add_row(tibble_row(x = 15, y = +3, meanRT = NA, sdRT = NA)) 

rtplot_advancedGON <- meanSensitivity %>%
  ggplot(aes(x, y, meanRT)) +
  geom_raster(aes(x = x, y = y, fill = meanRT)) +
  geom_text(aes(label = ifelse(is.na(meanRT), "", paste0(round(meanRT, 2), "\n(", round(sdRT, 2), ")")),
                x = x, y = y), size = 3, colour="white") +
  coord_fixed(ratio = 1) +
  scale_fill_gradientn(colours = viridis(51), limits = c(-1, 49),
                       na.value="darkred") +
  theme_bw() +
  ggtitle("Retinalogik", subtitle = paste0("(n = ", nRT, " advanced GON eyes)")) +
  theme(legend.position="none")

hfaplot_advancedGON <- meanSensitivity %>%
  ggplot(aes(x, y, meanHFA)) +
  geom_raster(aes(x = x, y = y, fill = meanHFA)) +
  geom_text(aes(label = ifelse(is.na(meanHFA), "", paste0(round(meanHFA, 2), "\n(", round(sdHFA, 2), ")")),
                x = x, y = y), size = 3, colour="white") +
  coord_fixed(ratio = 1) +
  scale_fill_gradientn(colours = viridis(51), limits = c(-1, 49),
                       na.value="darkred") +
  theme_bw() +
  ggtitle("HFA", subtitle = paste0("(n = ", nHFA, " advanced GON eyes)")) +
  theme(legend.position="none")

diffplot_advancedGON <- meanSensitivity %>%
  mutate(diff = meanRT-meanHFA) %>%
  ggplot(aes(x, y, diff)) +
  geom_raster(aes(x = x, y = y, fill = diff)) +
  geom_text(aes(label = ifelse(is.na(diff), "", paste0(round(diff, 2))),
                x = x, y = y), size = 3, colour="black") +
  coord_fixed(ratio = 1) +
  scale_fill_gradientn(colours = colorRampPalette(c("red", "white", "red"))(27), limits = c(-13, 13), 
                       na.value="darkred") +
  theme_bw() +
  ggtitle("Difference (Retinalogik-HFA)") +
  theme(legend.position="none")

pplot_advancedGON <- meanSensitivity %>%
  ggplot(aes(x, y)) +
  geom_tile(fill = "white", colour = "grey90") +
  geom_text(aes(label = ifelse(pvalue < 0.01, "<.01", round(pvalue, 2)),
                x = x, y = y),  size = 3, colour="black") +
  coord_fixed(ratio = 1) +
  scale_fill_gradient(low = "gray", high = "white", na.value="darkred") +
  theme_bw() +
  ggtitle("Wilcoxon p-values (Retinalogik-HFA)") +
  theme(legend.position="none",
        panel.grid = element_blank())

# rtplot_earlyGON + hfaplot_earlyGON
# ggsave("plots/earlygon.png", width = 10, height = 6)
# 
# rtplot_moderateGON + hfaplot_moderateGON
# ggsave("plots/moderategon.png", width = 10, height = 6)
# 
# rtplot_advancedGON + hfaplot_advancedGON
# ggsave("plots/advancedgon.png", width = 10, height = 6)
# 
# rtplot_earlyGON + hfaplot_earlyGON + diffplot_earlyGON
# ggsave("plots/earlygon_withdiff.png", width = 14, height = 6)
# 
# rtplot_moderateGON + hfaplot_moderateGON + diffplot_moderateGON
# ggsave("plots/moderategon_withdiff.png", width = 14, height = 6)
# 
# rtplot_advancedGON + hfaplot_advancedGON + diffplot_advancedGON
# ggsave("plots/advancedgon_withdiff.png", width = 14, height = 6)

# rtplot_earlyGON + hfaplot_earlyGON + pplot_earlyGON
# ggsave("plots/earlygon_withp.png", width = 14, height = 6)
# 
# rtplot_moderateGON + hfaplot_moderateGON + pplot_moderateGON
# ggsave("plots/moderategon_withp.png", width = 14, height = 6)
# 
# rtplot_advancedGON + hfaplot_advancedGON + pplot_advancedGON
# ggsave("plots/advancedgon_withp.png", width = 14, height = 6)

rtplot_earlyGON + hfaplot_earlyGON
diffplot_earlyGON+ pplot_earlyGON

rtplot_moderateGON + hfaplot_moderateGON
diffplot_moderateGON + pplot_moderateGON

rtplot_advancedGON + hfaplot_advancedGON
diffplot_advancedGON + pplot_advancedGON
```

### Scotoma analysis
Please refer to [link](https://vvnejl.github.io/retinalogik-study/scotoma_analysis.html).

### Test duration
>The duration of the RL-VF was significantly shorter than that of the Humphrey and did not vary with glaucoma severity in a similar manner to SAP administered with the HFA.

```{r duration, warning=FALSE}
duration <- dBdat_filtered %>%
  group_by(classification, device) %>%
  mutate(duration = as.duration(duration)) %>%
  mutate(mean = mean(duration)/60,
         sd = sd(duration)/60,
         median = median(duration)/60,
         iqr_q1 = quantile(duration, .25)/60, 
         iqr_q3 = quantile(duration, .75)/60) %>%
  distinct(classification, device, mean, sd, median, iqr_q1, iqr_q3) 

duration %>%
  mutate(classification = factor(classification, levels = c("Mild", "Moderate", "Advanced")),
         device = factor(device, levels = c("retinalogik", "hfa"))) %>%
  arrange(classification, device) %>%
  nice_table()

durationdat <- dBdat_filtered %>%
  group_by(id, device) %>%
  mutate(duration = mean(as.duration(duration))) %>%
  distinct(id, classification, device, duration) %>%
  pivot_wider(names_from = device, values_from = duration) %>%
  na.omit()

durationtest <- wilcox.test(durationdat$retinalogik, durationdat$hfa, paired=TRUE)

```

Median test duration time was `r round(median(durationdat$retinalogik)/60, 2)` (IQR = `r round(quantile(durationdat$retinalogik, .25)/60, 2)`, `r round(quantile(durationdat$retinalogik, .75)/60, 2)`) mins for Retinalogik and `r round(median(durationdat$hfa)/60, 2)` (IQR = `r round(quantile(durationdat$hfa, .25)/60, 2)`, `r round(quantile(durationdat$hfa, .75)/60, 2)`) mins for HFA. Wilcoxon signed rank test indicated that this `r round((median(durationdat$hfa)-median(durationdat$retinalogik))/60, 2)` mins difference was statistically significant, *V* = `r durationtest$statistic`, `r if (durationtest$p.value < 0.01) {"$p < 0.01$"} else {paste("$p = ", round(durationtest$p.value, 3), "$", sep="")}`.

### ICC

```{r icc}
# ICC (1,3) Agreement between tests comparing "retinalogik" and "hfa" measurements.
dB_wide_agreement <- dBdat_filtered %>%
  group_by(id, device) %>%
  summarise(mean_dB = mean(dB, na.rm = TRUE)) %>%
  pivot_wider(names_from = device, values_from = mean_dB)

icc_data_agreement <- dB_wide_agreement %>%
  ungroup() %>%
  select(retinalogik, hfa) %>%
  mutate(
    retinalogik = as.numeric(retinalogik),
    hfa = as.numeric(hfa)
  ) %>%
  as.matrix() 

#icc_result_1_3 <- ICC(icc_data_agreement, lmer = FALSE, missing=FALSE) # lmer=FALSE for base R stats
icc_result_1_3 <- ICC(icc_data_agreement, lmer = TRUE, missing=FALSE)
# use result from ICC3 Single_fixed_raters, e.g.,
# Single_fixed_raters      ICC3 0.93 27  32  32 1.4e-15        0.86        0.96

### For repeatability within a device
retinalogik_visits <- dBdat_filtered %>%
  filter(device == "retinalogik") %>%
  # If you have varying numbers of visits, you might need to pad with NA or take a consistent number.
  # For demonstration, let's assume we pivot directly.
  # It's crucial that 'visit' actually represents repeated measures of the *same thing* to be averaged.
  group_by(id, visit) %>%
  summarise(mean_dB_visit = mean(dB, na.rm = TRUE)) %>% # If multiple dB per id, visit, device
  pivot_wider(names_from = visit, values_from = mean_dB_visit, names_prefix = "Visit_") %>%
  ungroup() # Ungroup after pivot_wider

icc_data_retinalogik <- retinalogik_visits %>%
  select(starts_with("Visit_")) %>%
  as.matrix()

icc_result_1_2_retinalogik <- ICC(icc_data_retinalogik)
icc_result_1_2_retinalogik # use ICC2k results

# repeat for HFA
hfa_visits <- dBdat_filtered %>%
  filter(device == "hfa") %>%
  group_by(id, visit) %>%
  summarise(mean_dB_visit = mean(dB, na.rm = TRUE)) %>%
  pivot_wider(names_from = visit, values_from = mean_dB_visit, names_prefix = "Visit_") %>%
  ungroup()

icc_data_hfa <- hfa_visits %>%
  select(starts_with("Visit_")) %>%
  as.matrix()

icc_result_1_2_hfa <- ICC(icc_data_hfa)
icc_result_1_2_hfa # use ICC2k results
```

The ICC values for comparison between Retinalogik and HFA were `r round(icc_result_1_3$results[icc_result_1_3$results$type == "ICC3", "ICC"], 2)` (95% confidence interval [CI]: `r round(icc_result_1_3$results[icc_result_1_3$results$type == "ICC3", "lower bound"], 2)`-`r round(icc_result_1_3$results[icc_result_1_3$results$type == "ICC3", "upper bound"], 2)`). The ICCs for intra-device reproducibility were `r round(icc_result_1_2_retinalogik$results[icc_result_1_2_retinalogik$results$type == "ICC2k", "ICC"], 2)` (95% CI: `r round(icc_result_1_2_retinalogik$results[icc_result_1_2_retinalogik$results$type == "ICC2k", "lower bound"], 2)`-`r round(icc_result_1_2_retinalogik$results[icc_result_1_2_retinalogik$results$type == "ICC2k", "upper bound"], 2)`) for RetinaLogik, and `r round(icc_result_1_2_hfa$results[icc_result_1_2_hfa$results$type == "ICC2k", "ICC"], 2)` (95% CI: `r round(icc_result_1_2_hfa$results[icc_result_1_2_hfa$results$type == "ICC2k", "lower bound"], 2)`-`r round(icc_result_1_2_hfa$results[icc_result_1_2_hfa$results$type == "ICC2k", "upper bound"], 2)`) for the HFA. 

### Questionnaire

```{r questionnaire}
qdat <- read.csv("RetinaLogik Questionnaire Data(Sheet1).csv") %>%
  mutate(
    # Extract the numeric part after "RL"
    numeric_part = as.numeric(str_extract(Patient.ID, "\\d+")),
    # Format the numeric part to have at least two digits with leading zeros
    formatted_numeric = str_pad(numeric_part, width = 2, side = "left", pad = "0"),
    # Combine "RL" with the formatted number
    ID = paste0("RL", formatted_numeric)
  ) %>%
  select(-Patient.ID)

# Recode the columns 
qdat_numeric <- qdat %>%
  mutate(
    Q1..Desktop.difficulty = case_match(Q1..Desktop.difficulty,
      "very easy" ~ 5,
      "somewhat easy" ~ 4,
      "neutral" ~ 3,
      "somewhat difficult" ~ 2,
      "very difficult" ~ 1
    ),
    Q2..Head.mounted.difficulty = case_match(Q2..Head.mounted.difficulty,
      "very easy" ~ 5,
      "somewhat easy" ~ 4,
      "neutral" ~ 3,
      "somewhat difficult" ~ 2,
      "very difficult" ~ 1
    ),
    Q3..Desktop.comfort = case_match(Q3..Desktop.comfort,
      "very comfortable" ~ 5,
      "somewhat comfortable" ~ 4,
      "neutral" ~ 3,
      "somewhat uncomfortable" ~ 2,
      "very uncomfortable" ~ 1
    ),
    Q4..Head.mounted.comfort = case_match(Q4..Head.mounted.comfort,
      "very comfortable" ~ 5,
      "somewhat comfortable" ~ 4,
      "neutral" ~ 3,
      "somewhat uncomfortable" ~ 2,
      "very uncomfortable" ~ 1
    )
  )

summary_stats <- qdat_numeric %>%
  summarise(
    Med_Q1 = median(Q1..Desktop.difficulty, na.rm = TRUE),
    SD_Q1 = sd(Q1..Desktop.difficulty, na.rm = TRUE),
    Med_Q2 = median(Q2..Head.mounted.difficulty, na.rm = TRUE),
    SD_Q2 = sd(Q2..Head.mounted.difficulty, na.rm = TRUE),
    Med_Q3 = median(Q3..Desktop.comfort, na.rm = TRUE),
    SD_Q3 = sd(Q3..Desktop.comfort, na.rm = TRUE),
    Med_Q4 = median(Q4..Head.mounted.comfort, na.rm = TRUE),
    SD_Q4 = sd(Q4..Head.mounted.comfort, na.rm = TRUE)
  )

print(summary_stats)

patient_summary <- qdat_numeric %>%
  group_by(ID) %>%
  summarise(
    Med_Q1 = median(Q1..Desktop.difficulty, na.rm = TRUE),
    Med_Q2 = median(Q2..Head.mounted.difficulty, na.rm = TRUE),
    Med_Q3 = median(Q3..Desktop.comfort, na.rm = TRUE),
    Med_Q4 = median(Q4..Head.mounted.comfort, na.rm = TRUE)
  )

print(patient_summary)

questionnaire_plot <- summary_stats %>%
  tidyr::pivot_longer(
    cols = c(Med_Q1, Med_Q2, Med_Q3, Med_Q4),
    names_to = "Question",
    values_to = "Median_Score"
  )

questionnaire_plot %>%
  ggplot(aes(x = Question, y = Median_Score)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  labs(
    title = "Median Scores for each Question",
    x = "Questions",
    y = "Median Score (1-5 Likert Scale)",
    caption = "On the difficulty scale: 1 is very difficult, 5 is very easy;
    On the comfort scale: 1 is very uncomfortable, 5 is very comfortable."
  ) +
  scale_x_discrete(labels = c("HFA difficulty", "Retinalogik difficulty",
                              "HFA comfort", "Retinalogik comfort")) +
  theme_minimal()

# Wilcoxon signed-rank test
wilcox_difficulty <- wilcox.test(qdat_numeric$Q1..Desktop.difficulty, 
                                 qdat_numeric$Q2..Head.mounted.difficulty, 
                                 paired = TRUE)

wilcox_comfort <- wilcox.test(qdat_numeric$Q3..Desktop.comfort, 
                              qdat_numeric$Q4..Head.mounted.comfort, 
                              paired = TRUE)

q5dat <- qdat %>%
  select(ID, Q5..Overall.preference) %>%
  mutate(Q5..Overall.preference = as.factor(Q5..Overall.preference)) 

q5dat <- q5dat[q5dat$Q5..Overall.preference != "", ] %>%
  count(Q5..Overall.preference)

```
**Q5 is "Which test would you prefer to perform on an on-going basis?"**

`r round(q5dat$n[1]/sum(q5dat$n)*100, 2)`% prefer "desktop" aka HFA, `r round(q5dat$n[2]/sum(q5dat$n)*100, 2)`% prefer head-mounted aka Retinalogik, and `r round(q5dat$n[3]/sum(q5dat$n)*100, 2)`% has no preference for either.

