---
title: "Retinalogik vs HFA study JAMA Ophthalmology paper"
author: "[Eadie Technologies, Inc.](https://www.eadietech.com/)"
date: "2025-06-11"
output: 
  html_document:
    theme: yeti
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: 
      collapsed: true
    includes:
        after_body: footer.html
---

## Setup

### Versioning

Last updated on `r format(Sys.time(), '%Y-%b-%d')` at `r format(Sys.time(), '%I:%M %p')`.

* 2025-Jun-11: Created separate .Rmd file as index.Rmd was getting very cluttered.
* 2025-Jun-26: Extract data based on classification during screening.
* 2025-Aug-25: Extract, clean, and run stats on questionnaire data.
* 2026-Jan-16: Added difference plots, p-value difference plots, LMM etc.
* 2026-Apr-21: Fixed left/eye swap discrepancies and added PSD.

### Acknowledgements
Data collected by Research Assistants Luc St-Pierre, Reann Post, and Research Coordinator Lynn Murphy.

```{r setup, collapse=TRUE}

# load packages
library(tidyverse)
library(here)
library(DT)
library(plotly)
library(viridis)
library(svglite)
library(htmltools)
library(gridExtra)
library(sf)
library(spdep)
library(RColorBrewer)
library(patchwork)
library(report)
library(effectsize)
library(rempsyc)
library(psych)
library(BlandAltmanLeh)

sessionInfo()

# default chunk options
knitr::opts_chunk$set(
  comment = '>', collapse = TRUE, cache = FALSE, dev= c("png")
  )

rm(list = ls()) 

# load processed data
load(here("dBdat.Rda"))

load(here("retinalogik_hfadatawpsd.Rda"))

classification <- readxl::read_excel("RetinaLogik Data Classification.xlsx") %>%
  mutate(
    # Extract the numeric part after "RL"
    numeric_part = as.numeric(str_extract(ID, "\\d+")),
    # Format the numeric part to have at least two digits with leading zeros
    formatted_numeric = str_pad(numeric_part, width = 2, side = "left", pad = "0"),
    # Combine "RL" with the formatted number
    ID = paste0("RL", formatted_numeric)
  ) %>%
  # Remove the temporary columns
  select(-numeric_part, -formatted_numeric) %>%
  rename(id = ID, classification = Classifcation, eye = `Eye (OD/OS)`) %>%
  mutate(eye = case_when(eye == "OS" ~ "L",
                         eye == "OD" ~ "R",
                         eye == "OU" ~ "R"))

dBdat_filtered <- dBdat %>%
  ungroup() %>%
  mutate(x = as.integer(x), y = as.integer(y)) %>%
  mutate(x = case_when(
    device == "retinalogik" & eye == "L" ~ x*-1,
    TRUE ~ as.integer(x)
  )) %>%
  # remove blind spots
  filter(!(x == 15 & y == -3)) %>%
  filter(!(x == 15 & y == +3)) %>%
  #remove first 2 visits
  #filter(visit != 1 & visit != 2) %>%
  #remove reliability indices <30%
  #filter(flPerc < 30 & fnPerc < 30 & fpPerc < 30) %>%
  group_by(id, device, eye) %>%
  mutate(meanmd = mean(md))

dBdat_filtered <- merge(dBdat_filtered, classification) %>%
  filter(classification != "Control") %>%
  mutate(classification = factor(classification, levels = c("Mild", "Moderate", "Advanced")))

```
### Download plots
[click me to find all generated plots](https://github.com/vvnejl/retinalogik-study/tree/cc84aa781bde9225ff235dc1ab89a7aad0ded830/plots)

# Study summary
We compared 24-2 visual field (VF) test results between Retinalogik and the Humphrey Field Analyzer (HFA) using SITA Standard test strategy in patients with early glaucoma and moderate to advanced glaucoma [here](https://vvnejl.github.io/retinalogik-study/). A scotoma analysis was performed following glaucomatous VF definitions found in [Park et al. (2011)](https://www.aaojournal.org/article/S0161-6420(11)00172-2/abstract). Scotoma analysis results may be found [here](https://vvnejl.github.io/retinalogik-study/scotoma_analysis.html).

# Methods (very briefly)
## Subjects
Eligible participants were identified among patients of Dr. Brennan Eadie at the Eadie Eye Centre. If deemed eligible for the study, subjects were recruited consecutively.

Each participant underwent five study visits. At each visit, they performed a VF test on both eyes using the two devices. The order of device tested was randomized at the first (baseline) study visit.

Exclusion criteria were:

* age <18 years
* diagnosis of secondary glaucoma
* diagnosis of non-glaucomatous optic neuropathy
* significant media opacity
* previous intraocular surgery
* pregnancy
* seizure disorder
* cardiac pacemaker or other implantable device
* severe vertigo or balance disturbance
* inability to demonstrate competence to make an informed decision regarding study participation

The study adhered to the tenets of the Declaration of Helsinki for research involving human subjects and the protocol was approved by the Nova Scotia Health Research Ethics Board (#1030608). All participants gave their written informed consent before enrollment in the study.

### Demographics
```{r demographics}
vft <- dBdat_filtered %>%
  distinct(id, visit, eye) %>%
  count(eye)

gender <- dBdat_filtered %>%
  distinct(id, gender) %>%
  count(gender)

groups <- dBdat_filtered %>%
  distinct(id, classification) %>%
  mutate(classification = as.factor(classification)) %>%
  count(classification)

```

All OS data are transposed to OD format before analyses. `r sum(vft$n)` VF tests from `r length(unique(dBdat_filtered$id))` participants (`r sum(gender$n[gender$gender == "F"], na.rm = T)` females, `r sum(gender$n[gender$gender == "M"], na.rm = T)` males) aged `r min(dBdat_filtered$age)` to `r max(dBdat_filtered$age)` (*M* = `r round(mean(dBdat_filtered$age), 2)`, *SD* = `r round(sd(dBdat_filtered$age), 2)`) were included in the analysis (Early n = `r groups$n[groups$classification == "Mild"]`; Moderate n = `r groups$n[groups$classification == "Moderate"]`; Advanced n = `r groups$n[groups$classification == "Advanced"]`).

# Results
## Analysis
Of the 5 visits, the first 2 visits were removed from analysis. VFTs where reliability indices (FL, FP, and FN) > 30% were removed from analysis as well.

### Mean/median sensitivity
In patients with advanced glaucoma, the median sensitivity was significantly higher when using the VR-RL perimeter. Wilcoxon signed-rank tests confirmed this difference when looking at the median of the patient-level mean differences between visits.

```{r meansens}
dBdat_filtered <- dBdat_filtered %>%
  #remove first 2 visits
  filter(visit != 1 & visit != 2) %>%
  #remove reliability indices <30%
  filter(flPerc < 30 & fnPerc < 30 & fpPerc < 30)

dBdat_filtered %>%
  group_by(id, device) %>%
  mutate(mean = mean(dB)) %>%
  distinct(id, classification, device, mean) %>%
  pivot_wider(names_from = device, values_from = mean) %>%
  na.omit() %>%
  mutate(classification = factor(classification, levels = c("Mild", "Moderate", "Advanced"))) %>%
  group_by(classification) %>%
  summarise(
    n = n(),
    mean_retinalogik = mean(retinalogik, na.rm = TRUE),
    sd_retinalogik = sd(retinalogik, na.rm = TRUE),
    mean_hfa = mean(hfa, na.rm = TRUE),
    sd_hfa = sd(hfa, na.rm = TRUE)
  ) %>%
  mutate(difference = mean_retinalogik-mean_hfa) %>%
  nice_table()

#########
dBdat_medians <- dBdat_filtered %>%
  group_by(id, device) %>%
  mutate(median = median(dB)) %>%
  distinct(id, classification, device, median) %>%
  pivot_wider(names_from = device, values_from = median) %>%
  na.omit() %>%
  mutate(difference = retinalogik-hfa)

# 1. Group the cleaned, wide data by classification
wilcoxon_results <- dBdat_medians %>%
  group_by(classification) %>%
  
  # 2. Nest the data for group-wise application
  nest() %>%
  
  # 3. Apply the Wilcoxon test to the paired columns in each nested group
  mutate(
    wilcox_test = map(data, ~ wilcox.test(.x$retinalogik, .x$hfa, paired = TRUE)),
    
    # 4. Extract the key results cleanly
    W_statistic = map_dbl(wilcox_test, "statistic"),
    p_value = map_dbl(wilcox_test, "p.value")
  ) %>%
  
  # 5. Select the final output columns
  select(classification, W_statistic, p_value)

wilcoxon_results %>%
  mutate(classification = factor(classification, levels = c("Mild", "Moderate", "Advanced"))) %>%
  arrange(classification) %>%
  nice_table(title = c("Wilcoxon signed rank test on paired samples"))

dBdat_filtered %>%
  group_by(id, device) %>%
  mutate(median = median(dB)) %>%
  distinct(id, classification, device, median) %>%
  pivot_wider(names_from = device, values_from = median) %>%
  na.omit() %>%
  mutate(classification = factor(classification, levels = c("Mild", "Moderate", "Advanced"))) %>%
  group_by(classification) %>%
  summarise(
    n = n(),
    median_retinalogik = median(retinalogik, na.rm = TRUE),
    q1_retinalogik = quantile(retinalogik, .25, na.rm = TRUE),
    q3_retinalogik = quantile(retinalogik, .75, na.rm = TRUE),
    iqr_retinalogik = IQR(retinalogik, na.rm = TRUE),
    median_hfa = median(hfa, na.rm = TRUE),
    q1_hfa = quantile(hfa, .25, na.rm = TRUE),
    q3_hfa = quantile(hfa, .75, na.rm = TRUE),
    iqr_hfa = IQR(hfa, na.rm = TRUE)
  ) %>%
  nice_table()

# mild_data <- subset(dBdat_means, classification == "Mild")
# t.test(mild_data$retinalogik, mild_data$hfa, paired = TRUE) %>%
#   report()
# 
# moderate_data <- subset(dBdat_means, classification == "Moderate")
# t.test(moderate_data$retinalogik, moderate_data$hfa, paired = TRUE) %>%
#   report()
# 
# advanced_data <- subset(dBdat_means, classification == "Advanced")
# t.test(advanced_data$retinalogik, advanced_data$hfa, paired = TRUE) %>%
#   report() %>% summary()
# 
# dBdat_medians <- dBdat_filtered %>%
#   group_by(id, device) %>%
#   mutate(median = median(dB)) %>%
#   distinct(id, classification, device, median) %>%
#   pivot_wider(names_from = device, values_from = median) %>%
#   na.omit()
# 
# mild_tmp <- dBdat_filtered %>%
#   subset(classification == "Mild") %>%
#   pivot_wider(names_from = device, values_from = dB) %>%
#   na.omit()
# wilcox.test(mild_tmp$retinalogik, mild_tmp$hfa, paired=TRUE)
# 
# mild_data <- subset(dBdat_medians, classification == "Mild")
# wilcox.test(mild_data$retinalogik, mild_data$hfa, paired = TRUE)
# 
# moderate_data <- subset(dBdat_medians, classification == "Moderate")
# wilcox.test(moderate_data$retinalogik, moderate_data$hfa, paired = TRUE)
# 
# advanced_data <- subset(dBdat_medians, classification == "Advanced")
# wilcox.test(advanced_data$retinalogik, advanced_data$hfa, paired = TRUE)

```

```{r hexbin}
dBdat_filtered %>%
  group_by(id, device, x, y) %>%
  mutate(meandB = mean(dB)) %>%
  distinct(id, device, x, y, meandB) %>%
  pivot_wider(names_from = device, values_from = meandB) %>%
  mutate(retinalogik = as.numeric(retinalogik), hfa = as.numeric(hfa)) %>%
  ggplot(aes(x = hfa, y = retinalogik)) +
  geom_hex(bins = 30) +
  scale_fill_viridis_c() +
  geom_smooth(method = "loess", se = TRUE, colour = "red") +
  geom_abline(intercept = 0, slope = 1) +
  labs(title="Hexbin plot of threshold sensitivities for all locations",
       subtitle="Averaged across reliable datasets between visits 3 to 5\n(Size of bin = 30)") +
  xlab("HFA (dB)") +
  ylab("VR device (dB)") +
  xlim(-2, 40) +
  ylim(-2, 40) +
  theme_bw()

ggsave("plots/hexbin30.png", width = 5, height = 4)
```
```{r dbblandaltman}
df <- dBdat_filtered %>%
  group_by(id, device, x, y) %>%
  mutate(dB = mean(dB)) %>%
  distinct(id, device, x, y, dB) %>%
  pivot_wider(names_from = device, values_from = dB) %>%
  mutate(retinalogik = as.numeric(retinalogik), hfa = as.numeric(hfa))

blandplot <- bland.altman.plot(df$retinalogik, df$hfa, graph.sys = "ggplot2", geom_count=T)
bias <- mean(df$retinalogik, na.rm=T) - mean(df$hfa, na.rm=T)

print(blandplot +
        geom_smooth(method = "lm", se = FALSE) +
        # geom_point(position = "jitter") +
        geom_hline(yintercept = 0, color = "black") +
        geom_hline(yintercept = bias, color = "red", linetype = "solid", linewidth = 1) +
        #ylim(c(-10, 10)) +
        xlab("Threshold sensitivity (dB)") +
        ylab("Difference in dB (VR device-HFA)") +
        labs(subtitle=paste("Bias =", round(bias, 2), "dB")) +
        ggtitle("Bland-Altman plot comparing threshold sensitivity across all stages of glaucoma"))

ggsave("plots/blandaltman-dB-allgon.png", width = 7, height = 5)

```

```{r msblandaltman}
ms_means <- dBdat_filtered %>%
  group_by(id, device, visit) %>%
  summarize(ms = mean(dB)) %>%
  group_by(id, device) %>%
  summarize(meanms = mean(ms)) %>%
  pivot_wider(names_from = device, values_from = meanms) %>%
  na.omit() %>%
  mutate(difference = retinalogik-hfa)

df <- ms_means

blandplot <- bland.altman.plot(df$retinalogik, df$hfa, graph.sys = "ggplot2", geom_count=T)
bias <- mean(df$retinalogik, na.rm=T) - mean(df$hfa, na.rm=T)

print(blandplot +
        geom_smooth(method = "lm", se = FALSE) +
        # geom_point(position = "jitter") +
        geom_hline(yintercept = 0, color = "black") +
        geom_hline(yintercept = bias, color = "red", linetype = "solid", linewidth = 1) +
        ylim(c(-10, 10)) +
        xlab("Mean Sensitivity (MS)") +
        ylab("Difference in MS (VR device-HFA)") +
        labs(subtitle=paste("Bias =", round(bias, 2), "dB")) +
        ggtitle("Bland-Altman plot comparing MS across all stages of glaucoma"))

ggsave("plots/blandaltman-ms-allgon.png", width = 7, height = 5)

```

### Mean deviation
In patients with moderate and advanced glaucoma, the mean deviation was significantly higher when using the VR-RL perimeter. Wilcoxon signed-rank tests confirmed this difference when looking at the median of the patient-level mean differences between visits.

```{r mdtable}
dBdat_filtered %>%
  group_by(id, device) %>%
  mutate(mean = mean(md)) %>%
  distinct(id, classification, device, mean) %>%
  pivot_wider(names_from = device, values_from = mean) %>%
  na.omit() %>%
  mutate(classification = factor(classification, levels = c("Mild", "Moderate", "Advanced"))) %>%
  group_by(classification) %>%
  summarise(
    n = n(),
    mean_retinalogik = mean(retinalogik, na.rm = TRUE),
    sd_retinalogik = sd(retinalogik, na.rm = TRUE),
    mean_hfa = mean(hfa, na.rm = TRUE),
    sd_hfa = sd(hfa, na.rm = TRUE)
  ) %>%
  mutate(difference = mean_retinalogik-mean_hfa) %>%
  nice_table()

#########
md_means <- dBdat_filtered %>%
  group_by(id, device) %>%
  mutate(mean = mean(md)) %>%
  distinct(id, classification, device, mean) %>%
  pivot_wider(names_from = device, values_from = mean) %>%
  na.omit() %>%
  mutate(difference = retinalogik-hfa)

md_medians <- dBdat_filtered %>%
  group_by(id, device) %>%
  mutate(median = median(md)) %>%
  distinct(id, classification, device, median) %>%
  pivot_wider(names_from = device, values_from = median) %>%
  na.omit() %>%
  mutate(difference = retinalogik-hfa)

# 1. Group the cleaned, wide data by classification
wilcoxon_results <- md_medians %>%
  group_by(classification) %>%
  
  # 2. Nest the data for group-wise application
  nest() %>%
  
  # 3. Apply the Wilcoxon test to the paired columns in each nested group
  mutate(
    wilcox_test = map(data, ~ wilcox.test(.x$retinalogik, .x$hfa, paired = TRUE)),
    
    # 4. Extract the key results cleanly
    W_statistic = map_dbl(wilcox_test, "statistic"),
    p_value = map_dbl(wilcox_test, "p.value")
  ) %>%
  
  # 5. Select the final output columns
  select(classification, W_statistic, p_value)

wilcoxon_results %>%
  mutate(classification = factor(classification, levels = c("Mild", "Moderate", "Advanced"))) %>%
  arrange(classification) %>%
  nice_table(title = c("Wilcoxon signed rank test on paired samples"))

dBdat_filtered %>%
  group_by(id, device) %>%
  mutate(median = median(md)) %>%
  distinct(id, classification, device, median) %>%
  pivot_wider(names_from = device, values_from = median) %>%
  na.omit() %>%
  mutate(classification = factor(classification, levels = c("Mild", "Moderate", "Advanced"))) %>%
  group_by(classification) %>%
  summarise(
    n = n(),
    median_retinalogik = median(retinalogik, na.rm = TRUE),
    q1_retinalogik = quantile(retinalogik, .25, na.rm = TRUE),
    q3_retinalogik = quantile(retinalogik, .75, na.rm = TRUE),
    iqr_retinalogik = IQR(retinalogik, na.rm = TRUE),
    median_hfa = median(hfa, na.rm = TRUE),
    q1_hfa = quantile(hfa, .25, na.rm = TRUE),
    q3_hfa = quantile(hfa, .75, na.rm = TRUE),
    iqr_hfa = IQR(hfa, na.rm = TRUE)
  ) %>%
  nice_table()

```

```{r mdscatterplot}
dBdat_filtered %>%
  group_by(id, device, visit) %>%
  summarize(md = mean(md)) %>%
  group_by(id, device) %>%
  summarize(meanmd = mean(md)) %>%
  pivot_wider(names_from = device, values_from = meanmd) %>%
  mutate(retinalogik = as.numeric(retinalogik), hfa = as.numeric(hfa)) %>%
  ggplot(., aes(x=hfa, y=retinalogik)) + 
  geom_point(position = "jitter") +
  scale_x_continuous(breaks = seq(-30, 2, by=2), limits=c(-30,2)) +
  scale_y_continuous(breaks = seq(-30, 2, by=2), limits=c(-30,2)) +
  geom_abline(intercept = 0, slope = 1, color = "gray50") +
  labs(title="Mean deviation (MD)") +
  xlab("HFA (dB)") +
  ylab("VR device (dB)") +
  theme_bw()

ggsave("plots/mdscatter.png", width = 5, height = 4)
```

```{r mdblandaltman}
library(BlandAltmanLeh)

df <- md_means

blandplot <- bland.altman.plot(df$retinalogik, df$hfa, graph.sys = "ggplot2", geom_count=T)
bias <- mean(df$retinalogik, na.rm=T) - mean(df$hfa, na.rm=T)

print(blandplot +
        geom_smooth(method = "lm", se = FALSE) +
        # geom_point(position = "jitter") +
        geom_hline(yintercept = 0, color = "black") +
        geom_hline(yintercept = bias, color = "red", linetype = "solid", linewidth = 1) +
        ylim(c(-15, 15)) +
        xlab("Mean deviation (MD)") +
        ylab("Difference in MD (VR device-HFA)") +
        labs(subtitle=paste("Bias =", round(bias, 2), "dB")) +
        ggtitle("Bland-Altman plot comparing MD across all stages of glaucoma"))

ggsave("plots/blandaltman-md-allgon.png", width = 7, height = 5)

milddf <- df %>%
  filter(classification == "Mild")

blandplotmild <- bland.altman.plot(milddf$retinalogik, milddf$hfa, graph.sys = "ggplot2", geom_count=T)
biasmild <- mean(milddf$retinalogik, na.rm=T) - mean(milddf$hfa, na.rm=T)

 print(blandplotmild +
        geom_smooth(method = "lm", se = FALSE) +
        # geom_point(position = "jitter") +
        geom_hline(yintercept = 0, color = "black") +
        geom_hline(yintercept = biasmild, color = "red", linetype = "solid", linewidth = 2) +
        ylim(c(-15, 15)) +
        xlab("MD") +
        ylab("Difference in MD (Retinalogik-HFA)") +
        labs(subtitle=paste("Bias =", round(biasmild, 2), "dB")) +
        ggtitle("Bland-Altman plot comparing MD in patients with mild glaucoma"))

#ggsave("plots/blandaltman-md-mild.png", width = 7, height = 5)

moderatedf <- df %>%
  filter(classification == "Moderate")

blandplotmoderate <- bland.altman.plot(moderatedf$retinalogik, moderatedf$hfa, graph.sys = "ggplot2", geom_count=T)
biasmoderate <- mean(moderatedf$retinalogik, na.rm=T) - mean(moderatedf$hfa, na.rm=T)

print(blandplotmoderate +
        geom_smooth(method = "lm", se = FALSE) +
        # geom_point(position = "jitter") +
        geom_hline(yintercept = 0, color = "black") +
        geom_hline(yintercept = biasmoderate, color = "red", linetype = "solid", linewidth = 2) +
        ylim(c(-15, 15)) +
        xlab("MD") +
        ylab("Difference in MD (Retinalogik-HFA)") +
        labs(subtitle=paste("Bias =", round(biasmoderate, 2), "dB")) +
        ggtitle("Bland-Altman plot comparing MD in patients with moderate glaucoma"))

#ggsave("plots/blandaltman-md-moderate.png", width = 7, height = 5)

advanceddf <- df %>%
  filter(classification == "Advanced")

blandplotadvanced <- bland.altman.plot(advanceddf$retinalogik, advanceddf$hfa, graph.sys = "ggplot2", geom_count=T)
biasadvanced <- mean(advanceddf$retinalogik, na.rm=T) - mean(advanceddf$hfa, na.rm=T)

print(blandplotadvanced +
        geom_smooth(method = "lm", se = FALSE) +
        # geom_point(position = "jitter") +
        geom_hline(yintercept = 0, color = "black") +
        geom_hline(yintercept = biasadvanced, color = "red", linetype = "solid", linewidth = 2) +
        ylim(c(-15, 15)) +
        xlab("MD") +
        ylab("Difference in MD (Retinalogik-HFA)") +
        labs(subtitle=paste("Bias =", round(biasadvanced, 2), "dB")) +
        ggtitle("Bland-Altman plot comparing MD in patients with advanced glaucoma"))

#ggsave("plots/blandaltman-md-advanced.png", width = 7, height = 5)

```

### Pattern standard deviation
In patients with moderate and advanced glaucoma, the pattern standard deviation was significantly lower when using the VR-RL perimeter. Wilcoxon signed-rank tests confirmed this difference when looking at the median of the patient-level mean differences between visits.

```{r psdtable}
dBdat_filtered %>%
  group_by(id, device) %>%
  mutate(mean = mean(psd)) %>%
  distinct(id, classification, device, mean) %>%
  pivot_wider(names_from = device, values_from = mean) %>%
  na.omit() %>%
  mutate(classification = factor(classification, levels = c("Mild", "Moderate", "Advanced"))) %>%
  group_by(classification) %>%
  summarise(
    n = n(),
    mean_retinalogik = mean(retinalogik, na.rm = TRUE),
    sd_retinalogik = sd(retinalogik, na.rm = TRUE),
    mean_hfa = mean(hfa, na.rm = TRUE),
    sd_hfa = sd(hfa, na.rm = TRUE)
  ) %>%
  mutate(difference = mean_retinalogik-mean_hfa) %>%
  nice_table()

#########
psd_means <- dBdat_filtered %>%
  group_by(id, device) %>%
  mutate(mean = mean(psd)) %>%
  distinct(id, classification, device, mean) %>%
  pivot_wider(names_from = device, values_from = mean) %>%
  na.omit() %>%
  mutate(difference = retinalogik-hfa)

psd_medians <- dBdat_filtered %>%
  group_by(id, device) %>%
  mutate(median = median(psd)) %>%
  distinct(id, classification, device, median) %>%
  pivot_wider(names_from = device, values_from = median) %>%
  na.omit() %>%
  mutate(difference = retinalogik-hfa)

# 1. Group the cleaned, wide data by classification
wilcoxon_results <- psd_medians %>%
  group_by(classification) %>%
  
  # 2. Nest the data for group-wise application
  nest() %>%
  
  # 3. Apply the Wilcoxon test to the paired columns in each nested group
  mutate(
    wilcox_test = map(data, ~ wilcox.test(.x$retinalogik, .x$hfa, paired = TRUE)),
    
    # 4. Extract the key results cleanly
    W_statistic = map_dbl(wilcox_test, "statistic"),
    p_value = map_dbl(wilcox_test, "p.value")
  ) %>%
  
  # 5. Select the final output columns
  select(classification, W_statistic, p_value)

wilcoxon_results %>%
  mutate(classification = factor(classification, levels = c("Mild", "Moderate", "Advanced"))) %>%
  arrange(classification) %>%
  nice_table(title = c("Wilcoxon signed rank test on paired samples"))

dBdat_filtered %>%
  group_by(id, device) %>%
  mutate(median = median(psd)) %>%
  distinct(id, classification, device, median) %>%
  pivot_wider(names_from = device, values_from = median) %>%
  na.omit() %>%
  mutate(classification = factor(classification, levels = c("Mild", "Moderate", "Advanced"))) %>%
  group_by(classification) %>%
  summarise(
    n = n(),
    median_retinalogik = median(retinalogik, na.rm = TRUE),
    q1_retinalogik = quantile(retinalogik, .25, na.rm = TRUE),
    q3_retinalogik = quantile(retinalogik, .75, na.rm = TRUE),
    iqr_retinalogik = IQR(retinalogik, na.rm = TRUE),
    median_hfa = median(hfa, na.rm = TRUE),
    q1_hfa = quantile(hfa, .25, na.rm = TRUE),
    q3_hfa = quantile(hfa, .75, na.rm = TRUE),
    iqr_hfa = IQR(hfa, na.rm = TRUE)
  ) %>%
  nice_table()

```

```{r psdscatterplot}
dBdat_filtered %>%
  group_by(id, device, visit) %>%
  summarize(psd = mean(psd)) %>%
  group_by(id, device) %>%
  summarize(meanpsd = mean(psd)) %>%
  pivot_wider(names_from = device, values_from = meanpsd) %>%
  mutate(retinalogik = as.numeric(retinalogik), hfa = as.numeric(hfa)) %>%
  ggplot(., aes(x=hfa, y=retinalogik)) + 
  geom_point(position = "jitter") +
  scale_x_continuous(breaks = seq(0, 16, by=2), limits=c(0,16)) +
  scale_y_continuous(breaks = seq(0, 16, by=2), limits=c(0,16)) +
  geom_abline(intercept = 0, slope = 1, color = "gray50") +
  labs(title="Pattern standard deviation (PSD)") +
  xlab("HFA (dB)") +
  ylab("VR device (dB)") +
  theme_bw()

ggsave("plots/psdscatter.png", width = 5, height = 4)
```

```{r psdblandaltman}
df <- psd_means

blandplot <- bland.altman.plot(df$retinalogik, df$hfa, graph.sys = "ggplot2", geom_count=T)
bias <- mean(df$retinalogik, na.rm=T) - mean(df$hfa, na.rm=T)

print(blandplot +
        geom_smooth(method = "lm", se = FALSE) +
        # geom_point(position = "jitter") +
        geom_hline(yintercept = 0, color = "black") +
        geom_hline(yintercept = bias, color = "red", linetype = "solid", linewidth = 1) +
        ylim(c(-8, 8)) +
        xlab("Pattern standard deviation (PSD)") +
        ylab("Difference in PSD (VR device-HFA)") +
        labs(subtitle=paste("Bias =", round(bias, 2), "dB")) +
        ggtitle("Bland-Altman plot comparing PSD across all stages of glaucoma"))

ggsave("plots/blandaltman-psd-allgon.png", width = 7, height = 5)
```

### Pointwise analysis
```{r lmm}
# Calculate eccentricity for each test point
dBdat_filtered$eccentricity <- sqrt(dBdat_filtered$x^2 + dBdat_filtered$y^2)

library(lme4)
library(lmerTest) # Adds p-values to lmer output

# The model
lmm_model <- lmer(dB ~ device * classification + eccentricity + visit + 
                       (1 + visit | id), 
                       data = dBdat_filtered)

summary(lmm_model)

library(emmeans)

# Compare devices within each glaucoma classification
post_hoc <- emmeans(lmm_model, pairwise ~ device | classification)
print(post_hoc)

# Extracting results from emmeans
ph_df <- as.data.frame(post_hoc$contrasts)

# Helper function for p-value formatting
format_p <- function(p) {
  if (p < 0.001) {
    return("p < .001")
  } else {
    return(paste0("p = ", round(p, 3)))
  }
}

# Extracting specific estimates and p-values
adv_diff  <- round(abs(ph_df[ph_df$classification == "Advanced", "estimate"]), 2)
adv_p     <- format_p(ph_df[ph_df$classification == "Advanced", "p.value"])

mod_diff  <- round(abs(ph_df[ph_df$classification == "Moderate", "estimate"]), 2)
mod_p     <- format_p(ph_df[ph_df$classification == "Moderate", "p.value"])

mild_diff <- round(abs(ph_df[ph_df$classification == "Mild", "estimate"]), 2)
mild_p    <- format_p(ph_df[ph_df$classification == "Mild", "p.value"])

# Extracting eccentricity slope from the main model summary
res <- as.data.frame(summary(lmm_model)$coefficients)
ecc_slope <- round(res["eccentricity", "Estimate"], 2)
ecc_p     <- format_p(res["eccentricity", "Pr(>|t|)"])

library(kableExtra)

# 1. Tidy the Post-Hoc contrasts (The specific Device Differences)
post_hoc_df <- as.data.frame(post_hoc$contrasts) %>%
  select(classification, estimate, SE, p.value) %>%
  mutate(
    # We take the absolute value for 'Overestimation' and round
    Estimate = round(abs(estimate), 2),
    SE = round(SE, 2),
    # Formatting p-values for the table
    `P-value` = ifelse(p.value < 0.001, "< .001", as.character(round(p.value, 3)))
  ) %>%
  select(classification, Estimate, SE, `P-value`)

# 2. Display the table
kable(post_hoc_df, 
      col.names = c("Glaucoma Severity", "Mean Overestimation (dB)", "SE", "P-value"),
      caption = "Table: Adjusted Mean Differences (VR-RL minus HFA) by Glaucoma Severity derived from Linear Mixed-Effects Model.") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F)
```

Initial non-parametric comparisons using Wilcoxon signed-rank tests on patient-level averages showed significant differences in the Moderate and Advanced groups. To further validate these findings, we employed a Linear Mixed-Effects Model (LMM) to account for spatial eccentricity and inter-visit variability. The model included dB as the dependent variable, with device, glaucoma classification, and their interaction as fixed effects. Eccentricity and visit number were included as covariates to adjust for the "Hill of Vision" and potential learning effects, respectively. To account for the correlation within subjects across multiple visits and test points, a random intercept and random slope for visit were included for each patient id.

The LMM revealed a significant interaction between device type and disease severity (p < .001), indicating that the level of agreement between the VR-RL and HFA perimeters is dependent on the stage of glaucoma. In early-stage disease, the devices showed the highest level of agreement with a non-significant difference of only `r mild_diff` dB (`r mild_p`). In moderate cases, the VR-RL recorded significantly higher sensitivities by `r mod_diff` dB (`r mod_p`). In the advanced group, the VR-RL perimeter significantly overestimated retinal sensitivity by `r adv_diff` dB compared to the HFA (`r adv_p`). Other than disease severity, differences between VR-RL and HFA is also modulated by eccentricity, where for every 1 degree away from the center, sensitivity drops by `r ecc_slope` dB (`r ecc_p`).

The VR-RL device significantly overestimates retinal sensitivity in damaged fields compared to the HFA. In advanced stages, the VR-RL reports values `r adv_diff` dB higher than the HFA. This suggests a potential lack of sensitivity to dense scotomas. Even after adjusting for the Hill of Vision, the VR-RL fails to match the HFA’s deeper characterization of peripheral loss, potentially masking the true extent of glaucomatous damage.

```{r pointwise}

## Early GON
nRT <- dBdat_filtered %>%
  filter(classification == "Mild" & device == "retinalogik") %>%
  distinct(id, device, visit, eye) %>%
  n_distinct()

nHFA <- dBdat_filtered %>%
  filter(classification == "Mild" & device == "hfa") %>%
  distinct(id, device, visit, eye) %>%
  n_distinct()

meanSensitivity <- dBdat_filtered %>%
  filter(classification == "Mild") %>%
  distinct(id, device, visit, eye, x, y, dB) %>%
  pivot_wider(names_from = device, values_from = dB) %>%
  group_by(x, y) %>%
  summarize(meanRT = mean(retinalogik, na.rm=T), sdRT = sd(retinalogik, na.rm=T),
            meanHFA = mean(hfa, na.rm=T), sdHFA = sd(hfa, na.rm=T),
            # Add a check for complete pairs
            pvalue  = if(sum(!is.na(retinalogik) & !is.na(hfa)) > 1) {
                wilcox.test(Pair(retinalogik, hfa) ~ 1)$p.value
              } else {
                NA_real_
              },
    .groups = "drop") %>%
  ungroup() %>%
  mutate(p_bonferroni = p.adjust(pvalue, method = "bonferroni", n=52)) %>%
  add_row(tibble_row(x = 15, y = -3, meanRT = NA, sdRT = NA)) %>%
  add_row(tibble_row(x = 15, y = +3, meanRT = NA, sdRT = NA)) 

rtplot_earlyGON <- meanSensitivity %>%
  ggplot(aes(x, y, meanRT)) +
  geom_raster(aes(x = x, y = y, fill = meanRT)) +
  geom_text(aes(label = ifelse(is.na(meanRT), "", paste0(round(meanRT, 2), "\n(", round(sdRT, 2), ")")),
                x = x, y = y), size = 3, colour="white") +
  coord_fixed(ratio = 1) +
  scale_fill_gradientn(colours = viridis(51), limits = c(-1, 49),
                       na.value="darkred") +
  theme_bw() +
  ggtitle("VR device", subtitle = paste0("(n = ", nRT, " early GON eyes)")) +
  theme(legend.position="none")

hfaplot_earlyGON <- meanSensitivity %>%
  ggplot(aes(x, y, meanHFA)) +
  geom_raster(aes(x = x, y = y, fill = meanHFA)) +
  geom_text(aes(label = ifelse(is.na(meanHFA), "", paste0(round(meanHFA, 2), "\n(", round(sdHFA, 2), ")")),
                x = x, y = y), size = 3, colour="white") +
  coord_fixed(ratio = 1) +
  scale_fill_gradientn(colours = viridis(51), limits = c(-1, 49),
                       na.value="darkred") +
  theme_bw() +
  ggtitle("HFA", subtitle = paste0("(n = ", nHFA, " early GON eyes)")) +
  theme(legend.position="none")

diffplot_earlyGON <- meanSensitivity %>%
  mutate(diff = meanRT-meanHFA) %>%
  ggplot(aes(x, y, diff)) +
  geom_raster(aes(x = x, y = y, fill = diff)) +
  geom_text(aes(label = ifelse(is.na(diff), "", paste0(round(diff, 2))),
                x = x, y = y), size = 3, colour="black") +
  coord_fixed(ratio = 1) +
  scale_fill_gradientn(colours = colorRampPalette(c("red", "white", "red"))(27), limits = c(-13, 13), 
                       na.value="darkred") +
  theme_bw() +
  ggtitle("Difference (VR device-HFA)") +
  theme(legend.position="none")

# pplot_earlyGON <- meanSensitivity %>%
#   ggplot(aes(x, y)) +
#   geom_tile(fill = "white", colour = "grey90") +
#   geom_text(aes(label = ifelse(pvalue < 0.01, "<.01", round(pvalue, 2)),
#                 x = x, y = y),  size = 3, colour="black") +
#   coord_fixed(ratio = 1) +
#   #scale_fill_gradient(low = "gray", high = "white", na.value="darkred") +
#   theme_bw() +
#   ggtitle("Wilcoxon p-values (Retinalogik-HFA)") +
#   theme(legend.position="none",
#         panel.grid = element_blank())

pplot_earlyGON <- meanSensitivity %>%
  mutate(
    diff = meanRT - meanHFA,
    
    # 1. BACKGROUND COLORS (Red-Blue Palette)
    box_color = case_when(
      is.na(pvalue) ~ "white",
      
      # --- COOL COLORS (Blue) ---
      diff > 0 & pvalue < 0.001 ~ "#0571b0",  # Dark Blue
      diff > 0 & pvalue < 0.01  ~ "#92c5de",  # Medium Blue 
      diff > 0 & pvalue < 0.05  ~ "#d1e5f0",  # Pale Blue

      # --- WARM COLORS (Red) ---
      diff <= 0 & pvalue < 0.001 ~ "#ca0020", # Dark Red
      diff <= 0 & pvalue < 0.01  ~ "#f4a582", # Medium Red
      diff <= 0 & pvalue < 0.05  ~ "#fddbc7", # Pale Red

      TRUE ~ "white"
    ),
    
    # 2. TEXT COLOR LOGIC
    # ONLY turn white if p < 0.001 (Darkest boxes)
    text_color = case_when(
      pvalue < 0.001 ~ "white", 
      TRUE           ~ "black"
    )
  ) %>%
  ggplot(aes(x, y)) +
  geom_tile(aes(fill = box_color), colour = "grey90") +
  scale_fill_identity() + 
  
  # 3. THE FIX: Note that 'colour = text_color' is INSIDE aes()
  # and there is NO colour="black" outside aes()
  geom_text(aes(label = ifelse(is.na(pvalue), "", 
                        ifelse(pvalue < 0.001, "<.001", 
                        ifelse(pvalue < 0.01, "<.01", round(pvalue, 2)))),
                x = x, y = y, 
                colour = text_color), 
            size = 3) +
  
  # 4. Tell ggplot to use the "white"/"black" strings as real colors
  scale_colour_identity() +
  
  coord_fixed(ratio = 1) +
  theme_bw() +
  ggtitle("Wilcoxon p-values (VR device - HFA)", 
          subtitle = "Warm = VR < HFA | Cool = VR > HFA") +
  theme(legend.position="none",
        panel.grid = element_blank())

# Plot using the STRICT Bonferroni values
pplot_bonferroni_earlyGON <- meanSensitivity %>%
  mutate(box_color = case_when(
    p_bonferroni < 0.001 ~ "red",
    p_bonferroni < 0.01  ~ "orange",
    p_bonferroni < 0.05  ~ "yellow",
    TRUE                 ~ "white"
  )) %>%
  ggplot(aes(x, y)) +
  geom_tile(aes(fill = box_color), colour = "grey90") +
  scale_fill_identity() +
  geom_text(aes(label = ifelse(p_bonferroni < 0.001, "<.001", 
                        ifelse(p_bonferroni < 0.01, "<.01", round(p_bonferroni, 2))),
                x = x, y = y), size = 3, colour="black") +
  coord_fixed(ratio = 1) +
  theme_bw() +
  ggtitle("Bonferroni Corrected") +
  theme(legend.position="none", panel.grid = element_blank())

## Moderate GON
nRT <- dBdat_filtered %>%
  filter(classification == "Moderate" & device == "retinalogik") %>%
  distinct(id, device, visit, eye) %>%
  n_distinct()

nHFA <- dBdat_filtered %>%
  filter(classification == "Moderate" & device == "hfa") %>%
  distinct(id, device, visit, eye) %>%
  n_distinct()

meanSensitivity <- dBdat_filtered %>%
  filter(classification == "Moderate") %>%
  distinct(id, device, visit, eye, x, y, dB) %>%
  pivot_wider(names_from = device, values_from = dB) %>%
  group_by(x, y) %>%
  summarize(meanRT = mean(retinalogik, na.rm=T), sdRT = sd(retinalogik, na.rm=T),
            meanHFA = mean(hfa, na.rm=T), sdHFA = sd(hfa, na.rm=T),
            # Add a check for complete pairs
            pvalue  = if(sum(!is.na(retinalogik) & !is.na(hfa)) > 1) {
                wilcox.test(Pair(retinalogik, hfa) ~ 1)$p.value
              } else {
                NA_real_
              },
    .groups = "drop") %>%
  ungroup() %>%
  mutate(p_bonferroni = p.adjust(pvalue, method = "bonferroni", n = 52)) %>%
  add_row(tibble_row(x = 15, y = -3, meanRT = NA, sdRT = NA)) %>%
  add_row(tibble_row(x = 15, y = +3, meanRT = NA, sdRT = NA)) 

rtplot_moderateGON <- meanSensitivity %>%
  ggplot(aes(x, y, meanRT)) +
  geom_raster(aes(x = x, y = y, fill = meanRT)) +
  geom_text(aes(label = ifelse(is.na(meanRT), "", paste0(round(meanRT, 2), "\n(", round(sdRT, 2), ")")),
                x = x, y = y), size = 3, colour="white") +
  coord_fixed(ratio = 1) +
  scale_fill_gradientn(colours = viridis(51), limits = c(-1, 49),
                       na.value="darkred") +
  theme_bw() +
  ggtitle("VR device", subtitle = paste0("(n = ", nRT, " moderate GON eyes)")) +
  theme(legend.position="none")

hfaplot_moderateGON <- meanSensitivity %>%
  ggplot(aes(x, y, meanHFA)) +
  geom_raster(aes(x = x, y = y, fill = meanHFA)) +
  geom_text(aes(label = ifelse(is.na(meanHFA), "", paste0(round(meanHFA, 2), "\n(", round(sdHFA, 2), ")")),
                x = x, y = y), size = 3, colour="white") +
  coord_fixed(ratio = 1) +
  scale_fill_gradientn(colours = viridis(51), limits = c(-1, 49),
                       na.value="darkred") +
  theme_bw() +
  ggtitle("HFA", subtitle = paste0("(n = ", nHFA, " moderate GON eyes)")) +
  theme(legend.position="none")

diffplot_moderateGON <- meanSensitivity %>%
  mutate(diff = meanRT-meanHFA) %>%
  ggplot(aes(x, y, diff)) +
  geom_raster(aes(x = x, y = y, fill = diff)) +
  geom_text(aes(label = ifelse(is.na(diff), "", paste0(round(diff, 2))),
                x = x, y = y), size = 3, colour="black") +
  coord_fixed(ratio = 1) +
  scale_fill_gradientn(colours = colorRampPalette(c("red", "white", "red"))(27), limits = c(-13, 13), 
                       na.value="darkred") +
  theme_bw() +
  ggtitle("Difference (VR device-HFA)") +
  theme(legend.position="none")

pplot_moderateGON <- meanSensitivity %>%
  mutate(
    diff = meanRT - meanHFA,
    
    # 1. BACKGROUND COLORS (Red-Blue Palette)
    box_color = case_when(
      is.na(pvalue) ~ "white",
      
      # --- COOL COLORS (Blue) ---
      diff > 0 & pvalue < 0.001 ~ "#0571b0",  # Dark Blue
      diff > 0 & pvalue < 0.01  ~ "#92c5de",  # Medium Blue 
      diff > 0 & pvalue < 0.05  ~ "#d1e5f0",  # Pale Blue

      # --- WARM COLORS (Red) ---
      diff <= 0 & pvalue < 0.001 ~ "#ca0020", # Dark Red
      diff <= 0 & pvalue < 0.01  ~ "#f4a582", # Medium Red
      diff <= 0 & pvalue < 0.05  ~ "#fddbc7", # Pale Red

      TRUE ~ "white"
    ),
    
    # 2. TEXT COLOR LOGIC
    # ONLY turn white if p < 0.001 (Darkest boxes)
    text_color = case_when(
      pvalue < 0.001 ~ "white", 
      TRUE           ~ "black"
    )
  ) %>%
  ggplot(aes(x, y)) +
  geom_tile(aes(fill = box_color), colour = "grey90") +
  scale_fill_identity() + 
  
  # 3. THE FIX: Note that 'colour = text_color' is INSIDE aes()
  # and there is NO colour="black" outside aes()
  geom_text(aes(label = ifelse(is.na(pvalue), "", 
                        ifelse(pvalue < 0.001, "<.001", 
                        ifelse(pvalue < 0.01, "<.01", round(pvalue, 2)))),
                x = x, y = y, 
                colour = text_color), 
            size = 3) +
  
  # 4. Tell ggplot to use the "white"/"black" strings as real colors
  scale_colour_identity() +
  
  coord_fixed(ratio = 1) +
  theme_bw() +
  ggtitle("Wilcoxon p-values (VR device - HFA)", 
          subtitle = "Warm = VR < HFA | Cool = VR > HFA") +
  theme(legend.position="none",
        panel.grid = element_blank())

# Plot using the STRICT Bonferroni values
pplot_bonferroni_moderateGON <- meanSensitivity %>%
  mutate(box_color = case_when(
    p_bonferroni < 0.001 ~ "red",
    p_bonferroni < 0.01  ~ "orange",
    p_bonferroni < 0.05  ~ "yellow",
    TRUE                 ~ "white"
  )) %>%
  ggplot(aes(x, y)) +
  geom_tile(aes(fill = box_color), colour = "grey90") +
  scale_fill_identity() +
  geom_text(aes(label = ifelse(p_bonferroni < 0.001, "<.001", 
                        ifelse(p_bonferroni < 0.01, "<.01", round(p_bonferroni, 2))),
                x = x, y = y), size = 3, colour="black") +
  coord_fixed(ratio = 1) +
  theme_bw() +
  ggtitle("Bonferroni Corrected") +
  theme(legend.position="none", panel.grid = element_blank())

## Advanced GON
nRT <- dBdat_filtered %>%
  filter(classification == "Advanced" & device == "retinalogik") %>%
  distinct(id, device, visit, eye) %>%
  n_distinct()

nHFA <- dBdat_filtered %>%
  filter(classification == "Advanced" & device == "hfa") %>%
  distinct(id, device, visit, eye) %>%
  n_distinct()

meanSensitivity <- dBdat_filtered %>%
  filter(classification == "Advanced") %>%
  distinct(id, device, visit, eye, x, y, dB) %>%
  pivot_wider(names_from = device, values_from = dB) %>%
  group_by(x, y) %>%
  summarize(meanRT = mean(retinalogik, na.rm=T), sdRT = sd(retinalogik, na.rm=T),
            meanHFA = mean(hfa, na.rm=T), sdHFA = sd(hfa, na.rm=T),
            # Add a check for complete pairs
            pvalue  = if(sum(!is.na(retinalogik) & !is.na(hfa)) > 1) {
                wilcox.test(Pair(retinalogik, hfa) ~ 1)$p.value
              } else {
                NA_real_
              },
    .groups = "drop") %>%
  ungroup() %>%
  mutate(p_bonferroni = p.adjust(pvalue, method = "bonferroni", n = 52)) %>%
  add_row(tibble_row(x = 15, y = -3, meanRT = NA, sdRT = NA)) %>%
  add_row(tibble_row(x = 15, y = +3, meanRT = NA, sdRT = NA)) 

rtplot_advancedGON <- meanSensitivity %>%
  ggplot(aes(x, y, meanRT)) +
  geom_raster(aes(x = x, y = y, fill = meanRT)) +
  geom_text(aes(label = ifelse(is.na(meanRT), "", paste0(round(meanRT, 2), "\n(", round(sdRT, 2), ")")),
                x = x, y = y), size = 3, colour="white") +
  coord_fixed(ratio = 1) +
  scale_fill_gradientn(colours = viridis(51), limits = c(-1, 49),
                       na.value="darkred") +
  theme_bw() +
  ggtitle("VR device", subtitle = paste0("(n = ", nRT, " advanced GON eyes)")) +
  theme(legend.position="none")

hfaplot_advancedGON <- meanSensitivity %>%
  ggplot(aes(x, y, meanHFA)) +
  geom_raster(aes(x = x, y = y, fill = meanHFA)) +
  geom_text(aes(label = ifelse(is.na(meanHFA), "", paste0(round(meanHFA, 2), "\n(", round(sdHFA, 2), ")")),
                x = x, y = y), size = 3, colour="white") +
  coord_fixed(ratio = 1) +
  scale_fill_gradientn(colours = viridis(51), limits = c(-1, 49),
                       na.value="darkred") +
  theme_bw() +
  ggtitle("HFA", subtitle = paste0("(n = ", nHFA, " advanced GON eyes)")) +
  theme(legend.position="none")

diffplot_advancedGON <- meanSensitivity %>%
  mutate(diff = meanRT-meanHFA) %>%
  ggplot(aes(x, y, diff)) +
  geom_raster(aes(x = x, y = y, fill = diff)) +
  geom_text(aes(label = ifelse(is.na(diff), "", paste0(round(diff, 2))),
                x = x, y = y), size = 3, colour="black") +
  coord_fixed(ratio = 1) +
  scale_fill_gradientn(colours = colorRampPalette(c("red", "white", "red"))(27), limits = c(-13, 13), 
                       na.value="darkred") +
  theme_bw() +
  ggtitle("Difference (VR device-HFA)") +
  theme(legend.position="none")

pplot_advancedGON <- meanSensitivity %>%
  mutate(
    diff = meanRT - meanHFA,
    
    # 1. BACKGROUND COLORS (Red-Blue Palette)
    box_color = case_when(
      is.na(pvalue) ~ "white",
      
      # --- COOL COLORS (Blue) ---
      diff > 0 & pvalue < 0.001 ~ "#0571b0",  # Dark Blue
      diff > 0 & pvalue < 0.01  ~ "#92c5de",  # Medium Blue 
      diff > 0 & pvalue < 0.05  ~ "#d1e5f0",  # Pale Blue

      # --- WARM COLORS (Red) ---
      diff <= 0 & pvalue < 0.001 ~ "#ca0020", # Dark Red
      diff <= 0 & pvalue < 0.01  ~ "#f4a582", # Medium Red
      diff <= 0 & pvalue < 0.05  ~ "#fddbc7", # Pale Red

      TRUE ~ "white"
    ),
    
    # 2. TEXT COLOR LOGIC
    # ONLY turn white if p < 0.001 (Darkest boxes)
    text_color = case_when(
      pvalue < 0.001 ~ "white", 
      TRUE           ~ "black"
    )
  ) %>%
  ggplot(aes(x, y)) +
  geom_tile(aes(fill = box_color), colour = "grey90") +
  scale_fill_identity() + 
  
  # 3. THE FIX: Note that 'colour = text_color' is INSIDE aes()
  # and there is NO colour="black" outside aes()
  geom_text(aes(label = ifelse(is.na(pvalue), "", 
                        ifelse(pvalue < 0.001, "<.001", 
                        ifelse(pvalue < 0.01, "<.01", round(pvalue, 2)))),
                x = x, y = y, 
                colour = text_color), 
            size = 3) +
  
  # 4. Tell ggplot to use the "white"/"black" strings as real colors
  scale_colour_identity() +
  
  coord_fixed(ratio = 1) +
  theme_bw() +
  ggtitle("Wilcoxon p-values (VR device - HFA)", 
          subtitle = "Warm = VR < HFA | Cool = VR > HFA") +
  theme(legend.position="none",
        panel.grid = element_blank())

# Plot using the STRICT Bonferroni values
pplot_bonferroni_advancedGON <- meanSensitivity %>%
  mutate(box_color = case_when(
    p_bonferroni < 0.001 ~ "red",
    p_bonferroni < 0.01  ~ "orange",
    p_bonferroni < 0.05  ~ "yellow",
    TRUE                 ~ "white"
  )) %>%
  ggplot(aes(x, y)) +
  geom_tile(aes(fill = box_color), colour = "grey90") +
  scale_fill_identity() +
  geom_text(aes(label = ifelse(p_bonferroni < 0.001, "<.001", 
                        ifelse(p_bonferroni < 0.01, "<.01", round(p_bonferroni, 2))),
                x = x, y = y), size = 3, colour="black") +
  coord_fixed(ratio = 1) +
  theme_bw() +
  ggtitle("Bonferroni Corrected") +
  theme(legend.position="none", panel.grid = element_blank())

# rtplot_earlyGON + hfaplot_earlyGON
# ggsave("plots/earlygon.png", width = 10, height = 6)
# 
# rtplot_moderateGON + hfaplot_moderateGON
# ggsave("plots/moderategon.png", width = 10, height = 6)
# 
# rtplot_advancedGON + hfaplot_advancedGON
# ggsave("plots/advancedgon.png", width = 10, height = 6)
# 
# rtplot_earlyGON + hfaplot_earlyGON + diffplot_earlyGON
# ggsave("plots/earlygon_withdiff.png", width = 14, height = 6)
# 
# rtplot_moderateGON + hfaplot_moderateGON + diffplot_moderateGON
# ggsave("plots/moderategon_withdiff.png", width = 14, height = 6)
# 
# rtplot_advancedGON + hfaplot_advancedGON + diffplot_advancedGON
# ggsave("plots/advancedgon_withdiff.png", width = 14, height = 6)

# rtplot_earlyGON + hfaplot_earlyGON + pplot_earlyGON
# ggsave("plots/earlygon_withpcolours.png", width = 14, height = 6)
# 
# rtplot_moderateGON + hfaplot_moderateGON + pplot_moderateGON
# ggsave("plots/moderategon_withpcolours.png", width = 14, height = 6)
# 
# rtplot_advancedGON + hfaplot_advancedGON + pplot_advancedGON
# ggsave("plots/advancedgon_withpcolours.png", width = 14, height = 6)

rtplot_earlyGON + hfaplot_earlyGON
diffplot_earlyGON+ pplot_earlyGON

rtplot_moderateGON + hfaplot_moderateGON
diffplot_moderateGON + pplot_moderateGON

rtplot_advancedGON + hfaplot_advancedGON
diffplot_advancedGON + pplot_advancedGON
```

### Scotoma analysis
Please refer to [link](https://vvnejl.github.io/retinalogik-study/scotoma_analysis.html).

### Test duration
>The duration of the RL-VF was significantly shorter than that of the Humphrey and did not vary with glaucoma severity in a similar manner to SAP administered with the HFA.

```{r duration, warning=FALSE}
duration <- dBdat_filtered %>%
  group_by(classification, device) %>%
  mutate(duration = as.duration(duration)) %>%
  mutate(mean = mean(duration)/60,
         sd = sd(duration)/60,
         median = median(duration)/60,
         iqr_q1 = quantile(duration, .25)/60, 
         iqr_q3 = quantile(duration, .75)/60) %>%
  distinct(classification, device, mean, sd, median, iqr_q1, iqr_q3) 

duration %>%
  mutate(classification = factor(classification, levels = c("Mild", "Moderate", "Advanced")),
         device = factor(device, levels = c("retinalogik", "hfa"))) %>%
  arrange(classification, device) %>%
  nice_table()

durationdat <- dBdat_filtered %>%
  group_by(id, device) %>%
  mutate(duration = mean(as.duration(duration))) %>%
  distinct(id, classification, device, duration) %>%
  pivot_wider(names_from = device, values_from = duration) %>%
  na.omit()

durationtest <- wilcox.test(durationdat$retinalogik, durationdat$hfa, paired=TRUE)

```

Median test duration time was `r round(median(durationdat$retinalogik)/60, 2)` (IQR = `r round(quantile(durationdat$retinalogik, .25)/60, 2)`, `r round(quantile(durationdat$retinalogik, .75)/60, 2)`) mins for Retinalogik and `r round(median(durationdat$hfa)/60, 2)` (IQR = `r round(quantile(durationdat$hfa, .25)/60, 2)`, `r round(quantile(durationdat$hfa, .75)/60, 2)`) mins for HFA. Wilcoxon signed rank test indicated that this `r round((median(durationdat$hfa)-median(durationdat$retinalogik))/60, 2)` mins difference was statistically significant, *V* = `r durationtest$statistic`, `r if (durationtest$p.value < 0.01) {"$p < 0.01$"} else {paste("$p = ", round(durationtest$p.value, 3), "$", sep="")}`.

### ICC

```{r icc}
# ICC (1,3) Agreement between tests comparing "retinalogik" and "hfa" measurements.
dB_wide_agreement <- dBdat_filtered %>%
  group_by(id, device) %>%
  summarise(mean_dB = mean(dB, na.rm = TRUE)) %>%
  pivot_wider(names_from = device, values_from = mean_dB)

icc_data_agreement <- dB_wide_agreement %>%
  ungroup() %>%
  select(retinalogik, hfa) %>%
  mutate(
    retinalogik = as.numeric(retinalogik),
    hfa = as.numeric(hfa)
  ) %>%
  as.matrix() 

#icc_result_1_3 <- ICC(icc_data_agreement, lmer = FALSE, missing=FALSE) # lmer=FALSE for base R stats
icc_result_1_3 <- ICC(icc_data_agreement, lmer = TRUE, missing=FALSE)
# use result from ICC3 Single_fixed_raters, e.g.,
# Single_fixed_raters      ICC3 0.93 27  32  32 1.4e-15        0.86        0.96

### For repeatability within a device
retinalogik_visits <- dBdat_filtered %>%
  filter(device == "retinalogik") %>%
  # If you have varying numbers of visits, you might need to pad with NA or take a consistent number.
  # For demonstration, let's assume we pivot directly.
  # It's crucial that 'visit' actually represents repeated measures of the *same thing* to be averaged.
  group_by(id, visit) %>%
  summarise(mean_dB_visit = mean(dB, na.rm = TRUE)) %>% # If multiple dB per id, visit, device
  pivot_wider(names_from = visit, values_from = mean_dB_visit, names_prefix = "Visit_") %>%
  ungroup() # Ungroup after pivot_wider

icc_data_retinalogik <- retinalogik_visits %>%
  select(starts_with("Visit_")) %>%
  as.matrix()

icc_result_1_2_retinalogik <- ICC(icc_data_retinalogik)
icc_result_1_2_retinalogik # use ICC2k results

# repeat for HFA
hfa_visits <- dBdat_filtered %>%
  filter(device == "hfa") %>%
  group_by(id, visit) %>%
  summarise(mean_dB_visit = mean(dB, na.rm = TRUE)) %>%
  pivot_wider(names_from = visit, values_from = mean_dB_visit, names_prefix = "Visit_") %>%
  ungroup()

icc_data_hfa <- hfa_visits %>%
  select(starts_with("Visit_")) %>%
  as.matrix()

icc_result_1_2_hfa <- ICC(icc_data_hfa)
icc_result_1_2_hfa # use ICC2k results
```

The ICC values for comparison between Retinalogik and HFA were `r round(icc_result_1_3$results[icc_result_1_3$results$type == "ICC3", "ICC"], 2)` (95% confidence interval [CI]: `r round(icc_result_1_3$results[icc_result_1_3$results$type == "ICC3", "lower bound"], 2)`-`r round(icc_result_1_3$results[icc_result_1_3$results$type == "ICC3", "upper bound"], 2)`). The ICCs for intra-device reproducibility were `r round(icc_result_1_2_retinalogik$results[icc_result_1_2_retinalogik$results$type == "ICC2k", "ICC"], 2)` (95% CI: `r round(icc_result_1_2_retinalogik$results[icc_result_1_2_retinalogik$results$type == "ICC2k", "lower bound"], 2)`-`r round(icc_result_1_2_retinalogik$results[icc_result_1_2_retinalogik$results$type == "ICC2k", "upper bound"], 2)`) for RetinaLogik, and `r round(icc_result_1_2_hfa$results[icc_result_1_2_hfa$results$type == "ICC2k", "ICC"], 2)` (95% CI: `r round(icc_result_1_2_hfa$results[icc_result_1_2_hfa$results$type == "ICC2k", "lower bound"], 2)`-`r round(icc_result_1_2_hfa$results[icc_result_1_2_hfa$results$type == "ICC2k", "upper bound"], 2)`) for the HFA. 

### Questionnaire

```{r questionnaire}
qdat <- read.csv("RetinaLogik Questionnaire Data(Sheet1).csv") %>%
  mutate(
    # Extract the numeric part after "RL"
    numeric_part = as.numeric(str_extract(Patient.ID, "\\d+")),
    # Format the numeric part to have at least two digits with leading zeros
    formatted_numeric = str_pad(numeric_part, width = 2, side = "left", pad = "0"),
    # Combine "RL" with the formatted number
    ID = paste0("RL", formatted_numeric)
  ) %>%
  select(-Patient.ID)

# Recode the columns 
qdat_numeric <- qdat %>%
  mutate(
    Q1..Desktop.difficulty = case_match(Q1..Desktop.difficulty,
      "very easy" ~ 5,
      "somewhat easy" ~ 4,
      "neutral" ~ 3,
      "somewhat difficult" ~ 2,
      "very difficult" ~ 1
    ),
    Q2..Head.mounted.difficulty = case_match(Q2..Head.mounted.difficulty,
      "very easy" ~ 5,
      "somewhat easy" ~ 4,
      "neutral" ~ 3,
      "somewhat difficult" ~ 2,
      "very difficult" ~ 1
    ),
    Q3..Desktop.comfort = case_match(Q3..Desktop.comfort,
      "very comfortable" ~ 5,
      "somewhat comfortable" ~ 4,
      "neutral" ~ 3,
      "somewhat uncomfortable" ~ 2,
      "very uncomfortable" ~ 1
    ),
    Q4..Head.mounted.comfort = case_match(Q4..Head.mounted.comfort,
      "very comfortable" ~ 5,
      "somewhat comfortable" ~ 4,
      "neutral" ~ 3,
      "somewhat uncomfortable" ~ 2,
      "very uncomfortable" ~ 1
    )
  )

summary_stats <- qdat_numeric %>%
  summarise(
    Med_Q1 = median(Q1..Desktop.difficulty, na.rm = TRUE),
    SD_Q1 = sd(Q1..Desktop.difficulty, na.rm = TRUE),
    Med_Q2 = median(Q2..Head.mounted.difficulty, na.rm = TRUE),
    SD_Q2 = sd(Q2..Head.mounted.difficulty, na.rm = TRUE),
    Med_Q3 = median(Q3..Desktop.comfort, na.rm = TRUE),
    SD_Q3 = sd(Q3..Desktop.comfort, na.rm = TRUE),
    Med_Q4 = median(Q4..Head.mounted.comfort, na.rm = TRUE),
    SD_Q4 = sd(Q4..Head.mounted.comfort, na.rm = TRUE)
  )

print(summary_stats)

patient_summary <- qdat_numeric %>%
  group_by(ID) %>%
  summarise(
    Med_Q1 = median(Q1..Desktop.difficulty, na.rm = TRUE),
    Med_Q2 = median(Q2..Head.mounted.difficulty, na.rm = TRUE),
    Med_Q3 = median(Q3..Desktop.comfort, na.rm = TRUE),
    Med_Q4 = median(Q4..Head.mounted.comfort, na.rm = TRUE)
  )

print(patient_summary)

questionnaire_plot <- summary_stats %>%
  tidyr::pivot_longer(
    cols = c(Med_Q1, Med_Q2, Med_Q3, Med_Q4),
    names_to = "Question",
    values_to = "Median_Score"
  )

questionnaire_plot %>%
  ggplot(aes(x = Question, y = Median_Score)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  labs(
    title = "Median Scores for each Question",
    x = "Questions",
    y = "Median Score (1-5 Likert Scale)",
    caption = "On the difficulty scale: 1 is very difficult, 5 is very easy;
    On the comfort scale: 1 is very uncomfortable, 5 is very comfortable."
  ) +
  scale_x_discrete(labels = c("HFA difficulty", "Retinalogik difficulty",
                              "HFA comfort", "Retinalogik comfort")) +
  theme_minimal()

# Wilcoxon signed-rank test
wilcox_difficulty <- wilcox.test(qdat_numeric$Q1..Desktop.difficulty, 
                                 qdat_numeric$Q2..Head.mounted.difficulty, 
                                 paired = TRUE)

wilcox_comfort <- wilcox.test(qdat_numeric$Q3..Desktop.comfort, 
                              qdat_numeric$Q4..Head.mounted.comfort, 
                              paired = TRUE)

q5dat <- qdat %>%
  select(ID, Q5..Overall.preference) %>%
  mutate(Q5..Overall.preference = as.factor(Q5..Overall.preference)) 

q5dat <- q5dat[q5dat$Q5..Overall.preference != "", ] %>%
  count(Q5..Overall.preference)

```
**Q5 is "Which test would you prefer to perform on an on-going basis?"**

`r round(q5dat$n[1]/sum(q5dat$n)*100, 2)`% prefer "desktop" aka HFA, `r round(q5dat$n[2]/sum(q5dat$n)*100, 2)`% prefer head-mounted aka Retinalogik, and `r round(q5dat$n[3]/sum(q5dat$n)*100, 2)`% has no preference for either.

