---
title: "Retinalogik vs HFA study JAMA Ophthalmology paper"
author: "[Eadie Technologies, Inc.](https://www.eadietech.com/)"
date: "2025-06-11"
output: 
  html_document:
    theme: yeti
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: 
      collapsed: true
    includes:
        after_body: footer.html
---

## Setup

### Versioning

Last updated on `r format(Sys.time(), '%Y-%b-%d')` at `r format(Sys.time(), '%I:%M %p')`.

* 2025-Jun-11: Created separate .Rmd file as index.Rmd was getting very cluttered.

### Acknowledgements
Data collected by Research Assistants Luc St-Pierre, Reann Post, and Research Coordinator Lynn Murphy.

```{r setup, collapse=TRUE}

# load packages
library(tidyverse)
library(here)
library(DT)
library(plotly)
library(viridis)
library(svglite)
library(htmltools)
library(gridExtra)
library(sf)
library(spdep)
library(RColorBrewer)
library(patchwork)
library(report)
library(effectsize)

sessionInfo()

# default chunk options
knitr::opts_chunk$set(
  comment = '>', collapse = TRUE, cache = FALSE, dev= c("png")
  )

rm(list = ls()) 

# load processed data
load(here("dBdat.Rda"))

load(here("retinalogik_hfadatawpsd.Rda"))

dBdat_filtered <- dBdat %>%
  mutate(x = as.numeric(x), y = as.numeric(y)) %>%
  mutate(x = case_when(
    eye == "L" & device == "retinalogik" ~ x*-1,
    TRUE ~ as.numeric(x)
  )) %>%
  # remove blind spots
  filter(!(x == 15 & y == -3)) %>%
  filter(!(x == 15 & y == +3)) %>%
  #remove first 2 visits
  filter(visit != 1 & visit != 2) %>%
  #remove reliability indices <30%
  filter(flPerc < 30 & fnPerc < 30 & fpPerc < 30) %>%
  group_by(id, device, eye) %>%
  mutate(meanmd = mean(md))

df <- dBdat_filtered %>%
  distinct(id, device, visit, eye, x, y, dB) %>%
  group_by(id, device, eye, x, y) %>%
  mutate(meandB = mean(dB)) %>%
  distinct(id, device, eye, x, y, meandB) %>%
  pivot_wider(names_from = device, values_from = meandB) %>%
  mutate(retinalogik = as.numeric(retinalogik), hfa = as.numeric(hfa))

```

# Study summary
We compared 24-2 visual field (VF) test results between Retinalogik and the Humphrey Field Analyzer (HFA) using SITA Standard test strategy in patients with early glaucoma and moderate to advanced glaucoma [here](https://vvnejl.github.io/retinalogik-study/). A scotoma analysis was performed following glaucomatous VF definitions found in [Park et al. (2011)](https://www.aaojournal.org/article/S0161-6420(11)00172-2/abstract).

# Methods (very briefly)
We define a scotoma as the presence of at least 3 vertically, horizontally, or diagonally contiguous test points within the same hemifield on the pattern deviation probability plot at p <5%, with at least 1 point at p <1%, excluding points directly above or below the blind spot. Eyes with consistent glaucomatous VF loss on at least 2 consecutive examinations with reliability indices better than 30% were included. 

## Analysis

### Mean sensitivity
> The mean sensitivity was significantly higher in the VR-RL perimeter (stats).

```{r meansens}
# meansens <- dBdat_filtered %>%
#   group_by(device) %>%
#   summarise(meandB = mean(dB), sd = sd(dB))

df <- dBdat_filtered %>%
  group_by(id, device) %>%
  mutate(meandB = mean(dB)) %>%
  distinct(id, device, meandB) %>%
  pivot_wider(names_from = device, values_from = meandB) %>%
  mutate(retinalogik = as.numeric(retinalogik), hfa = as.numeric(hfa)) %>%
  na.omit()

tmp <- t.test(df$retinalogik, df$hfa, paired=T) %>%
  report()

```

Mean sensitivity was significantly higher in the VR-RL perimeter (*M* = `r round(mean(df$retinalogik),2)`, *SD* = `r round(sd(df$retinalogik), 2)`) than HFA (*M* = `r round(mean(df$hfa), 2)`, *SD* = `r round(sd(df$hfa), 2)`). Paired t-test suggests that the  difference in mean sensitivity (`r round(mean(df$retinalogik)-mean(df$hfa), 2)` dB, 95% CI [0.58, 4.11]) is statistically significant (t(25) = 2.73, p = 0.011) and moderate (Cohen’s d = 0.54).

### Test duration

```{r duration, warning=FALSE}
duration <- dBdat_filtered %>%
  group_by(device) %>%
  mutate(mean = mean(as.duration(duration)), sd = sd(as.duration(duration))) %>%
  distinct(device, mean, sd)

rthfaduration <- dBdat_filtered %>%
  distinct(id, visit, eye, duration) %>%
  group_by(id, device) %>%
  mutate(duration = mean(duration)) %>%
  distinct(device, id, duration) %>%
  pivot_wider(names_from = device, values_from = duration) %>%
  mutate(retinalogik = as.numeric(retinalogik), hfa = as.numeric(hfa)) %>%
  na.omit()

t.test(as.numeric(rthfaduration$retinalogik), as.numeric(rthfaduration$hfa), paired=T)

```

Mean test duration time was `r round(duration$mean[duration$device=="retinalogik"]/60, 2)` (SD = `r round(duration$sd[duration$device=="retinalogik"]/60, 2)`) mins for Retinalogik and `r round(duration$mean[duration$device=="hfa"]/60, 2)` (SD = `r round(duration$sd[duration$device=="hfa"]/60, 2)`) mins for HFA (p < .01).

### Threshold sensitivities
Threshold sensitivities were significantly higher in peripheral locations with the VR-RL perimeter. Note: Do a ?linear regression with dB ~ perimeter * age * x * y 

### Test-retest variability
> with conventional perimetry, test-retest variability increases with defect depth and eccentricity. [(Heijl, Lindgren & Lindgren, 1989)](https://pubmed.ncbi.nlm.nih.gov/2757094/)

> The global test-retest variability was assessed by the root mean square deviation (RMSD) of
the TD values. We define RMSD as the square root of average square of TD differences over
all 52 locations between two repeated measurements... [(Choi et al., 2021)](https://www.sciencedirect.com/science/article/abs/pii/S2589419620303197).

> The test-retest variability at each location on the 24-2 grid of each participant was calculated as the absolute value of the difference in pointwise sensitivity of VF test 1 and test 2 at that location... The mean pointwise test-retest variability at a particular grid
location was calculated as the mean value of the variability at that
particular location of each of the 1426 participants... [(Tan, Agar, Kalloniatis & Phu, 2023)](https://doi.org/10.1016/j.ophtha.2023.12.018)

```{r test-retest}



```
