---
title: "Retinalogik vs HFA study scotoma analysis"
author: "[Eadie Technologies, Inc.](https://www.eadietech.com/)"
date: "2025-06-04"
output: 
  html_document:
    theme: yeti
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: 
      collapsed: true
    includes:
        after_body: footer.html
---

## Setup

### Versioning

Last updated on `r format(Sys.time(), '%Y-%b-%d')` at `r format(Sys.time(), '%I:%M %p')`.

* 2025-Jun-04: Created separate .Rmd file as index.Rmd was getting very cluttered.

### Acknowledgements
Data collected by Research Assistants Luc St-Pierre, Reann Post, and Research Coordinator Lynn Murphy.

```{r setup, collapse=TRUE}

# load packages
library(tidyverse)
library(here)
library(DT)
library(plotly)
library(viridis)
library(svglite)
library(htmltools)
library(gridExtra)
library(sf)
library(spdep)
library(RColorBrewer)
library(patchwork)
library(spatstat.geom)

sessionInfo()

# default chunk options
knitr::opts_chunk$set(
  comment = '>', collapse = TRUE, cache = FALSE, dev= c("png")
  )

# load processed data
load(here("dBdat.Rda"))

load(here("retinalogik_hfadatawpsd.Rda"))

```

# Study summary
We compared 24-2 visual field (VF) test results between Retinalogik and the Humphrey Field Analyzer (HFA) using SITA Standard test strategy in patients with early glaucoma and moderate to advanced glaucoma [here](https://vvnejl.github.io/retinalogik-study/). A scotoma analysis was performed following glaucomatous VF definitions found in [Park et al. (2011)](https://www.aaojournal.org/article/S0161-6420(11)00172-2/abstract).

# Methods (very briefly)
We define a scotoma as the presence of at least 3 vertically, horizontally, or diagonally contiguous test points within the same hemifield on the pattern deviation probability plot at p <5%, with at least 1 point at p <1%, excluding points directly above or below the blind spot. Eyes with consistent glaucomatous VF loss on at least 2 consecutive examinations with reliability indices better than 30% were included. 

## Analysis {.tabset}
```{r functions}

# --- Define the main analysis function ---
analyze_visual_field_data <- function(my_data_input, rttmp_raw_input) {

  # 1. Data Preparation and Spatial Object Creation
  # Create a clean psd_p column, treating NA as non-defective (e.g., value > 5)
  my_data_processed <- my_data_input %>%
    mutate(psd_p_clean = ifelse(is.na(psd_p), 100, psd_p))

  # Convert the data frame to an sf (simple features) object
  points_sf <- st_as_sf(my_data_processed, coords = c("x", "y"), crs = NA)

  # 2. Filtering Defective Points
  # Filter for points where psd_p is less than 5% (including 0.5% and 5.0%)
  filtered_points_sf <- points_sf %>%
    filter(psd_p_clean < 5)

  # Initialize outputs
  identified_scotomas_table_df <- data.frame(
    Scotoma_ID = integer(),
    Number_of_Points = integer(),
    Minimum_psd_p_percent = numeric(),
    Constituent_Points = character(),
    stringsAsFactors = FALSE
  )
  all_scotoma_points_for_plot <- data.frame()

  # Ensure there are defective points to analyze
  if (nrow(filtered_points_sf) > 0) {
    # 3. Connected Components Labeling using Distance-Based Neighbors
    # Create a spatial weights matrix based on distance for the filtered points
    # d1 = 0: minimum distance (points must be distinct)
    # d2 = 9: maximum distance (captures vertical, horizontal, and diagonal neighbors
    #         given your coordinate spacing of 6 units, sqrt(6^2 + 6^2) approx 8.485)
    filtered_neighbors <- dnearneigh(filtered_points_sf, d1 = 0, d2 = 9)

    # Check if there are any connections among the filtered points
    if (length(unlist(filtered_neighbors)) > 0) {
      # Find connected components (contiguous groups)
      components <- n.comp.nb(filtered_neighbors)

      # Assign component IDs back to the filtered data
      filtered_points_sf$component_id <- components$comp.id

      # 4. Filtering Connected Components by Scotoma Criteria
      scotomas_found <- list()
      unique_component_ids <- unique(filtered_points_sf$component_id)

      for (comp_id in unique_component_ids) {
        current_cluster <- filtered_points_sf %>%
          filter(component_id == comp_id)

        num_points <- nrow(current_cluster)
        # Check if any point in this cluster has psd_p < 1% (i.e., psd_p_clean == 0.5)
        has_p_lt_1_percent <- any(current_cluster$psd_p_clean < 1)

        # Apply scotoma definition: at least 3 points AND at least 1 location at p < 1%
        if (num_points >= 3 && has_p_lt_1_percent) {
          scotomas_found[[as.character(comp_id)]] <- current_cluster %>%
            # Extract x and y coordinates from the geometry column
            mutate(
              x_coord = st_coordinates(.)[, "X"],
              y_coord = st_coordinates(.)[, "Y"]
            ) %>%
            # Now select the desired columns, including the newly extracted x_coord and y_coord
            select(x_coord, y_coord, psd_p) %>%
            # Drop the geometry column as it's no longer needed for the table output
            st_drop_geometry() %>%
            # Rename x_coord and y_coord back to x and y for consistency with the table format
            rename(x = x_coord, y = y_coord)
        }
      }

      # Prepare results for the report table
      scotoma_results_list <- list()
      if (length(scotomas_found) > 0) {
        scotoma_id_counter <- 1
        for (scotoma_label in names(scotomas_found)) {
          scotoma_data <- scotomas_found[[scotoma_label]]
          num_points <- nrow(scotoma_data)
          min_psd_p_actual <- min(scotoma_data$psd_p, na.rm = TRUE) # Use original psd_p for reporting

          # Format constituent points string
          constituent_points_str <- paste(
            paste0("(", scotoma_data$x, ",", scotoma_data$y, ") [p=", scotoma_data$psd_p, "%]"),
            collapse = "; "
          )

          scotoma_results_list[[scotoma_id_counter]] <- data.frame(
            Scotoma_ID = scotoma_id_counter,
            Number_of_Points = num_points,
            Minimum_psd_p_percent = min_psd_p_actual,
            Constituent_Points = constituent_points_str,
            stringsAsFactors = FALSE
          )
          scotoma_id_counter <- scotoma_id_counter + 1
        }
      }

      # Final table of identified scotomas
      identified_scotomas_table_df <- if (length(scotoma_results_list) > 0) {
        do.call(rbind, scotoma_results_list)
      } else {
        data.frame(
          Scotoma_ID = integer(),
          Number_of_Points = integer(),
          Minimum_psd_p_percent = numeric(),
          Constituent_Points = character(),
          stringsAsFactors = FALSE
        )
      }

      # Prepare all scotoma points for plotting with a sequential ID
      if (length(scotomas_found) > 0) {
        scotoma_id_map <- 1
        for (original_comp_id in names(scotomas_found)) {
          temp_df <- scotomas_found[[original_comp_id]]
          temp_df$scotoma_plot_id <- scotoma_id_map # Assign a sequential ID for plotting colors
          all_scotoma_points_for_plot <- bind_rows(all_scotoma_points_for_plot, temp_df)
          scotoma_id_map <- scotoma_id_map + 1
        }
      }
    }
  }

  # Prepare data for Retinalogik dB plot
  # Filter rttmp_raw_input to only include points present in my_data_processed (50 points)
  rttmp_filtered_for_plot <- rttmp_raw_input %>%
    semi_join(my_data_processed, by = c("x", "y"))

  # Join filtered_rttmp with my_data_processed to get psd_p for coloring the background
  rttmp_for_plot_with_psdp <- rttmp_filtered_for_plot %>%
    left_join(my_data_processed %>% select(x, y, psd_p), by = c("x", "y"))

  return(list(
    scotoma_table = identified_scotomas_table_df,
    hfa_data_processed = my_data_processed,
    scotoma_points_for_plot = all_scotoma_points_for_plot,
    rttmp_data_for_plot = rttmp_for_plot_with_psdp
  ))
}

# --- Define plotting functions ---

generate_hfa_psd_plot <- function(hfa_data_processed) {
  customcolors <- brewer.pal(4, "YlOrRd")
  customcolors <- rev(customcolors)

  hfa_data_processed %>%
    ggplot(aes(x, y)) +
    geom_raster(aes(fill = factor(psd_p))) +
    geom_text(aes(label = ifelse(is.na(psd_p), "", psd_p)), size = 3) +
    coord_fixed(ratio = 1) +
    scale_fill_manual(values = customcolors, na.value = "gray") +
    theme_bw() +
    ggtitle("HFA", subtitle = "PSD p-values") +
    theme(legend.position = "none")
}

generate_hfa_db_plot <- function(hfa_data_processed) {
  customcolors <- brewer.pal(4, "YlOrRd")
  customcolors <- rev(customcolors)

  hfa_data_processed %>%
    ggplot(aes(x, y)) +
    geom_raster(aes(fill = factor(psd_p))) +
    geom_text(aes(label = ifelse(is.na(dB), "", dB)), size = 3) +
    coord_fixed(ratio = 1) +
    scale_fill_manual(values = customcolors, na.value = "gray") +
    theme_bw() +
    ggtitle("HFA", subtitle = "Threshold sensitivity dB values") +
    theme(legend.position = "none")
}

generate_scotoma_highlight_plot <- function(hfa_data_processed, scotoma_points_for_plot) {
  ggplot() +
    geom_point(data = hfa_data_processed, aes(x = x, y = y), color = "grey", size = 1, alpha = 0.5) +
    {
      if (nrow(scotoma_points_for_plot) > 0) {
        list(
          geom_point(data = scotoma_points_for_plot,
                     aes(x = x, y = y, color = as.factor(scotoma_plot_id)),
                     size = 3, shape = 16),
          scale_color_viridis_d(option = "viridis", name = "Scotoma ID"),
          geom_point(data = scotoma_points_for_plot,
                     aes(x = x, y = y),
                     color = "black", size = 5, shape = 1)
        )
      }
    } +
    labs(title = "Identified Scotomas on Visual Field",
         x = "X-coordinate (degrees)", y = "Y-coordinate (degrees)") +
    theme_minimal() +
    theme(legend.position = "right")
}

generate_retinalogik_plot <- function(rttmp_data_for_plot) {
  retinalogik_colors <- rev(brewer.pal(3, "YlOrRd"))

  rttmp_data_for_plot %>%
    ggplot(aes(x, y)) +
    geom_raster(aes(fill = factor(psd_p))) +
    geom_text(aes(label = ifelse(is.na(dB), "", dB)), size = 3) +
    coord_fixed(ratio = 1) +
    scale_fill_manual(values = retinalogik_colors, na.value = "gray") +
    theme_bw() +
    ggtitle("Retinalogik", subtitle = "Threshold sensitivity dB values") +
    theme(legend.position = "none")
}

```
### RL35 {.tabset}
#### Visit 1
```{r scotoma1, warning=FALSE, fig.width=7.5, fig.height=7.5}
# remove locations directly above and below blindspots
psdpvalues_hfa <- alldat %>%
  filter(!(x==15 & y==9) & !(x==15 & y==-9))

# remove reliability indices <30%
psdpvalues_hfa <- psdpvalues_hfa %>%
  filter((fl/flTest) < 0.3 & fnPerc < 0.3 & fpPerc < 0.3)

tmp <- psdpvalues_hfa %>%
  filter(id=="RL35" & eye=="R")

customcolors <- brewer.pal(4, "YlOrRd")
customcolors <- rev(customcolors)

my_data <- psdpvalues_hfa %>%
  filter(datetime==ymd_hms("2025-05-20 09:48:31 ") & eye=="R")

rttmp <- dBdat %>%
  filter(id=="RL35", visit==1, eye=="R", device=="retinalogik") %>%
  select(id, eye, x, y, dB)

# 1. Data Preparation and Binary Grid Creation
my_data_processed <- my_data %>%
  mutate(psd_p_clean = ifelse(is.na(psd_p), 100, psd_p))

# Convert the data frame to an sf (simple features) object
points_sf <- st_as_sf(my_data_processed, coords = c("x", "y"), crs = NA)

# 2. Filtering Defective Points
# Filter for points where psd_p is less than 5% (including 0.5% and 5.0%)
filtered_points_sf <- points_sf %>%
  filter(psd_p_clean < 5)

# Ensure there are defective points to analyze
if (nrow(filtered_points_sf) > 0) {
  # 3. Connected Components Labeling using Distance-Based Neighbors
  # Create a spatial weights matrix based on distance for the filtered points
  # d1 = 0: minimum distance (points must be distinct)
  # d2 = 9: maximum distance (captures vertical, horizontal, and diagonal neighbors
  #         given your coordinate spacing of 6 units, sqrt(6^2 + 6^2) approx 8.485)
  filtered_neighbors <- dnearneigh(filtered_points_sf, d1 = 0, d2 = 9)

  # Check if there are any connections among the filtered points
  if (length(unlist(filtered_neighbors)) > 0) {
    # Find connected components (contiguous groups)
    components <- n.comp.nb(filtered_neighbors)

    # Assign component IDs back to the filtered data
    filtered_points_sf$component_id <- components$comp.id

    # 4. Filtering Connected Components by Scotoma Criteria
    scotomas_found <- list()
    unique_component_ids <- unique(filtered_points_sf$component_id)

    for (comp_id in unique_component_ids) {
      current_cluster <- filtered_points_sf %>%
        filter(component_id == comp_id)

      num_points <- nrow(current_cluster)
      # Check if any point in this cluster has psd_p < 1% (i.e., psd_p_clean == 0.5)
      has_p_lt_1_percent <- any(current_cluster$psd_p_clean < 1)

      # Apply scotoma definition: at least 3 points AND at least 1 location at p < 1%
      # Apply scotoma definition: at least 3 points AND at least 1 location at p < 1%
      if (num_points >= 3 && has_p_lt_1_percent) {
        scotomas_found[[as.character(comp_id)]] <- current_cluster %>%
          # Extract x and y coordinates from the geometry column
          mutate(
            x_coord = st_coordinates(.)[, "X"],
            y_coord = st_coordinates(.)[, "Y"]
          ) %>%
          # Now select the desired columns, including the newly extracted x_coord and y_coord
          select(x_coord, y_coord, psd_p) %>%
          # Drop the geometry column as it's no longer needed for the table output
          st_drop_geometry() %>%
          # Rename x_coord and y_coord back to x and y for consistency with the table format
          rename(x = x_coord, y = y_coord)
      }
    }

    # Prepare results for the report table
    scotoma_results_list <- list()
    if (length(scotomas_found) > 0) {
      scotoma_id_counter <- 1
      for (scotoma_label in names(scotomas_found)) {
        scotoma_data <- scotomas_found[[scotoma_label]]
        num_points <- nrow(scotoma_data)
        min_psd_p_actual <- min(scotoma_data$psd_p, na.rm = TRUE) # Use original psd_p for reporting

        # Format constituent points string
        constituent_points_str <- paste(
          paste0("(", scotoma_data$x, ",", scotoma_data$y, ") [p=", scotoma_data$psd_p, "%]"),
          collapse = "; "
        )

        scotoma_results_list[[scotoma_id_counter]] <- data.frame(
          Scotoma_ID = scotoma_id_counter,
          Number_of_Points = num_points,
          Minimum_psd_p_percent = min_psd_p_actual,
          Constituent_Points = constituent_points_str,
          stringsAsFactors = FALSE
        )
        scotoma_id_counter <- scotoma_id_counter + 1
      }
    }

    # Final table of identified scotomas
    identified_scotomas_table_df <- if (length(scotoma_results_list) > 0) {
      do.call(rbind, scotoma_results_list)
    } else {
      data.frame(
        Scotoma_ID = integer(),
        Number_of_Points = integer(),
        Minimum_psd_p_percent = numeric(),
        Constituent_Points = character(),
        stringsAsFactors = FALSE
      )
    }

  } else {
    identified_scotomas_table_df <- data.frame(
      Scotoma_ID = integer(),
      Number_of_Points = integer(),
      Minimum_psd_p_percent = numeric(),
      Constituent_Points = character(),
      stringsAsFactors = FALSE
    )
  }
} else {
  identified_scotomas_table_df <- data.frame(
    Scotoma_ID = integer(),
    Number_of_Points = integer(),
    Minimum_psd_p_percent = numeric(),
    Constituent_Points = character(),
    stringsAsFactors = FALSE
  )
}

# Print the table (for verification within the R environment)
print(identified_scotomas_table_df)

# Define custom colors for psd_p values for the background raster
customcolors <- brewer.pal(4, "YlOrRd")
customcolors <- rev(customcolors)

# Plot 1: HFA Pattern Deviation Probability Plot (similar to user's 'p' plot)
# This plot shows the psd_p values for all test points.
p_plot <- my_data_processed %>%
  ggplot(aes(x, y)) +
  geom_raster(aes(fill = factor(psd_p))) + # Use original psd_p for fill
  geom_text(aes(label = ifelse(is.na(psd_p), "", psd_p)), size = 3) +
  coord_fixed(ratio = 1) +
  scale_fill_manual(values = customcolors, na.value = "gray") +
  theme_bw() +
  ggtitle("HFA", subtitle = "PSD p-values") +
  theme(legend.position = "none")

dB_plot <- my_data_processed %>%
  ggplot(aes(x, y)) +
  geom_raster(aes(fill = factor(psd_p))) + # Use original psd_p for fill
  geom_text(aes(label = ifelse(is.na(dB), "", dB)), size = 3) +
  coord_fixed(ratio = 1) +
  scale_fill_manual(values = customcolors, na.value = "gray") +
  theme_bw() +
  ggtitle("HFA", subtitle = "Threshold sensitivity dB values") +
  theme(legend.position = "none")

rttmp_filtered_for_plot <- rttmp %>%
  semi_join(my_data_processed, by = c("x", "y"))

# Join filtered_rttmp with my_data_processed to get psd_p for coloring the background
rttmp_for_plot_with_psdp <- rttmp_filtered_for_plot %>%
  left_join(my_data_processed %>% select(x, y, psd_p), by = c("x", "y"))

# Plot 4: Retinalogik Threshold Sensitivity dB values
rt_plot <- rttmp_for_plot_with_psdp %>%
  ggplot(aes(x, y)) +
  geom_raster(aes(fill = factor(psd_p))) + # Use HFA psd_p for fill as per user's example
  geom_text(aes(label = ifelse(is.na(dB), "", dB)), size = 3) + # Use Retinalogik dB for label
  coord_fixed(ratio = 1) +
  scale_fill_manual(values = customcolors, na.value = "gray") +
  theme_bw() +
  ggtitle("Retinalogik", subtitle = "Threshold sensitivity dB values") +
  theme(legend.position = "none")

# Prepare data for the scotoma highlight plot
# Combine all identified scotoma points into a single data frame for plotting
all_scotoma_points_for_plot <- data.frame()
if (length(scotomas_found) > 0) {
  scotoma_id_map <- 1
  for (original_comp_id in names(scotomas_found)) {
    temp_df <- scotomas_found[[original_comp_id]]
    temp_df$scotoma_plot_id <- scotoma_id_map # Assign a sequential ID for plotting colors
    all_scotoma_points_for_plot <- bind_rows(all_scotoma_points_for_plot, temp_df)
    scotoma_id_map <- scotoma_id_map + 1
  }
}

# Plot 2: Identified Scotomas Highlighted (similar to user's 'plot_scotoma')
# This plot overlays the identified scotoma points on a grey background of all points.
plot_scotoma_highlight <- ggplot() +
  geom_point(data = my_data_processed, aes(x = x, y = y), color = "grey", size = 1, alpha = 0.5) + # All points as background
  {
    if (nrow(all_scotoma_points_for_plot) > 0) {
      list(
        geom_point(data = all_scotoma_points_for_plot,
                   aes(x = x, y = y, color = as.factor(scotoma_plot_id)),
                   size = 3, shape = 16), # Highlight scotoma points
        scale_color_viridis_d(option = "viridis", name = "Scotoma ID"), # Use viridis for distinct scotoma colors
        geom_point(data = all_scotoma_points_for_plot,
                   aes(x = x, y = y),
                   color = "black", size = 5, shape = 1) # Add black outline for emphasis
      )
    }
  } +
  labs(title = "Identified Scotomas",
       x = "x", y = "y") +
  theme_minimal() +
  theme(legend.position = "right")

(p_plot + plot_scotoma_highlight)/
  (dB_plot + rt_plot)

```

#### Visit 2
```{r scotoma2, warning=FALSE, fig.width=7.5, fig.height=7.5}
tmp <- psdpvalues_hfa %>%
  filter(id=="35" & eye=="R")

my_data <- psdpvalues_hfa %>%
  filter(datetime==ymd_hms("2025-05-22 09:40:31") & eye=="R")

customcolors <- brewer.pal(4, "YlOrRd")
customcolors <- rev(customcolors)

p <- my_data %>%
  ggplot(aes(x, y, psd_p)) +
  geom_raster(aes(x = x, y = y, fill = factor(psd_p))) +
  geom_text(aes(label = ifelse(is.na(psd_p), "", psd_p),
                x = x, y = y), size = 3) +
  coord_fixed(ratio = 1) +
  scale_fill_manual(values = customcolors, na.value="gray") +
  theme_bw() +
  ggtitle("HFA", subtitle = "PSD p-values") +
  theme(legend.position="none")

dBplot <- my_data %>%
  ggplot(aes(x, y, psd_p, dB)) +
  geom_raster(aes(x = x, y = y, fill = factor(psd_p))) +
  geom_text(aes(label = dB, x = x, y = y), size = 3) +
  coord_fixed(ratio = 1) +
  scale_fill_manual(values = customcolors, na.value="gray") +
  theme_bw() +
  ggtitle("HFA", subtitle = "Threshold sensitivity dB values") +
  theme(legend.position="none")
  
my_sf <- my_data %>%
  select(datetime, eye, id, x, y, dB, psd_p) %>%
  st_as_sf(coords = c("x", "y"), crs = NA)

# You can view the data
#plot(my_sf)

# Convert to sf object
# It's good practice to provide a CRS if your coordinates represent real-world locations.
# For simple relative coordinates, NA is fine.
points_sf <- st_as_sf(my_sf, wkt = "geometry", crs = NA)

# Create a spatial weights matrix based on distance for point data
# d1 = 0: minimum distance (points must be distinct)
# d2 = 9: maximum distance (captures vertical, horizontal, and diagonal neighbors
#         given your coordinate spacing of 6 units)
neighbors <- dnearneigh(points_sf, d1 = 0, d2 = 9)

# Convert the neighbor list to a spatial weights list (binary style is common)
weights_list <- nb2listw(neighbors, style = "B")

# Filter for points where psd_p <= 5
# Using is.finite() to handle NA values in psd_p, as NA <= 5 would also be NA, not FALSE.
filtered_points <- points_sf %>%
  filter(is.finite(psd_p) & psd_p <= 5)

if(nrow(filtered_points) > 0) {
  # Create a spatial weights matrix for the filtered points (re-evaluate neighbors)
  # This ensures we only consider contiguity among the *filtered* points.
  filtered_neighbors <- dnearneigh(filtered_points, d1 = 0, d2 = 9)
  
  # Check if there are any connections among the filtered points
  if (length(unlist(filtered_neighbors)) > 0) { # Check if any neighbors exist
    # Find connected components (contiguous groups)
    components <- n.comp.nb(filtered_neighbors)
    
    # Assign component IDs back to the filtered data
    filtered_points$component_id <- components$comp.id
    
    # Group by component and count points
    contiguous_groups <- filtered_points %>%
      group_by(component_id) %>%
      summarize(n_points = n()) %>%
      arrange(desc(n_points))
    
    # Print the results
    #print("Contiguous groups where psd_p <= 5:")
    #print(contiguous_groups)
    
    # Find the largest contiguous group
    largest_group_id <- contiguous_groups$component_id[1]
    
    largest_group <- filtered_points %>%
      filter(component_id == largest_group_id)
    
    #print("Largest Contiguous Group:")
    #print(largest_group)
    
    if (nrow(largest_group) >= 3) {
      #print("Largest contiguous group has at least 3 points.")
      # Optionally, extract and print the coordinates of the largest group
      #print("Coordinates of the largest contiguous group:")
      #print(st_coordinates(largest_group))
    } else {
      #print("No contiguous group with at least 3 points found that has at least 3 points.")
    }
  } else {
    #print("No contiguous points found after filtering for psd_p <= 5.")
  }
  
} else {
  #print("No points found where psd_p <= 5.")
}

# --- Visualization ---

# Plot all original points
# plot_all_points <- ggplot() +
#   geom_sf(data = points_sf, color = "grey", size = 1, alpha = 0.5) +
#   labs(title = "All Original Points (Grey) and Filtered Contiguous Groups (Colored)",
#        x = "X-coordinate", y = "Y-coordinate") +
#   theme_minimal()

plot_all_points <- ggplot() +
  geom_sf(data = points_sf, color = "grey", size = 1, alpha = 0.5) +
  labs(title = "Contiguous Groups",
       x = "x", y = "y") +
  theme_bw()

# Add the filtered and grouped points on top, colored by their component_id
# Use a specific color palette if you have many groups
plot_scotoma <- plot_all_points +
  geom_sf(data = filtered_points, aes(color = as.factor(component_id)), size = 3) +
  scale_color_viridis_d(option = "viridis", name = "Contiguous Group ID") # Viridis is a good perceptually uniform palette

# Highlight the largest contiguous group
# (Optional: you can make these points larger or a different shape)
plot_scotoma <- plot_scotoma +
  geom_sf(data = largest_group, color = "black", size = 5) +
  theme(legend.position="none")

# Print the plot
#print(plot_all_points)

rttmp <- dBdat %>%
  filter(id=="RL35", visit==2, eye=="R", device=="retinalogik") %>%
  select(id, eye, x, y, dB)

hfatmp <- largest_group %>%
  mutate(x = st_coordinates(.)[, "X"],
         y = st_coordinates(.)[, "Y"]) %>%
  st_drop_geometry() 

mergeddat <- left_join(rttmp, hfatmp, by=c("id","eye","x","y")) %>%
  # remove blindspots and locations directly above and below blindspots
  # remove locations directly above and below blindspots
  filter(!(x==15 & y==3) & !(x==15 & y==-3) & !(x==15 & y==9) & !(x==15 & y==-9))

mergeddat <- mergeddat %>%
  merge(my_data, by=c("id","eye","x","y","psd_p"))

rtplot <- mergeddat %>%
  ggplot(aes(x, y, psd_p, dB.x)) +
  geom_raster(aes(x = x, y = y, fill = factor(psd_p))) +
  geom_text(aes(label = dB.x, x = x, y = y), size = 3) +
  coord_fixed(ratio = 1) +
  scale_fill_manual(values = rev(brewer.pal(3, "YlOrRd")), na.value="gray") +
  theme_bw() +
  ggtitle("Retinalogik", subtitle = "Threshold sensitivity dB values") +
  theme(legend.position="none")

hfaplot <- mergeddat %>%
  ggplot(aes(x, y, psd_p, dB)) +
  geom_raster(aes(x = x, y = y, fill = factor(psd_p))) +
  geom_text(aes(label = dB, x = x, y = y), size = 3) +
  coord_fixed(ratio = 1) +
  scale_fill_manual(values = rev(brewer.pal(3, "YlOrRd")), na.value="gray") +
  theme_bw() +
  ggtitle("HFA", subtitle = "Threshold sensitivity dB values") +
  theme(legend.position="none")

# mergeddat %>%
#   ggplot(aes(x, y, psd_p)) +
#   geom_raster(aes(x = x, y = y, fill = factor(psd_p))) +
#   geom_text(aes(label = psd_p, x = x, y = y), size = 3) +
#   coord_fixed(ratio = 1) +
#   scale_fill_manual(values = customcolors, na.value="gray") +
#   theme_bw() +
#   ggtitle("Retinalogik", subtitle = "Threshold sensitivity dB values") +
#   theme(legend.position="none")

#(p + plot_scotoma) / (hfaplot + rtplot)

# 1. Data Preparation and Binary Grid Creation
my_data_processed <- my_data %>%
  mutate(psd_p_clean = ifelse(is.na(psd_p), 100, psd_p))

# Convert the data frame to an sf (simple features) object
points_sf <- st_as_sf(my_data_processed, coords = c("x", "y"), crs = NA)

# 2. Filtering Defective Points
# Filter for points where psd_p is less than 5% (including 0.5% and 5.0%)
filtered_points_sf <- points_sf %>%
  filter(psd_p_clean < 5)

# Ensure there are defective points to analyze
if (nrow(filtered_points_sf) > 0) {
  # 3. Connected Components Labeling using Distance-Based Neighbors
  # Create a spatial weights matrix based on distance for the filtered points
  # d1 = 0: minimum distance (points must be distinct)
  # d2 = 9: maximum distance (captures vertical, horizontal, and diagonal neighbors
  #         given your coordinate spacing of 6 units, sqrt(6^2 + 6^2) approx 8.485)
  filtered_neighbors <- dnearneigh(filtered_points_sf, d1 = 0, d2 = 9)

  # Check if there are any connections among the filtered points
  if (length(unlist(filtered_neighbors)) > 0) {
    # Find connected components (contiguous groups)
    components <- n.comp.nb(filtered_neighbors)

    # Assign component IDs back to the filtered data
    filtered_points_sf$component_id <- components$comp.id

    # 4. Filtering Connected Components by Scotoma Criteria
    scotomas_found <- list()
    unique_component_ids <- unique(filtered_points_sf$component_id)

    for (comp_id in unique_component_ids) {
      current_cluster <- filtered_points_sf %>%
        filter(component_id == comp_id)

      num_points <- nrow(current_cluster)
      # Check if any point in this cluster has psd_p < 1% (i.e., psd_p_clean == 0.5)
      has_p_lt_1_percent <- any(current_cluster$psd_p_clean < 1)

      # Apply scotoma definition: at least 3 points AND at least 1 location at p < 1%
      # Apply scotoma definition: at least 3 points AND at least 1 location at p < 1%
      if (num_points >= 3 && has_p_lt_1_percent) {
        scotomas_found[[as.character(comp_id)]] <- current_cluster %>%
          # Extract x and y coordinates from the geometry column
          mutate(
            x_coord = st_coordinates(.)[, "X"],
            y_coord = st_coordinates(.)[, "Y"]
          ) %>%
          # Now select the desired columns, including the newly extracted x_coord and y_coord
          select(x_coord, y_coord, psd_p) %>%
          # Drop the geometry column as it's no longer needed for the table output
          st_drop_geometry() %>%
          # Rename x_coord and y_coord back to x and y for consistency with the table format
          rename(x = x_coord, y = y_coord)
      }
    }

    # Prepare results for the report table
    scotoma_results_list <- list()
    if (length(scotomas_found) > 0) {
      scotoma_id_counter <- 1
      for (scotoma_label in names(scotomas_found)) {
        scotoma_data <- scotomas_found[[scotoma_label]]
        num_points <- nrow(scotoma_data)
        min_psd_p_actual <- min(scotoma_data$psd_p, na.rm = TRUE) # Use original psd_p for reporting

        # Format constituent points string
        constituent_points_str <- paste(
          paste0("(", scotoma_data$x, ",", scotoma_data$y, ") [p=", scotoma_data$psd_p, "%]"),
          collapse = "; "
        )

        scotoma_results_list[[scotoma_id_counter]] <- data.frame(
          Scotoma_ID = scotoma_id_counter,
          Number_of_Points = num_points,
          Minimum_psd_p_percent = min_psd_p_actual,
          Constituent_Points = constituent_points_str,
          stringsAsFactors = FALSE
        )
        scotoma_id_counter <- scotoma_id_counter + 1
      }
    }

    # Final table of identified scotomas
    identified_scotomas_table_df <- if (length(scotoma_results_list) > 0) {
      do.call(rbind, scotoma_results_list)
    } else {
      data.frame(
        Scotoma_ID = integer(),
        Number_of_Points = integer(),
        Minimum_psd_p_percent = numeric(),
        Constituent_Points = character(),
        stringsAsFactors = FALSE
      )
    }

  } else {
    identified_scotomas_table_df <- data.frame(
      Scotoma_ID = integer(),
      Number_of_Points = integer(),
      Minimum_psd_p_percent = numeric(),
      Constituent_Points = character(),
      stringsAsFactors = FALSE
    )
  }
} else {
  identified_scotomas_table_df <- data.frame(
    Scotoma_ID = integer(),
    Number_of_Points = integer(),
    Minimum_psd_p_percent = numeric(),
    Constituent_Points = character(),
    stringsAsFactors = FALSE
  )
}

# Print the table (for verification within the R environment)
print(identified_scotomas_table_df)

# Define custom colors for psd_p values for the background raster
customcolors <- brewer.pal(4, "YlOrRd")
customcolors <- rev(customcolors)

# Plot 1: HFA Pattern Deviation Probability Plot (similar to user's 'p' plot)
# This plot shows the psd_p values for all test points.
p_plot <- my_data_processed %>%
  ggplot(aes(x, y)) +
  geom_raster(aes(fill = factor(psd_p))) + # Use original psd_p for fill
  geom_text(aes(label = ifelse(is.na(psd_p), "", psd_p)), size = 3) +
  coord_fixed(ratio = 1) +
  scale_fill_manual(values = customcolors, na.value = "gray") +
  theme_bw() +
  ggtitle("HFA", subtitle = "PSD p-values") +
  theme(legend.position = "none")

dB_plot <- my_data_processed %>%
  ggplot(aes(x, y)) +
  geom_raster(aes(fill = factor(psd_p))) + # Use original psd_p for fill
  geom_text(aes(label = ifelse(is.na(dB), "", dB)), size = 3) +
  coord_fixed(ratio = 1) +
  scale_fill_manual(values = customcolors, na.value = "gray") +
  theme_bw() +
  ggtitle("HFA", subtitle = "Threshold sensitivity dB values") +
  theme(legend.position = "none")

rttmp_filtered_for_plot <- rttmp %>%
  semi_join(my_data_processed, by = c("x", "y"))

# Join filtered_rttmp with my_data_processed to get psd_p for coloring the background
rttmp_for_plot_with_psdp <- rttmp_filtered_for_plot %>%
  left_join(my_data_processed %>% select(x, y, psd_p), by = c("x", "y"))

# Plot 4: Retinalogik Threshold Sensitivity dB values
rt_plot <- rttmp_for_plot_with_psdp %>%
  ggplot(aes(x, y)) +
  geom_raster(aes(fill = factor(psd_p))) + # Use HFA psd_p for fill as per user's example
  geom_text(aes(label = ifelse(is.na(dB), "", dB)), size = 3) + # Use Retinalogik dB for label
  coord_fixed(ratio = 1) +
  scale_fill_manual(values = customcolors, na.value = "gray") +
  theme_bw() +
  ggtitle("Retinalogik", subtitle = "Threshold sensitivity dB values") +
  theme(legend.position = "none")

# Prepare data for the scotoma highlight plot
# Combine all identified scotoma points into a single data frame for plotting
all_scotoma_points_for_plot <- data.frame()
if (length(scotomas_found) > 0) {
  scotoma_id_map <- 1
  for (original_comp_id in names(scotomas_found)) {
    temp_df <- scotomas_found[[original_comp_id]]
    temp_df$scotoma_plot_id <- scotoma_id_map # Assign a sequential ID for plotting colors
    all_scotoma_points_for_plot <- bind_rows(all_scotoma_points_for_plot, temp_df)
    scotoma_id_map <- scotoma_id_map + 1
  }
}

# Plot 2: Identified Scotomas Highlighted (similar to user's 'plot_scotoma')
# This plot overlays the identified scotoma points on a grey background of all points.
plot_scotoma_highlight <- ggplot() +
  geom_point(data = my_data_processed, aes(x = x, y = y), color = "grey", size = 1, alpha = 0.5) + # All points as background
  {
    if (nrow(all_scotoma_points_for_plot) > 0) {
      list(
        geom_point(data = all_scotoma_points_for_plot,
                   aes(x = x, y = y, color = as.factor(scotoma_plot_id)),
                   size = 3, shape = 16), # Highlight scotoma points
        scale_color_viridis_d(option = "viridis", name = "Scotoma ID"), # Use viridis for distinct scotoma colors
        geom_point(data = all_scotoma_points_for_plot,
                   aes(x = x, y = y),
                   color = "black", size = 5, shape = 1) # Add black outline for emphasis
      )
    }
  } +
  labs(title = "Identified Scotomas",
       x = "x", y = "y") +
  theme_minimal() +
  theme(legend.position = "right")

# To display these plots in your R environment, you would use:
# print(p_plot)
# print(plot_scotoma_highlight)
# If you have the 'patchwork' package installed, you can combine them like:
# library(patchwork)
(p_plot + plot_scotoma_highlight)/
  (dB_plot + rt_plot)

```

#### Visit 3
```{r scotoma3, warning=FALSE, fig.width=7.5, fig.height=7.5}
my_data <- psdpvalues_hfa %>%
  filter(datetime==ymd_hms("2025-05-23 10:00:11") & eye=="R")

rttmp <- dBdat %>%
  filter(id=="RL35", visit==3, eye=="R", device=="retinalogik") %>%
  select(id, eye, x, y, dB)

# 1. Data Preparation and Binary Grid Creation
my_data_processed <- my_data %>%
  mutate(psd_p_clean = ifelse(is.na(psd_p), 100, psd_p))

# Convert the data frame to an sf (simple features) object
points_sf <- st_as_sf(my_data_processed, coords = c("x", "y"), crs = NA)

# 2. Filtering Defective Points
# Filter for points where psd_p is less than 5% (including 0.5% and 5.0%)
filtered_points_sf <- points_sf %>%
  filter(psd_p_clean < 5)

# Ensure there are defective points to analyze
if (nrow(filtered_points_sf) > 0) {
  # 3. Connected Components Labeling using Distance-Based Neighbors
  # Create a spatial weights matrix based on distance for the filtered points
  # d1 = 0: minimum distance (points must be distinct)
  # d2 = 9: maximum distance (captures vertical, horizontal, and diagonal neighbors
  #         given your coordinate spacing of 6 units, sqrt(6^2 + 6^2) approx 8.485)
  filtered_neighbors <- dnearneigh(filtered_points_sf, d1 = 0, d2 = 9)

  # Check if there are any connections among the filtered points
  if (length(unlist(filtered_neighbors)) > 0) {
    # Find connected components (contiguous groups)
    components <- n.comp.nb(filtered_neighbors)

    # Assign component IDs back to the filtered data
    filtered_points_sf$component_id <- components$comp.id

    # 4. Filtering Connected Components by Scotoma Criteria
    scotomas_found <- list()
    unique_component_ids <- unique(filtered_points_sf$component_id)

    for (comp_id in unique_component_ids) {
      current_cluster <- filtered_points_sf %>%
        filter(component_id == comp_id)

      num_points <- nrow(current_cluster)
      # Check if any point in this cluster has psd_p < 1% (i.e., psd_p_clean == 0.5)
      has_p_lt_1_percent <- any(current_cluster$psd_p_clean < 1)

      # Apply scotoma definition: at least 3 points AND at least 1 location at p < 1%
      # Apply scotoma definition: at least 3 points AND at least 1 location at p < 1%
      if (num_points >= 3 && has_p_lt_1_percent) {
        scotomas_found[[as.character(comp_id)]] <- current_cluster %>%
          # Extract x and y coordinates from the geometry column
          mutate(
            x_coord = st_coordinates(.)[, "X"],
            y_coord = st_coordinates(.)[, "Y"]
          ) %>%
          # Now select the desired columns, including the newly extracted x_coord and y_coord
          select(x_coord, y_coord, psd_p) %>%
          # Drop the geometry column as it's no longer needed for the table output
          st_drop_geometry() %>%
          # Rename x_coord and y_coord back to x and y for consistency with the table format
          rename(x = x_coord, y = y_coord)
      }
    }

    # Prepare results for the report table
    scotoma_results_list <- list()
    if (length(scotomas_found) > 0) {
      scotoma_id_counter <- 1
      for (scotoma_label in names(scotomas_found)) {
        scotoma_data <- scotomas_found[[scotoma_label]]
        num_points <- nrow(scotoma_data)
        min_psd_p_actual <- min(scotoma_data$psd_p, na.rm = TRUE) # Use original psd_p for reporting

        # Format constituent points string
        constituent_points_str <- paste(
          paste0("(", scotoma_data$x, ",", scotoma_data$y, ") [p=", scotoma_data$psd_p, "%]"),
          collapse = "; "
        )

        scotoma_results_list[[scotoma_id_counter]] <- data.frame(
          Scotoma_ID = scotoma_id_counter,
          Number_of_Points = num_points,
          Minimum_psd_p_percent = min_psd_p_actual,
          Constituent_Points = constituent_points_str,
          stringsAsFactors = FALSE
        )
        scotoma_id_counter <- scotoma_id_counter + 1
      }
    }

    # Final table of identified scotomas
    identified_scotomas_table_df <- if (length(scotoma_results_list) > 0) {
      do.call(rbind, scotoma_results_list)
    } else {
      data.frame(
        Scotoma_ID = integer(),
        Number_of_Points = integer(),
        Minimum_psd_p_percent = numeric(),
        Constituent_Points = character(),
        stringsAsFactors = FALSE
      )
    }

  } else {
    identified_scotomas_table_df <- data.frame(
      Scotoma_ID = integer(),
      Number_of_Points = integer(),
      Minimum_psd_p_percent = numeric(),
      Constituent_Points = character(),
      stringsAsFactors = FALSE
    )
  }
} else {
  identified_scotomas_table_df <- data.frame(
    Scotoma_ID = integer(),
    Number_of_Points = integer(),
    Minimum_psd_p_percent = numeric(),
    Constituent_Points = character(),
    stringsAsFactors = FALSE
  )
}

# Print the table (for verification within the R environment)
print(identified_scotomas_table_df)

# Define custom colors for psd_p values for the background raster
customcolors <- brewer.pal(4, "YlOrRd")
customcolors <- rev(customcolors)

# Plot 1: HFA Pattern Deviation Probability Plot (similar to user's 'p' plot)
# This plot shows the psd_p values for all test points.
p_plot <- my_data_processed %>%
  ggplot(aes(x, y)) +
  geom_raster(aes(fill = factor(psd_p))) + # Use original psd_p for fill
  geom_text(aes(label = ifelse(is.na(psd_p), "", psd_p)), size = 3) +
  coord_fixed(ratio = 1) +
  scale_fill_manual(values = customcolors, na.value = "gray") +
  theme_bw() +
  ggtitle("HFA", subtitle = "PSD p-values") +
  theme(legend.position = "none")

dB_plot <- my_data_processed %>%
  ggplot(aes(x, y)) +
  geom_raster(aes(fill = factor(psd_p))) + # Use original psd_p for fill
  geom_text(aes(label = ifelse(is.na(dB), "", dB)), size = 3) +
  coord_fixed(ratio = 1) +
  scale_fill_manual(values = customcolors, na.value = "gray") +
  theme_bw() +
  ggtitle("HFA", subtitle = "Threshold sensitivity dB values") +
  theme(legend.position = "none")

rttmp_filtered_for_plot <- rttmp %>%
  semi_join(my_data_processed, by = c("x", "y"))

# Join filtered_rttmp with my_data_processed to get psd_p for coloring the background
rttmp_for_plot_with_psdp <- rttmp_filtered_for_plot %>%
  left_join(my_data_processed %>% select(x, y, psd_p), by = c("x", "y"))

# Plot 4: Retinalogik Threshold Sensitivity dB values
rt_plot <- rttmp_for_plot_with_psdp %>%
  ggplot(aes(x, y)) +
  geom_raster(aes(fill = factor(psd_p))) + # Use HFA psd_p for fill as per user's example
  geom_text(aes(label = ifelse(is.na(dB), "", dB)), size = 3) + # Use Retinalogik dB for label
  coord_fixed(ratio = 1) +
  scale_fill_manual(values = customcolors, na.value = "gray") +
  theme_bw() +
  ggtitle("Retinalogik", subtitle = "Threshold sensitivity dB values") +
  theme(legend.position = "none")

# Prepare data for the scotoma highlight plot
# Combine all identified scotoma points into a single data frame for plotting
all_scotoma_points_for_plot <- data.frame()
if (length(scotomas_found) > 0) {
  scotoma_id_map <- 1
  for (original_comp_id in names(scotomas_found)) {
    temp_df <- scotomas_found[[original_comp_id]]
    temp_df$scotoma_plot_id <- scotoma_id_map # Assign a sequential ID for plotting colors
    all_scotoma_points_for_plot <- bind_rows(all_scotoma_points_for_plot, temp_df)
    scotoma_id_map <- scotoma_id_map + 1
  }
}

# Plot 2: Identified Scotomas Highlighted (similar to user's 'plot_scotoma')
# This plot overlays the identified scotoma points on a grey background of all points.
plot_scotoma_highlight <- ggplot() +
  geom_point(data = my_data_processed, aes(x = x, y = y), color = "grey", size = 1, alpha = 0.5) + # All points as background
  {
    if (nrow(all_scotoma_points_for_plot) > 0) {
      list(
        geom_point(data = all_scotoma_points_for_plot,
                   aes(x = x, y = y, color = as.factor(scotoma_plot_id)),
                   size = 3, shape = 16), # Highlight scotoma points
        scale_color_viridis_d(option = "viridis", name = "Scotoma ID"), # Use viridis for distinct scotoma colors
        geom_point(data = all_scotoma_points_for_plot,
                   aes(x = x, y = y),
                   color = "black", size = 5, shape = 1) # Add black outline for emphasis
      )
    }
  } +
  labs(title = "Identified Scotomas",
       x = "x", y = "y") +
  theme_minimal() +
  theme(legend.position = "right")

(p_plot + plot_scotoma_highlight)/
  (dB_plot + rt_plot)

```

#### Visit 4
```{r scotoma4, warning=FALSE, fig.width=7.5, fig.height=7.5}
my_data <- psdpvalues_hfa %>%
  filter(datetime==ymd_hms("2025-05-27 10:07:48") & eye=="R")

rttmp <- dBdat %>%
  filter(id=="RL35", visit==4, eye=="R", device=="retinalogik") %>%
  select(id, eye, x, y, dB)

# 1. Data Preparation and Binary Grid Creation
my_data_processed <- my_data %>%
  mutate(psd_p_clean = ifelse(is.na(psd_p), 100, psd_p))

# Convert the data frame to an sf (simple features) object
points_sf <- st_as_sf(my_data_processed, coords = c("x", "y"), crs = NA)

# 2. Filtering Defective Points
# Filter for points where psd_p is less than 5% (including 0.5% and 5.0%)
filtered_points_sf <- points_sf %>%
  filter(psd_p_clean < 5)

# Ensure there are defective points to analyze
if (nrow(filtered_points_sf) > 0) {
  # 3. Connected Components Labeling using Distance-Based Neighbors
  # Create a spatial weights matrix based on distance for the filtered points
  # d1 = 0: minimum distance (points must be distinct)
  # d2 = 9: maximum distance (captures vertical, horizontal, and diagonal neighbors
  #         given your coordinate spacing of 6 units, sqrt(6^2 + 6^2) approx 8.485)
  filtered_neighbors <- dnearneigh(filtered_points_sf, d1 = 0, d2 = 9)

  # Check if there are any connections among the filtered points
  if (length(unlist(filtered_neighbors)) > 0) {
    # Find connected components (contiguous groups)
    components <- n.comp.nb(filtered_neighbors)

    # Assign component IDs back to the filtered data
    filtered_points_sf$component_id <- components$comp.id

    # 4. Filtering Connected Components by Scotoma Criteria
    scotomas_found <- list()
    unique_component_ids <- unique(filtered_points_sf$component_id)

    for (comp_id in unique_component_ids) {
      current_cluster <- filtered_points_sf %>%
        filter(component_id == comp_id)

      num_points <- nrow(current_cluster)
      # Check if any point in this cluster has psd_p < 1% (i.e., psd_p_clean == 0.5)
      has_p_lt_1_percent <- any(current_cluster$psd_p_clean < 1)

      # Apply scotoma definition: at least 3 points AND at least 1 location at p < 1%
      # Apply scotoma definition: at least 3 points AND at least 1 location at p < 1%
      if (num_points >= 3 && has_p_lt_1_percent) {
        scotomas_found[[as.character(comp_id)]] <- current_cluster %>%
          # Extract x and y coordinates from the geometry column
          mutate(
            x_coord = st_coordinates(.)[, "X"],
            y_coord = st_coordinates(.)[, "Y"]
          ) %>%
          # Now select the desired columns, including the newly extracted x_coord and y_coord
          select(x_coord, y_coord, psd_p) %>%
          # Drop the geometry column as it's no longer needed for the table output
          st_drop_geometry() %>%
          # Rename x_coord and y_coord back to x and y for consistency with the table format
          rename(x = x_coord, y = y_coord)
      }
    }

    # Prepare results for the report table
    scotoma_results_list <- list()
    if (length(scotomas_found) > 0) {
      scotoma_id_counter <- 1
      for (scotoma_label in names(scotomas_found)) {
        scotoma_data <- scotomas_found[[scotoma_label]]
        num_points <- nrow(scotoma_data)
        min_psd_p_actual <- min(scotoma_data$psd_p, na.rm = TRUE) # Use original psd_p for reporting

        # Format constituent points string
        constituent_points_str <- paste(
          paste0("(", scotoma_data$x, ",", scotoma_data$y, ") [p=", scotoma_data$psd_p, "%]"),
          collapse = "; "
        )

        scotoma_results_list[[scotoma_id_counter]] <- data.frame(
          Scotoma_ID = scotoma_id_counter,
          Number_of_Points = num_points,
          Minimum_psd_p_percent = min_psd_p_actual,
          Constituent_Points = constituent_points_str,
          stringsAsFactors = FALSE
        )
        scotoma_id_counter <- scotoma_id_counter + 1
      }
    }

    # Final table of identified scotomas
    identified_scotomas_table_df <- if (length(scotoma_results_list) > 0) {
      do.call(rbind, scotoma_results_list)
    } else {
      data.frame(
        Scotoma_ID = integer(),
        Number_of_Points = integer(),
        Minimum_psd_p_percent = numeric(),
        Constituent_Points = character(),
        stringsAsFactors = FALSE
      )
    }

  } else {
    identified_scotomas_table_df <- data.frame(
      Scotoma_ID = integer(),
      Number_of_Points = integer(),
      Minimum_psd_p_percent = numeric(),
      Constituent_Points = character(),
      stringsAsFactors = FALSE
    )
  }
} else {
  identified_scotomas_table_df <- data.frame(
    Scotoma_ID = integer(),
    Number_of_Points = integer(),
    Minimum_psd_p_percent = numeric(),
    Constituent_Points = character(),
    stringsAsFactors = FALSE
  )
}

# Print the table (for verification within the R environment)
print(identified_scotomas_table_df)

# Define custom colors for psd_p values for the background raster
customcolors <- brewer.pal(4, "YlOrRd")
customcolors <- rev(customcolors)

# Plot 1: HFA Pattern Deviation Probability Plot (similar to user's 'p' plot)
# This plot shows the psd_p values for all test points.
p_plot <- my_data_processed %>%
  ggplot(aes(x, y)) +
  geom_raster(aes(fill = factor(psd_p))) + # Use original psd_p for fill
  geom_text(aes(label = ifelse(is.na(psd_p), "", psd_p)), size = 3) +
  coord_fixed(ratio = 1) +
  scale_fill_manual(values = customcolors, na.value = "gray") +
  theme_bw() +
  ggtitle("HFA", subtitle = "PSD p-values") +
  theme(legend.position = "none")

dB_plot <- my_data_processed %>%
  ggplot(aes(x, y)) +
  geom_raster(aes(fill = factor(psd_p))) + # Use original psd_p for fill
  geom_text(aes(label = ifelse(is.na(dB), "", dB)), size = 3) +
  coord_fixed(ratio = 1) +
  scale_fill_manual(values = customcolors, na.value = "gray") +
  theme_bw() +
  ggtitle("HFA", subtitle = "Threshold sensitivity dB values") +
  theme(legend.position = "none")

rttmp_filtered_for_plot <- rttmp %>%
  semi_join(my_data_processed, by = c("x", "y"))

# Join filtered_rttmp with my_data_processed to get psd_p for coloring the background
rttmp_for_plot_with_psdp <- rttmp_filtered_for_plot %>%
  left_join(my_data_processed %>% select(x, y, psd_p), by = c("x", "y"))

# Plot 4: Retinalogik Threshold Sensitivity dB values
rt_plot <- rttmp_for_plot_with_psdp %>%
  ggplot(aes(x, y)) +
  geom_raster(aes(fill = factor(psd_p))) + # Use HFA psd_p for fill as per user's example
  geom_text(aes(label = ifelse(is.na(dB), "", dB)), size = 3) + # Use Retinalogik dB for label
  coord_fixed(ratio = 1) +
  scale_fill_manual(values = customcolors, na.value = "gray") +
  theme_bw() +
  ggtitle("Retinalogik", subtitle = "Threshold sensitivity dB values") +
  theme(legend.position = "none")

# Prepare data for the scotoma highlight plot
# Combine all identified scotoma points into a single data frame for plotting
all_scotoma_points_for_plot <- data.frame()
if (length(scotomas_found) > 0) {
  scotoma_id_map <- 1
  for (original_comp_id in names(scotomas_found)) {
    temp_df <- scotomas_found[[original_comp_id]]
    temp_df$scotoma_plot_id <- scotoma_id_map # Assign a sequential ID for plotting colors
    all_scotoma_points_for_plot <- bind_rows(all_scotoma_points_for_plot, temp_df)
    scotoma_id_map <- scotoma_id_map + 1
  }
}

# Plot 2: Identified Scotomas Highlighted (similar to user's 'plot_scotoma')
# This plot overlays the identified scotoma points on a grey background of all points.
plot_scotoma_highlight <- ggplot() +
  geom_point(data = my_data_processed, aes(x = x, y = y), color = "grey", size = 1, alpha = 0.5) + # All points as background
  {
    if (nrow(all_scotoma_points_for_plot) > 0) {
      list(
        geom_point(data = all_scotoma_points_for_plot,
                   aes(x = x, y = y, color = as.factor(scotoma_plot_id)),
                   size = 3, shape = 16), # Highlight scotoma points
        scale_color_viridis_d(option = "viridis", name = "Scotoma ID"), # Use viridis for distinct scotoma colors
        geom_point(data = all_scotoma_points_for_plot,
                   aes(x = x, y = y),
                   color = "black", size = 5, shape = 1) # Add black outline for emphasis
      )
    }
  } +
  labs(title = "Identified Scotomas",
       x = "x", y = "y") +
  theme_minimal() +
  theme(legend.position = "right")

(p_plot + plot_scotoma_highlight)/
  (dB_plot + rt_plot)

```

#### Visit 5
```{r scotoma5, warning=FALSE, fig.width=7.5, fig.height=7.5}
my_data <- psdpvalues_hfa %>%
  filter(datetime==ymd_hms("2025-05-29 09:47:54") & eye=="R")

rttmp <- dBdat %>%
  filter(id=="RL35", visit==5, eye=="R", device=="retinalogik") %>%
  select(id, eye, x, y, dB)

# 1. Data Preparation and Binary Grid Creation
my_data_processed <- my_data %>%
  mutate(psd_p_clean = ifelse(is.na(psd_p), 100, psd_p))

# Convert the data frame to an sf (simple features) object
points_sf <- st_as_sf(my_data_processed, coords = c("x", "y"), crs = NA)

# 2. Filtering Defective Points
# Filter for points where psd_p is less than 5% (including 0.5% and 5.0%)
filtered_points_sf <- points_sf %>%
  filter(psd_p_clean < 5)

# Ensure there are defective points to analyze
if (nrow(filtered_points_sf) > 0) {
  # 3. Connected Components Labeling using Distance-Based Neighbors
  # Create a spatial weights matrix based on distance for the filtered points
  # d1 = 0: minimum distance (points must be distinct)
  # d2 = 9: maximum distance (captures vertical, horizontal, and diagonal neighbors
  #         given your coordinate spacing of 6 units, sqrt(6^2 + 6^2) approx 8.485)
  filtered_neighbors <- dnearneigh(filtered_points_sf, d1 = 0, d2 = 9)

  # Check if there are any connections among the filtered points
  if (length(unlist(filtered_neighbors)) > 0) {
    # Find connected components (contiguous groups)
    components <- n.comp.nb(filtered_neighbors)

    # Assign component IDs back to the filtered data
    filtered_points_sf$component_id <- components$comp.id

    # 4. Filtering Connected Components by Scotoma Criteria
    scotomas_found <- list()
    unique_component_ids <- unique(filtered_points_sf$component_id)

    for (comp_id in unique_component_ids) {
      current_cluster <- filtered_points_sf %>%
        filter(component_id == comp_id)

      num_points <- nrow(current_cluster)
      # Check if any point in this cluster has psd_p < 1% (i.e., psd_p_clean == 0.5)
      has_p_lt_1_percent <- any(current_cluster$psd_p_clean < 1)

      # Apply scotoma definition: at least 3 points AND at least 1 location at p < 1%
      # Apply scotoma definition: at least 3 points AND at least 1 location at p < 1%
      if (num_points >= 3 && has_p_lt_1_percent) {
        scotomas_found[[as.character(comp_id)]] <- current_cluster %>%
          # Extract x and y coordinates from the geometry column
          mutate(
            x_coord = st_coordinates(.)[, "X"],
            y_coord = st_coordinates(.)[, "Y"]
          ) %>%
          # Now select the desired columns, including the newly extracted x_coord and y_coord
          select(x_coord, y_coord, psd_p) %>%
          # Drop the geometry column as it's no longer needed for the table output
          st_drop_geometry() %>%
          # Rename x_coord and y_coord back to x and y for consistency with the table format
          rename(x = x_coord, y = y_coord)
      }
    }

    # Prepare results for the report table
    scotoma_results_list <- list()
    if (length(scotomas_found) > 0) {
      scotoma_id_counter <- 1
      for (scotoma_label in names(scotomas_found)) {
        scotoma_data <- scotomas_found[[scotoma_label]]
        num_points <- nrow(scotoma_data)
        min_psd_p_actual <- min(scotoma_data$psd_p, na.rm = TRUE) # Use original psd_p for reporting

        # Format constituent points string
        constituent_points_str <- paste(
          paste0("(", scotoma_data$x, ",", scotoma_data$y, ") [p=", scotoma_data$psd_p, "%]"),
          collapse = "; "
        )

        scotoma_results_list[[scotoma_id_counter]] <- data.frame(
          Scotoma_ID = scotoma_id_counter,
          Number_of_Points = num_points,
          Minimum_psd_p_percent = min_psd_p_actual,
          Constituent_Points = constituent_points_str,
          stringsAsFactors = FALSE
        )
        scotoma_id_counter <- scotoma_id_counter + 1
      }
    }

    # Final table of identified scotomas
    identified_scotomas_table_df <- if (length(scotoma_results_list) > 0) {
      do.call(rbind, scotoma_results_list)
    } else {
      data.frame(
        Scotoma_ID = integer(),
        Number_of_Points = integer(),
        Minimum_psd_p_percent = numeric(),
        Constituent_Points = character(),
        stringsAsFactors = FALSE
      )
    }

  } else {
    identified_scotomas_table_df <- data.frame(
      Scotoma_ID = integer(),
      Number_of_Points = integer(),
      Minimum_psd_p_percent = numeric(),
      Constituent_Points = character(),
      stringsAsFactors = FALSE
    )
  }
} else {
  identified_scotomas_table_df <- data.frame(
    Scotoma_ID = integer(),
    Number_of_Points = integer(),
    Minimum_psd_p_percent = numeric(),
    Constituent_Points = character(),
    stringsAsFactors = FALSE
  )
}

# Print the table (for verification within the R environment)
print(identified_scotomas_table_df)

# Define custom colors for psd_p values for the background raster
customcolors <- brewer.pal(4, "YlOrRd")
customcolors <- rev(customcolors)

# Plot 1: HFA Pattern Deviation Probability Plot (similar to user's 'p' plot)
# This plot shows the psd_p values for all test points.
p_plot <- my_data_processed %>%
  ggplot(aes(x, y)) +
  geom_raster(aes(fill = factor(psd_p))) + # Use original psd_p for fill
  geom_text(aes(label = ifelse(is.na(psd_p), "", psd_p)), size = 3) +
  coord_fixed(ratio = 1) +
  scale_fill_manual(values = customcolors, na.value = "gray") +
  theme_bw() +
  ggtitle("HFA", subtitle = "PSD p-values") +
  theme(legend.position = "none")

dB_plot <- my_data_processed %>%
  ggplot(aes(x, y)) +
  geom_raster(aes(fill = factor(psd_p))) + # Use original psd_p for fill
  geom_text(aes(label = ifelse(is.na(dB), "", dB)), size = 3) +
  coord_fixed(ratio = 1) +
  scale_fill_manual(values = customcolors, na.value = "gray") +
  theme_bw() +
  ggtitle("HFA", subtitle = "Threshold sensitivity dB values") +
  theme(legend.position = "none")

rttmp_filtered_for_plot <- rttmp %>%
  semi_join(my_data_processed, by = c("x", "y"))

# Join filtered_rttmp with my_data_processed to get psd_p for coloring the background
rttmp_for_plot_with_psdp <- rttmp_filtered_for_plot %>%
  left_join(my_data_processed %>% select(x, y, psd_p), by = c("x", "y"))

# Plot 4: Retinalogik Threshold Sensitivity dB values
rt_plot <- rttmp_for_plot_with_psdp %>%
  ggplot(aes(x, y)) +
  geom_raster(aes(fill = factor(psd_p))) + # Use HFA psd_p for fill as per user's example
  geom_text(aes(label = ifelse(is.na(dB), "", dB)), size = 3) + # Use Retinalogik dB for label
  coord_fixed(ratio = 1) +
  scale_fill_manual(values = customcolors, na.value = "gray") +
  theme_bw() +
  ggtitle("Retinalogik", subtitle = "Threshold sensitivity dB values") +
  theme(legend.position = "none")

# Prepare data for the scotoma highlight plot
# Combine all identified scotoma points into a single data frame for plotting
all_scotoma_points_for_plot <- data.frame()
if (length(scotomas_found) > 0) {
  scotoma_id_map <- 1
  for (original_comp_id in names(scotomas_found)) {
    temp_df <- scotomas_found[[original_comp_id]]
    temp_df$scotoma_plot_id <- scotoma_id_map # Assign a sequential ID for plotting colors
    all_scotoma_points_for_plot <- bind_rows(all_scotoma_points_for_plot, temp_df)
    scotoma_id_map <- scotoma_id_map + 1
  }
}

# Plot 2: Identified Scotomas Highlighted (similar to user's 'plot_scotoma')
# This plot overlays the identified scotoma points on a grey background of all points.
plot_scotoma_highlight <- ggplot() +
  geom_point(data = my_data_processed, aes(x = x, y = y), color = "grey", size = 1, alpha = 0.5) + # All points as background
  {
    if (nrow(all_scotoma_points_for_plot) > 0) {
      list(
        geom_point(data = all_scotoma_points_for_plot,
                   aes(x = x, y = y, color = as.factor(scotoma_plot_id)),
                   size = 3, shape = 16), # Highlight scotoma points
        scale_color_viridis_d(option = "viridis", name = "Scotoma ID"), # Use viridis for distinct scotoma colors
        geom_point(data = all_scotoma_points_for_plot,
                   aes(x = x, y = y),
                   color = "black", size = 5, shape = 1) # Add black outline for emphasis
      )
    }
  } +
  labs(title = "Identified Scotomas",
       x = "x", y = "y") +
  theme_minimal() +
  theme(legend.position = "right")

(p_plot + plot_scotoma_highlight)/
  (dB_plot + rt_plot)

```
