---
title: "Retinalogik vs HFA study scotoma analysis"
author: "[Eadie Technologies, Inc.](https://www.eadietech.com/)"
date: "2025-06-04"
output: 
  html_document:
    theme: yeti
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: 
      collapsed: true
    includes:
        after_body: footer.html
---

## Setup

### Versioning

Last updated on `r format(Sys.time(), '%Y-%b-%d')` at `r format(Sys.time(), '%I:%M %p')`.

* 2025-Jun-04: Created separate .Rmd file as index.Rmd was getting very cluttered.

### Acknowledgements
Data collected by Research Assistants Luc St-Pierre, Reann Post, and Research Coordinator Lynn Murphy.

```{r setup, collapse=TRUE}

# load packages
library(tidyverse)
library(here)
library(DT)
library(plotly)
library(viridis)
library(svglite)
library(htmltools)
library(gridExtra)

sessionInfo()

# default chunk options
knitr::opts_chunk$set(
  comment = '>', collapse = TRUE, cache = FALSE, dev= c("png")
  )

# load processed data
load(here("dBdat.Rda"))

load(here("retinalogik_hfadatawpsd.Rda"))

```

# Study summary
We compared 24-2 visual field (VF) test results between Retinalogik and the Humphrey Field Analyzer (HFA) using SITA Standard test strategy in patients with early glaucoma and moderate to advanced glaucoma [here](https://vvnejl.github.io/retinalogik-study/). A scotoma analysis was performed following glaucomatous VF definitions found in [Park et al. (2011)](https://www.aaojournal.org/article/S0161-6420(11)00172-2/abstract).

# Methods (very briefly)
We define a scotoma as the presence of at least 3 vertically, horizontally, or diagonally contiguous test points within the same hemifield on the pattern deviation probability plot at p <5%, with at least 1 point at p <1%, excluding points directly above or below the blind spot. Eyes with consistent glaucomatous VF loss on at least 2 consecutive examinations with reliability indices better than 30% were included. 

## Analysis {.tabset}
### RL24
```{r scotoma, echo=FALSE}

```
