---
title: "Retinalogik vs HFA study scotoma analysis"
author: "[Eadie Technologies, Inc.](https://www.eadietech.com/)"
date: "2025-06-04"
output: 
  html_document:
    theme: yeti
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: 
      collapsed: true
    includes:
        after_body: footer.html
---

## Setup

### Versioning

Last updated on `r format(Sys.time(), '%Y-%b-%d')` at `r format(Sys.time(), '%I:%M %p')`.

* 2025-Jun-04: Created separate .Rmd file as index.Rmd was getting very cluttered.

### Acknowledgements
Data collected by Research Assistants Luc St-Pierre, Reann Post, and Research Coordinator Lynn Murphy.

```{r setup, collapse=TRUE}

# load packages
library(tidyverse)
library(here)
library(DT)
library(plotly)
library(viridis)
library(svglite)
library(htmltools)
library(gridExtra)
library(sf)
library(spdep)
library(RColorBrewer)
library(patchwork)

sessionInfo()

# default chunk options
knitr::opts_chunk$set(
  comment = '>', collapse = TRUE, cache = FALSE, dev= c("png")
  )

# load processed data
load(here("dBdat.Rda"))

load(here("retinalogik_hfadatawpsd.Rda"))

```

# Study summary
We compared 24-2 visual field (VF) test results between Retinalogik and the Humphrey Field Analyzer (HFA) using SITA Standard test strategy in patients with early glaucoma and moderate to advanced glaucoma [here](https://vvnejl.github.io/retinalogik-study/). A scotoma analysis was performed following glaucomatous VF definitions found in [Park et al. (2011)](https://www.aaojournal.org/article/S0161-6420(11)00172-2/abstract).

# Methods (very briefly)
We define a scotoma as the presence of at least 3 vertically, horizontally, or diagonally contiguous test points within the same hemifield on the pattern deviation probability plot at p <5%, with at least 1 point at p <1%, excluding points directly above or below the blind spot. Eyes with consistent glaucomatous VF loss on at least 2 consecutive examinations with reliability indices better than 30% were included. 

## Analysis {.tabset}
### RL35 {.tabset}
#### Visit 1
```{r scotoma, warning=FALSE, fig.width=7.5, fig.height=7.5}
# remove locations directly above and below blindspots
psdpvalues_hfa <- alldat %>%
  filter(!(x==15 & y==9) & !(x==15 & y==-9))

# remove reliability indices <30%
psdpvalues_hfa <- psdpvalues_hfa %>%
  filter((fl/flTest) < 0.3 & fnPerc < 0.3 & fpPerc < 0.3)

tmp <- psdpvalues_hfa %>%
  filter(id=="RL35" & eye=="R")

my_data <- psdpvalues_hfa %>%
  filter(datetime==ymd_hms("2025-05-20 09:48:31") & eye=="R")

customcolors <- brewer.pal(4, "YlOrRd")
customcolors <- rev(customcolors)

p <- my_data %>%
  ggplot(aes(x, y, psd_p)) +
  geom_raster(aes(x = x, y = y, fill = factor(psd_p))) +
  geom_text(aes(label = ifelse(is.na(psd_p), "", psd_p),
                x = x, y = y), size = 3) +
  coord_fixed(ratio = 1) +
  scale_fill_manual(values = customcolors, na.value="gray") +
  theme_bw() +
  ggtitle("HFA", subtitle = "PSD p-values") +
  theme(legend.position="none")

dBplot <- my_data %>%
  ggplot(aes(x, y, psd_p, dB)) +
  geom_raster(aes(x = x, y = y, fill = factor(psd_p))) +
  geom_text(aes(label = dB, x = x, y = y), size = 3) +
  coord_fixed(ratio = 1) +
  scale_fill_manual(values = customcolors, na.value="gray") +
  theme_bw() +
  ggtitle("HFA", subtitle = "Threshold sensitivity dB values") +
  theme(legend.position="none")
  
my_sf <- my_data %>%
  select(datetime, eye, id, x, y, dB, psd_p) %>%
  st_as_sf(coords = c("x", "y"), crs = NA)

# You can view the data
#plot(my_sf)

# Convert to sf object
# It's good practice to provide a CRS if your coordinates represent real-world locations.
# For simple relative coordinates, NA is fine.
points_sf <- st_as_sf(my_sf, wkt = "geometry", crs = NA)

# Create a spatial weights matrix based on distance for point data
# d1 = 0: minimum distance (points must be distinct)
# d2 = 9: maximum distance (captures vertical, horizontal, and diagonal neighbors
#         given your coordinate spacing of 6 units)
neighbors <- dnearneigh(points_sf, d1 = 0, d2 = 9)

# Convert the neighbor list to a spatial weights list (binary style is common)
weights_list <- nb2listw(neighbors, style = "B")

# Filter for points where psd_p <= 5
# Using is.finite() to handle NA values in psd_p, as NA <= 5 would also be NA, not FALSE.
filtered_points <- points_sf %>%
  filter(is.finite(psd_p) & psd_p <= 5)

if(nrow(filtered_points) > 0) {
  # Create a spatial weights matrix for the filtered points (re-evaluate neighbors)
  # This ensures we only consider contiguity among the *filtered* points.
  filtered_neighbors <- dnearneigh(filtered_points, d1 = 0, d2 = 9)
  
  # Check if there are any connections among the filtered points
  if (length(unlist(filtered_neighbors)) > 0) { # Check if any neighbors exist
    # Find connected components (contiguous groups)
    components <- n.comp.nb(filtered_neighbors)
    
    # Assign component IDs back to the filtered data
    filtered_points$component_id <- components$comp.id
    
    # Group by component and count points
    contiguous_groups <- filtered_points %>%
      group_by(component_id) %>%
      summarize(n_points = n()) %>%
      arrange(desc(n_points))
    
    # Print the results
    #print("Contiguous groups where psd_p <= 5:")
    #print(contiguous_groups)
    
    # Find the largest contiguous group
    largest_group_id <- contiguous_groups$component_id[1]
    
    largest_group <- filtered_points %>%
      filter(component_id == largest_group_id)
    
    #print("Largest Contiguous Group:")
    #print(largest_group)
    
    if (nrow(largest_group) >= 3) {
      #print("Largest contiguous group has at least 3 points.")
      # Optionally, extract and print the coordinates of the largest group
      #print("Coordinates of the largest contiguous group:")
      #print(st_coordinates(largest_group))
    } else {
      #print("No contiguous group with at least 3 points found that has at least 3 points.")
    }
  } else {
    #print("No contiguous points found after filtering for psd_p <= 5.")
  }
  
} else {
  #print("No points found where psd_p <= 5.")
}

# --- Visualization ---

# Plot all original points
# plot_all_points <- ggplot() +
#   geom_sf(data = points_sf, color = "grey", size = 1, alpha = 0.5) +
#   labs(title = "All Original Points (Grey) and Filtered Contiguous Groups (Colored)",
#        x = "X-coordinate", y = "Y-coordinate") +
#   theme_minimal()

plot_all_points <- ggplot() +
  geom_sf(data = points_sf, color = "grey", size = 1, alpha = 0.5) +
  labs(title = "Contiguous Groups",
       x = "x", y = "y") +
  theme_bw()

# Add the filtered and grouped points on top, colored by their component_id
# Use a specific color palette if you have many groups
plot_scotoma <- plot_all_points +
  geom_sf(data = filtered_points, aes(color = as.factor(component_id)), size = 3) +
  scale_color_viridis_d(option = "viridis", name = "Contiguous Group ID") # Viridis is a good perceptually uniform palette

# Highlight the largest contiguous group
# (Optional: you can make these points larger or a different shape)
plot_scotoma <- plot_scotoma +
  geom_sf(data = largest_group, color = "black", size = 5) +
  theme(legend.position="none")

# Print the plot
#print(plot_all_points)

rttmp <- dBdat %>%
  filter(id=="RL35", visit==1, eye=="R", device=="retinalogik") %>%
  select(id, eye, x, y, dB)

hfatmp <- largest_group %>%
  mutate(x = st_coordinates(.)[, "X"],
         y = st_coordinates(.)[, "Y"]) %>%
  st_drop_geometry() 

mergeddat <- left_join(rttmp, hfatmp, by=c("id","eye","x","y")) %>%
  # remove blindspots and locations directly above and below blindspots
  # remove locations directly above and below blindspots
  filter(!(x==15 & y==3) & !(x==15 & y==-3) & !(x==15 & y==9) & !(x==15 & y==-9))

rtplot <- mergeddat %>%
  ggplot(aes(x, y, psd_p, dB.x)) +
  geom_raster(aes(x = x, y = y, fill = factor(psd_p))) +
  geom_text(aes(label = dB.x, x = x, y = y), size = 3) +
  coord_fixed(ratio = 1) +
  scale_fill_manual(values = rev(brewer.pal(3, "YlOrRd")), na.value="gray") +
  theme_bw() +
  ggtitle("Retinalogik", subtitle = "Threshold sensitivity dB values") +
  theme(legend.position="none")

hfaplot <- mergeddat %>%
  ggplot(aes(x, y, psd_p, dB.y)) +
  geom_raster(aes(x = x, y = y, fill = factor(psd_p))) +
  geom_text(aes(label = dB.y, x = x, y = y), size = 3) +
  coord_fixed(ratio = 1) +
  scale_fill_manual(values = rev(brewer.pal(3, "YlOrRd")), na.value="gray") +
  theme_bw() +
  ggtitle("HFA", subtitle = "Threshold sensitivity dB values") +
  theme(legend.position="none")

# mergeddat %>%
#   ggplot(aes(x, y, psd_p)) +
#   geom_raster(aes(x = x, y = y, fill = factor(psd_p))) +
#   geom_text(aes(label = psd_p, x = x, y = y), size = 3) +
#   coord_fixed(ratio = 1) +
#   scale_fill_manual(values = customcolors, na.value="gray") +
#   theme_bw() +
#   ggtitle("Retinalogik", subtitle = "Threshold sensitivity dB values") +
#   theme(legend.position="none")

(p + plot_scotoma) / (hfaplot + rtplot)

```

#### Visit 2
```{r scotoma2, warning=FALSE, fig.width=7.5, fig.height=7.5}
tmp <- psdpvalues_hfa %>%
  filter(id=="35" & eye=="R")

my_data <- psdpvalues_hfa %>%
  filter(datetime==ymd_hms("2025-05-22 09:40:31") & eye=="R")

customcolors <- brewer.pal(4, "YlOrRd")
customcolors <- rev(customcolors)

p <- my_data %>%
  ggplot(aes(x, y, psd_p)) +
  geom_raster(aes(x = x, y = y, fill = factor(psd_p))) +
  geom_text(aes(label = ifelse(is.na(psd_p), "", psd_p),
                x = x, y = y), size = 3) +
  coord_fixed(ratio = 1) +
  scale_fill_manual(values = customcolors, na.value="gray") +
  theme_bw() +
  ggtitle("HFA", subtitle = "PSD p-values") +
  theme(legend.position="none")

dBplot <- my_data %>%
  ggplot(aes(x, y, psd_p, dB)) +
  geom_raster(aes(x = x, y = y, fill = factor(psd_p))) +
  geom_text(aes(label = dB, x = x, y = y), size = 3) +
  coord_fixed(ratio = 1) +
  scale_fill_manual(values = customcolors, na.value="gray") +
  theme_bw() +
  ggtitle("HFA", subtitle = "Threshold sensitivity dB values") +
  theme(legend.position="none")
  
my_sf <- my_data %>%
  select(datetime, eye, id, x, y, dB, psd_p) %>%
  st_as_sf(coords = c("x", "y"), crs = NA)

# You can view the data
#plot(my_sf)

# Convert to sf object
# It's good practice to provide a CRS if your coordinates represent real-world locations.
# For simple relative coordinates, NA is fine.
points_sf <- st_as_sf(my_sf, wkt = "geometry", crs = NA)

# Create a spatial weights matrix based on distance for point data
# d1 = 0: minimum distance (points must be distinct)
# d2 = 9: maximum distance (captures vertical, horizontal, and diagonal neighbors
#         given your coordinate spacing of 6 units)
neighbors <- dnearneigh(points_sf, d1 = 0, d2 = 9)

# Convert the neighbor list to a spatial weights list (binary style is common)
weights_list <- nb2listw(neighbors, style = "B")

# Filter for points where psd_p <= 5
# Using is.finite() to handle NA values in psd_p, as NA <= 5 would also be NA, not FALSE.
filtered_points <- points_sf %>%
  filter(is.finite(psd_p) & psd_p <= 5)

if(nrow(filtered_points) > 0) {
  # Create a spatial weights matrix for the filtered points (re-evaluate neighbors)
  # This ensures we only consider contiguity among the *filtered* points.
  filtered_neighbors <- dnearneigh(filtered_points, d1 = 0, d2 = 9)
  
  # Check if there are any connections among the filtered points
  if (length(unlist(filtered_neighbors)) > 0) { # Check if any neighbors exist
    # Find connected components (contiguous groups)
    components <- n.comp.nb(filtered_neighbors)
    
    # Assign component IDs back to the filtered data
    filtered_points$component_id <- components$comp.id
    
    # Group by component and count points
    contiguous_groups <- filtered_points %>%
      group_by(component_id) %>%
      summarize(n_points = n()) %>%
      arrange(desc(n_points))
    
    # Print the results
    #print("Contiguous groups where psd_p <= 5:")
    #print(contiguous_groups)
    
    # Find the largest contiguous group
    largest_group_id <- contiguous_groups$component_id[1]
    
    largest_group <- filtered_points %>%
      filter(component_id == largest_group_id)
    
    #print("Largest Contiguous Group:")
    #print(largest_group)
    
    if (nrow(largest_group) >= 3) {
      #print("Largest contiguous group has at least 3 points.")
      # Optionally, extract and print the coordinates of the largest group
      #print("Coordinates of the largest contiguous group:")
      #print(st_coordinates(largest_group))
    } else {
      #print("No contiguous group with at least 3 points found that has at least 3 points.")
    }
  } else {
    #print("No contiguous points found after filtering for psd_p <= 5.")
  }
  
} else {
  #print("No points found where psd_p <= 5.")
}

# --- Visualization ---

# Plot all original points
# plot_all_points <- ggplot() +
#   geom_sf(data = points_sf, color = "grey", size = 1, alpha = 0.5) +
#   labs(title = "All Original Points (Grey) and Filtered Contiguous Groups (Colored)",
#        x = "X-coordinate", y = "Y-coordinate") +
#   theme_minimal()

plot_all_points <- ggplot() +
  geom_sf(data = points_sf, color = "grey", size = 1, alpha = 0.5) +
  labs(title = "Contiguous Groups",
       x = "x", y = "y") +
  theme_bw()

# Add the filtered and grouped points on top, colored by their component_id
# Use a specific color palette if you have many groups
plot_scotoma <- plot_all_points +
  geom_sf(data = filtered_points, aes(color = as.factor(component_id)), size = 3) +
  scale_color_viridis_d(option = "viridis", name = "Contiguous Group ID") # Viridis is a good perceptually uniform palette

# Highlight the largest contiguous group
# (Optional: you can make these points larger or a different shape)
plot_scotoma <- plot_scotoma +
  geom_sf(data = largest_group, color = "black", size = 5) +
  theme(legend.position="none")

# Print the plot
#print(plot_all_points)

rttmp <- dBdat %>%
  filter(id=="RL35", visit==2, eye=="R", device=="retinalogik") %>%
  select(id, eye, x, y, dB)

hfatmp <- largest_group %>%
  mutate(x = st_coordinates(.)[, "X"],
         y = st_coordinates(.)[, "Y"]) %>%
  st_drop_geometry() 

mergeddat <- left_join(rttmp, hfatmp, by=c("id","eye","x","y")) %>%
  # remove blindspots and locations directly above and below blindspots
  # remove locations directly above and below blindspots
  filter(!(x==15 & y==3) & !(x==15 & y==-3) & !(x==15 & y==9) & !(x==15 & y==-9))

mergeddat <- mergeddat %>%
  merge(my_data, by=c("id","eye","x","y","psd_p"))

rtplot <- mergeddat %>%
  ggplot(aes(x, y, psd_p, dB.x)) +
  geom_raster(aes(x = x, y = y, fill = factor(psd_p))) +
  geom_text(aes(label = dB.x, x = x, y = y), size = 3) +
  coord_fixed(ratio = 1) +
  scale_fill_manual(values = rev(brewer.pal(3, "YlOrRd")), na.value="gray") +
  theme_bw() +
  ggtitle("Retinalogik", subtitle = "Threshold sensitivity dB values") +
  theme(legend.position="none")

hfaplot <- mergeddat %>%
  ggplot(aes(x, y, psd_p, dB)) +
  geom_raster(aes(x = x, y = y, fill = factor(psd_p))) +
  geom_text(aes(label = dB, x = x, y = y), size = 3) +
  coord_fixed(ratio = 1) +
  scale_fill_manual(values = rev(brewer.pal(3, "YlOrRd")), na.value="gray") +
  theme_bw() +
  ggtitle("HFA", subtitle = "Threshold sensitivity dB values") +
  theme(legend.position="none")

# mergeddat %>%
#   ggplot(aes(x, y, psd_p)) +
#   geom_raster(aes(x = x, y = y, fill = factor(psd_p))) +
#   geom_text(aes(label = psd_p, x = x, y = y), size = 3) +
#   coord_fixed(ratio = 1) +
#   scale_fill_manual(values = customcolors, na.value="gray") +
#   theme_bw() +
#   ggtitle("Retinalogik", subtitle = "Threshold sensitivity dB values") +
#   theme(legend.position="none")

(p + plot_scotoma) / (hfaplot + rtplot)

```
